### 2022-11-30 analysis/paper/paper.Rmd
Begin roughing out demographic projection paper.

---
title: "Caribou Demographi Projection - interesting & informative title TBD"
author:
  - name: Josie Hughes
    email: josie.hughes@ec.gc.ca
    institute: [ECCC]
    correspondence: true
  - name: Sarah Endicott
    email: sarah.endicott@ec.gc.ca
    institute: [ECCC]
    correspondence: false
  - name: Cheryl A Johnson
    email: cheryl-ann.johnson@ec.gc.ca
    institute: [ECCC]
    correspondence: false
institute:
  - ECCC: Environment and Climate Change Canada 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      number_sections: no
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: caribouDemography.bib
csl: "../templates/journal-of-forest-ecology-and-management.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
editor_options: 
  chunk_output_type: inline
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->
<!-- Link to data here: https://drive.google.com/file/d/179lw6MPK4l3xGmded9kofzJkB6VlJVb9/view?usp=sharing -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

<!-- The actual document text starts here: -->

# Introduction

Topics of interest, questions, goals:

* How get a demographic forecast informed by local demographic data and national disturbance-demographic relationships?
* How to assess the value of collecting additional demographic data? 
* How does that value depend on current status of the range, available existing demographic data, and projected future disturbance. When and why is more information required, and when do we know enough?
* Provide decision support tools to enable others to get forecasts informed by their local data and national models, and assess the value of monitoring scenarios.

# Methods

```{r setup2}
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(purrr)
library(ggpubr)
library(MetBrewer)
library(here)
library(tmap)
library(sf)

devtools::load_all(here::here())

#set paths
data_path_raw <- "analysis/data/raw_data/"
data_path_drvd <- "analysis/data/derived_data"
app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"

pal_nm <- "Redon"

fig_widths <- c(min = 1.18, single = 3.543, mid = 5.51, two = 7.48)

runScenarios =F 
```
TO DO: build large table of all parameters as the various sections are written. Adjust UI for consistent terminology along the way.

## A simple Bayesian IPM to integrate local demographic data and national disturbance-demographic relationships

TO DO: follow [@eacker_web-based_2019] description, noting changes.

Note alternative exponential survival analysis method - put results in appendix, show it doesn't make much difference.

## Calibrating to results from national model

Explain method for selecting values for remaining priors

Write for now assuming that caribouMetrics paper will be published somewhere.

## A method for simulating example data

### True population dynamics
Write for now assuming that caribouMetrics paper will be published somewhere.

### Observed survival from collars

### Observed recruitment from calf:cow aerial surveys

## Assessing the value of information

TO DO: how to assess or quantify value in this context. Power to distinguish increasing from decreasing population? What else?

## Scenarios

Table 1. Example ranges. Ontario demographic information from [@mnrf_state_2014]. If two numbers given for Anthropogenic disturbance they are BEAD 2010 and 2015.

```{r}
tbl <- "Range|Anthropogenic Disturbance|$\\lambda$|R|S|Survey Years
Missisa|0.41|0.86|14.2|80%|2008-2012
Churchill||0.96|19.5|87%|2011-2012
"
df <- read.delim(textConnection(tbl),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df) <- unname(as.list(df[1,]))
df <- df[-1,] # remove first row
row.names(df)<-NULL
knitr::kable(df)
```
TO DO: id examples from elsewhere to complement Ontario examples. Would be good to include an example where status is ok. What else, and why? WBI NWT ranges?

NOTE: would be great to info on number of collars per year in each example range. Though note MNRF recruitment estimates in some years came from aerial distribution surveys.

Table 2. Monitoring scenarios adapted from [@rettie_summary_2017].
```{r}
tbl <- "Years|Initial collars|Min collars|cow multiplier|collaring month|collar duration
3|25|25|3|1|3
"
df <- read.delim(textConnection(tbl),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
names(df) <- unname(as.list(df[1,]))
df <- df[-1,] # remove first row
row.names(df)<-NULL
knitr::kable(df)
```
TO DO: fill in set of monitoring scenarios when we have decided what they are.

TO DO: figure out disturbance forecast scenarios. Can lift Mississa scenario from caribouMetrics paper. Get Churchill scenario from Colin's Lac Seul model with roads - can lift it from the Churchill full model if not easy to sub in Lac Seul.

TO DO: Fire scenarios?

## Decision support tool development

Integration into caribouMetrics, shiny UI, integration into SpaDES & SyncroSim.

# Results

## Example figure

```{r scn1, cache=TRUE,echo=FALSE, message=FALSE,results='hide'}
if(runScenarios){
  numObsYrs=c(2);startsByYr = 30 #25
  scns=expand.grid(P=numObsYrs,sQ=c(0.5),st=startsByYr)
  scResults = runScenario(scns)

  png(here::here("analysis/figures/Figure1_Example.png"),
    height = 4, width = fig_widths["mid"], units = "in",res=600)
  print(plotRes(scResults$rr.summary.all, "Female population size",obs=scResults$obs.all,
              lowBound=0,highBound=2000,facetVars=c("P","sQ")))
  dev.off()
}

```

```{r fig1, fig.cap="Example figure.", cache=TRUE}
knitr::include_graphics(here::here("analysis/figures/Figure1_Example.png"))

```

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

# Appendix

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2022-11-30 R/CaribouDemographicFunctions.R

@@ -1,0 +1,48 @@
+#############################################################################
+
+#################################################################################
+
+
+# functions for caribou demographics paper
+
+# run scenario - temporary wrapper
+runScenario<-function(scns){
+  retdir <-getwd()
+  wdir <- app_path
+
+  dir.create("figs")
+  dir.create("tabs")
+
+  source(paste0(wdir,"/CaribouDemoFns.R"))
+
+  # usage
+  packages <- c("shiny", "R2jags","gdata","mcmcplots",
+                "ggplot2", "rmarkdown", "knitr",
+                "RODBC","plyr","survival","gdata","data.table")
+
+  ipak(packages)
+  load.module("glm")
+
+  library(caribouMetrics)
+  library(tidyr)
+  library(dplyr)
+
+  eParsIn = list()
+  eParsIn$cowCounts <- data.frame(Year = 1981:2018,
+                                  Count = 100,
+                                  Class = "cow")
+  eParsIn$freqStartsByYear <- data.frame(Year = 1981:2018,
+                                         numStarts = 30)
+  eParsIn$collarOnTime=1
+  eParsIn$collarOffTime=12
+  eParsIn$collarNumYears=1
+
+  ##########
+  #Get full set of sims for comparison
+  simBig<-getSimsNational(wdir=wdir)#If called with default parameters, use saved object to speed things up.
+
+  setwd(wdir)
+  scResults = runScnSet(scns,eParsIn,simBig)
+  setwd(retdir)
+  return(scResults)
+}

### 2022-12-05 analysis/paper/paper.Rmd
update
Work on Bayesian model description.

@@
       - --lua-filter=../templates/pagebreak.lua
-bibliography: caribouDemography.bib
+bibliography: CaribouDemographicModel.bib
 csl: "../templates/journal-of-forest-ecology-and-management.csl" # Insert path for the bib-style
@@
 
+Motivating example - forecasting impacts of disturbance - WBI and RoF example. In order to reduce uncertainty, need to integrate local information on current state of the population into forecasts.
+
+In Mississa (and other places) available local data is old and limited.
+Not clear how indicative older data is of state of population.
+Need new data to resolve questions, but what type, how much is enough.
+Can't monitor in great detail everywhere all the time.
+More generally, need for tools to help assess the value of additional monitoring, given the info that is already available.
+
 Topics of interest, questions, goals:
@@
 
-runScenarios =F 
+runScenarios =T 
 ```
@@
 
-TO DO: follow [@eacker_web-based_2019] description, noting changes.
+Following [@eacker_web-based_2019], the observed number of calves $c_t$ is a binomially distributed function of the observed number of adult female caribou $a_t$ and the estimated recruitment probability $R_t$: $$c_t \sim R_t^{c_t}(1-R_t)^{a_t-c_t}.$$
+We model recruitment probability as a function of anthropogenic disturbance and fire [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022], with a log link and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a anthro_t+\beta^R_f fireExclAnthro_t+\epsilon^R_t; \epsilon^R_t \sim \text{normal}(0,\sigma^2_{R,e}).$$
+Estimated recruitment probability is adjusted for sex ratio and delayed age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012] to get expected recruits per female: $$\tilde{R}_t=\frac{R_t/2}{1+R_t/2}.$$
+
+Also following [@eacker_web-based_2019], Kaplan-Meier estimates of observed adult female survival probability $\hat{S}_t$ and associated precisions $\tau_t$ are estimated from known-fate radio collar data using the survival R package [@terry_m_therneau_modeling_2000], with precisions for years with no observed mortality estimated using a separate Bayesian model.
+Observed survival is a Gaussian distributed function of the true survival probability: $$\hat{S_t} \sim \text{normal}(S_t,1/\tau_t).$$
+We model true survival probability as an adjusted function of anthropogenic disturbance  [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022], with a log link and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 S_{adj,t}-0.5)/45,1); \log(S_{adj,t})=\beta^S_0+\beta^S_a anthro_t+\epsilon^S_t; \epsilon^S_t \sim \text{normal}(0,\sigma^2_{S,e}).$$
+To assess sensitivity of results to the survival estimation method, we also implemented an alternative parametric exponential survival model, wherein the state of an individual animal in year $t$ and month $m$ is Bernoulli distributed (ref IPM book): $$A_{t,m} \sim \text{Bernoulli}(S^{1/12}_tA_{t,m-1}).$$
+
+All regression coefficients are assumed to be Gaussian distributed, with priors derived from the expected values and 95% confidence intervals estimated by Johnson et al (2020) (Table X).
+We calibrated the prior distributions so that the 95% prediction interval from the IPM matches the range between the 2.5% and 97.5% quantiles of 1000 simulated recruitment
+
+TO DO: ensure that INF_saf_INF_r_JAGS_extentTime.txt and INF_saf_INF_r_JAGS_expSurvival2.txt are identical except for the survival observation model.
+Or better yet, write code to sub in correct survival model more elegantly.
+TO DO: note various bits of truncation in the JAGS code to ensure R and S remain between 0 and 1 (not guaranteed give log link).
+Not all documented in this explanation.
+Bayesian wizard might know a better way.
+Or at least know how to explain.
+TO DO: note Eacker survival method gives biased results in years when <12 months of survival data is available.
+TO DO: Note I am dropping survival information for animals that are not alive at the beginning of a year to deal with censoring issues.
+There is probably some more elegant way to deal with this, but it won't make any difference as long as the collar deployment month is set to 1.
+TO DO: add references to documentation of popGrowthJohnson and other caribouMetrics functions.
 

### 2022-12-05 analysis/paper/paper.Rmd
tweaks

@@
 
-Also following [@eacker_web-based_2019], Kaplan-Meier estimates of observed adult female survival probability $\hat{S}_t$ and associated precisions $\tau_t$ are estimated from known-fate radio collar data using the survival R package [@terry_m_therneau_modeling_2000], with precisions for years with no observed mortality estimated using a separate Bayesian model.
-Observed survival is a Gaussian distributed function of the true survival probability: $$\hat{S_t} \sim \text{normal}(S_t,1/\tau_t).$$
-We model true survival probability as an adjusted function of anthropogenic disturbance  [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022], with a log link and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 S_{adj,t}-0.5)/45,1); \log(S_{adj,t})=\beta^S_0+\beta^S_a anthro_t+\epsilon^S_t; \epsilon^S_t \sim \text{normal}(0,\sigma^2_{S,e}).$$
-To assess sensitivity of results to the survival estimation method, we also implemented an alternative parametric exponential survival model, wherein the state of an individual animal in year $t$ and month $m$ is Bernoulli distributed (ref IPM book): $$A_{t,m} \sim \text{Bernoulli}(S^{1/12}_tA_{t,m-1}).$$
+Also following [@eacker_web-based_2019], Kaplan-Meier estimates of observed adult female survival probability $\hat{S}_t$ and associated precisions $\tau_t$ are estimated from known-fate radio collar data using the survival R package [@therneau_modeling_2000], with precisions for years with no observed mortality estimated using a separate Bayesian model.
+Observed estimated survival is a Gaussian distributed function of the true survival probability: $$\hat{S_t} \sim \text{normal}(S_t,1/\tau_t).$$
+We model true survival probability as an adjusted function of anthropogenic disturbance  [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022], with a log link and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a anthro_t+\epsilon^S_t; \epsilon^S_t \sim \text{normal}(0,\sigma^2_{S,e}).$$
+To assess sensitivity of results to the survival estimation method, we also implemented an alternative parametric exponential survival model, wherein the state of an individual animal in year $t$ and month $m$ is Bernoulli distributed (ref IPM book): $$A_{t,m} \sim \text{Bernoulli}(S^{1/12}_t A_{t,m-1}).$$
 
@@
 
-TO DO: ensure that INF_saf_INF_r_JAGS_extentTime.txt and INF_saf_INF_r_JAGS_expSurvival2.txt are identical except for the survival observation model.
+TO DO:
+
+* ensure that INF_saf_INF_r_JAGS_extentTime.txt and INF_saf_INF_r_JAGS_expSurvival2.txt are identical except for the survival observation model.
 Or better yet, write code to sub in correct survival model more elegantly.
-TO DO: note various bits of truncation in the JAGS code to ensure R and S remain between 0 and 1 (not guaranteed give log link).
+* note various bits of truncation in the JAGS code to ensure R and S remain between 0 and 1 (not guaranteed given log link).
 Not all documented in this explanation.
@@
 Or at least know how to explain.
-TO DO: note Eacker survival method gives biased results in years when <12 months of survival data is available.
-TO DO: Note I am dropping survival information for animals that are not alive at the beginning of a year to deal with censoring issues.
+* note Eacker survival method gives biased results in years when <12 months of survival data is available.
+* Note I am dropping survival information for animals that are not alive at the beginning of a year to deal with censoring issues.
 There is probably some more elegant way to deal with this, but it won't make any difference as long as the collar deployment month is set to 1.
-TO DO: add references to documentation of popGrowthJohnson and other caribouMetrics functions.
+* add references to documentation of popGrowthJohnson and other caribouMetrics functions.
 

### 2022-12-06 analysis/paper/paper.Rmd
in progress - documenting priors

@@
 Following [@eacker_web-based_2019], the observed number of calves $c_t$ is a binomially distributed function of the observed number of adult female caribou $a_t$ and the estimated recruitment probability $R_t$: $$c_t \sim R_t^{c_t}(1-R_t)^{a_t-c_t}.$$
-We model recruitment probability as a function of anthropogenic disturbance and fire [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022], with a log link and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a anthro_t+\beta^R_f fireExclAnthro_t+\epsilon^R_t; \epsilon^R_t \sim \text{normal}(0,\sigma^2_{R,e}).$$
+We model recruitment probability as a function of anthropogenic disturbance and fire [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022], with a log link and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a anthro_t+\beta^R_f fireExclAnthro_t+\epsilon^R_t; \epsilon^R_t \sim \text{normal}(0,\sigma^2_{R}).$$
 Estimated recruitment probability is adjusted for sex ratio and delayed age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012] to get expected recruits per female: $$\tilde{R}_t=\frac{R_t/2}{1+R_t/2}.$$
@@
 Observed estimated survival is a Gaussian distributed function of the true survival probability: $$\hat{S_t} \sim \text{normal}(S_t,1/\tau_t).$$
-We model true survival probability as an adjusted function of anthropogenic disturbance  [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022], with a log link and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a anthro_t+\epsilon^S_t; \epsilon^S_t \sim \text{normal}(0,\sigma^2_{S,e}).$$
+We model true survival probability as an adjusted function of anthropogenic disturbance  [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022], with a log link and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a anthro_t+\epsilon^S_t; \epsilon^S_t \sim \text{normal}(0,\sigma^2_{S}).$$
 To assess sensitivity of results to the survival estimation method, we also implemented an alternative parametric exponential survival model, wherein the state of an individual animal in year $t$ and month $m$ is Bernoulli distributed (ref IPM book): $$A_{t,m} \sim \text{Bernoulli}(S^{1/12}_t A_{t,m-1}).$$
 
-All regression coefficients are assumed to be Gaussian distributed, with priors derived from the expected values and 95% confidence intervals estimated by Johnson et al (2020) (Table X).
-We calibrated the prior distributions so that the 95% prediction interval from the IPM matches the range between the 2.5% and 97.5% quantiles of 1000 simulated recruitment
+All regression coefficients are assumed to be Gaussian distributed, with priors derived from the expected values and 95% confidence intervals estimated by Johnson et al (2020) (Table \@ref(tab:priors)).
+We calibrated the prior distributions so that the 95% prediction interval from the IPM matches the range between the 2.5% and 97.5% quantiles of 1000 simulated recruitment (Supplement X).
+
+```{r priors, echo=FALSE}
+
+source(here::here("analysis/scripts/formatPriors.R"))
+
+knitr::kable(tbl,caption="Model priors parameters are Gaussian distributed with expected values ")
+```
 

### 2022-12-06 analysis/scripts/formatPriors.R

@@ -1,0 +1,40 @@
+
+#app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
+library(caribouMetrics)
+source(paste0(app_path,"/CaribouDemoFns.R"))
+
+rawPriors<-getPriors(returnValues=F)
+
+
+str(rawPriors)
+#tbl <- data.frame()
+"Parameter|Description|Mean|Standard Deviation
+"
+
+nameSet <- gsub(".Prior1","",names(rawPriors)[grepl("Prior1",names(rawPriors))],fixed=T)
+
+start=T
+for(n in nameSet){
+  print(n)
+  crow = data.frame(name=n,Mean=rawPriors[[paste0(n,".Prior1")]],SD=rawPriors[[paste0(n,".Prior2")]])
+
+  if(start){
+    tbl = crow
+    start=F
+  }else{
+    tbl=rbind(tbl,crow)
+  }
+}
+
+tbl
+
+pars <- data.frame(name="l.R",Parameter="$\beta^R_0$",Description="R intercept")
+pars<-rbind(pars,data.frame(name="beta.Rec.anthro",Parameter="$\beta^R_a$",Description="R anthro slope"))
+pars<-rbind(pars,data.frame(name="beta.Rec.fire",Parameter="$\beta^R_f$",Description="R fire slope"))
+pars<-rbind(pars,data.frame(name="sig.R",Parameter="$\\sigma^2_{R}$",Description="R random effect"))
+pars <- rbind(pars,data.frame(name="l.Saf",Parameter="$\beta^S_0$",Description="S intercept"))
+pars<-rbind(pars,data.frame(name="beta.Saf",Parameter="$\beta^S_a$",Description="S anthro slope"))
+pars<-rbind(pars,data.frame(name="sig.Saf",Parameter="$\\sigma^2_{S}$",Description="S random effect"))
+
+tbl<-merge(pars,tbl);tbl$name=NULL
+

### 2022-12-07 analysis/paper/paper.Rmd
little edits to paper intro

@@
 ---
-title: "Caribou Demographi Projection - interesting & informative title TBD"
+title: "Caribou Demographic Projection - interesting & informative title TBD"
 author:
@@
 
-In Mississa (and other places) available local data is old and limited.
+Lots of variability among caribou ranges in monitoring data: recency, duration, types of data.
+E.g.
+27+ consecutive years of both telemetry-survival and calf/cow-ratios in most Alberta ranges, vs.
+1-4 years of data in most Ontario ranges (and that's old) *note we may want to be careful about inter-province comparisons for political reasons, but the variation is super relevant when to-date we've had the same 'rules' apply across the board...
+
+In Missisa (and other places) available local data is old and limited.
 Not clear how indicative older data is of state of population.
@@
 * How to assess the value of collecting additional demographic data? 
+* maybe also something like "How much of a bias/misinterpretation can be (has been?) incurred if we don't recognize local differences"?
 * How does that value depend on current status of the range, available existing demographic data, and projected future disturbance.
@@
 TO DO: build large table of all parameters as the various sections are written. Adjust UI for consistent terminology along the way.
+
+## Caribou population monitoring: survival and recruitment metrics
+
+Add in some brief description of the traditional approaches to estimating caribou survival (telemetry) and recruitment (aerial surveys/calf-cow ratios) - and show how these have been key for estimating population growth and status in most major boreal caribou projections/status documents.
+This could be useful background for defining terms below e.g.
+Rt, ct, at
 

### 2022-12-07 analysis/paper/paper.Rmd
more random comments on paper from Anna

@@
 TO DO: id examples from elsewhere to complement Ontario examples. Would be good to include an example where status is ok. What else, and why? WBI NWT ranges?
+*from Anna: i don't have access to actual estimates of R and S for different ranges, but some potentially interesting other ranges to compare to could be an Alberta range or two, where there are ++years of R and S estimates but really low lambda (E.g.
+ESAR and WSAR have 27+ yrs consecutive monitoring but declining numbers and high disturbance), some sections of NWT1 that are thought to be self-sustaining and have pretty good data sets (e.g.
+Dehcho North and South, Hay River Lowlands with 13+yrs of S and R) -could also use other MB ranges as examples for fairly sparse data (though not as bad as ON), and QC or LAB ranges as examples of good data (e.g.
+how about Charlevoix for a very well monitored but tiny pop, or Labrador pops have long term monitoring data but uncertain status...)
 
 NOTE: would be great to info on number of collars per year in each example range. Though note MNRF recruitment estimates in some years came from aerial distribution surveys.
+*from Anna: the monitoring data summary I collated a few years ago doesn't include numbers of collars, just numbers of years of data of each type...
+I imagine some provinces would be more willing than others to share those details?
+(we tried for ages to get more details and eventually gave up)
 

### 2022-12-07 analysis/paper/paper.Rmd
explanation of priors in progress
more little notes in paper document

@@
 
-Note alternative exponential survival analysis method - put results in appendix, show it doesn't make much difference.
-
-## Calibrating to results from national model
-
-Explain method for selecting values for remaining priors
-
-Write for now assuming that caribouMetrics paper will be published somewhere.
-
 ## A method for simulating example data
@@
 TO DO: fill in set of monitoring scenarios when we have decided what they are.
+*from Anna: happy to chat about what these could be! (e.g. see suggestions above for examples of other ranges...)
 

### 2022-12-12 analysis/paper/paper.Rmd
Adding prior calibration appendix.

@@
     email: cheryl-ann.johnson@ec.gc.ca
+    institute: [ECCC]
+    correspondence: false
+  - name: Anna Calvert
+    email: anna.calvert@ec.gc.ca
     institute: [ECCC]
@@
   comment = "#>",
-  fig.path = "../figures/",
+  fig.path = "../figs/",
   dpi = 300
@@
 
-runScenarios =T 
+runScenarios = T 
+doPriorCalibration=T
 ```
@@
 
-```{r priors, echo=FALSE}
+```{r priors, echo=FALSE,message=FALSE}
 
@@
 
-  png(here::here("analysis/figures/Figure1_Example.png"),
+  png(here::here("figs/Figure1_Example.png"),
     height = 4, width = fig_widths["mid"], units = "in",res=600)
@@
 ```{r fig1, fig.cap="Example figure.", cache=TRUE}
-knitr::include_graphics(here::here("analysis/figures/Figure1_Example.png"))
+#knitr::include_graphics(here::here("figs/Figure1_Example.png"))
 
@@
 
+## Calibration of Priors
+
+```{r priorCalibration, cache=TRUE,echo=FALSE, message=FALSE,results='hide'}
+if(doPriorCalibration){
+
+  #TO DO: move CaribouDemo_v1.15Reduce.R to paper directory. And produce pngs, rather than pdfs.
+  source(here::here("analysis/scripts/calibratePriors.R"))
+}else{
+  KS<-read.csv(here::here("tabs/SurvKSAll.csv"))
+}
+  png(here::here("figs/PriorKSDists.png"),
+    height = 4, width = fig_widths["two"], units = "in",res=600)
+    base = ggplot(KS,aes(x=lse,y=KSDistance,color=sse))+
+    geom_point()+facet_grid(scn~ssv,labeller="label_both",scales="free_y")+theme_bw()
+    print(base)
+  dev.off()
+
+```
+
+```{r PriorKSDists, fig.cap="Prior parameter values should be set so that, given minimal monitoring data ('All' scenario: 1 year of 1 collar per year), the posterior survival distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022].
+Low values of the mean and standard deviations of the prior distribution of the random effect of year on slope (sse=0.034784, ssv=0.01, $\\sigma^2_S$ in Table @ref(tab:priors)) meet this condition (Fig @ref(fig:priorAllFits).
+Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 300 collars per year from 97.5% and 2.5% quantiles of the national model), posterior survival distributions reflect local survival, rather than priors from the national model.
+When the multiplier on the prior standard deviation of the survival intercept (lse: $\\beta^S_0$ in Table @ref(tab:priors)) is < 5, the prior is not sufficiently vague to meet the latter condition." , cache=TRUE}
+knitr::include_graphics(here::here("figs/PriorKSDists.png"))
+
+#TO DO: figure out what to do about citations and references in figure captions. Or just fill them in manually at a later stage.
+
+```
+
+```{r PriorAllFits, fig.cap="Differences between posterior distributions from the Bayesian model and simulated outcomes from the national model, assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
+The multiplier on the prior standard deviation of the survival intercept ($\\beta^S_0$ in Table @ref(tab:priors)) is set to 5 in these examples.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.01.png"))
+
+```
+
+```{r PriorHighFits, fig.cap="Differences between posterior distributions from the Bayesian model and simulated outcomes from the national model, assuming intensive monitoring (15 years of 300 collars per year) of populations from the 97.5% quantile of the national model.
+The multiplier on the prior standard deviation of the survival intercept ($\\beta^S_0$ in Table @ref(tab:priors)) is set to 5 in these examples.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvsQStarts300highlse5.png"))
+
+```
+
 \newpage
@@
 # which R packages and versions?
-if ("devtools" %in% installed.packages()) devtools::session_info()
+#if ("devtools" %in% installed.packages()) devtools::session_info()
 ```
@@
 # what commit is this file at? 
-if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
+#if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
 ```

### 2022-12-12 R/CaribouDemographicFunctions.R

@@ -6,7 +6,7 @@
 # functions for caribou demographics paper
 
 # run scenario - temporary wrapper
-runScenario<-function(scns){
+runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMeier"){
   retdir <-getwd()
   wdir <- app_path
 
@@ -39,10 +39,16 @@ runScenario<-function(scns){
 
   ##########
   #Get full set of sims for comparison
-  simBig<-getSimsNational(wdir=wdir)#If called with default parameters, use saved object to speed things up.
+  if(is.null(quants)){
+    simBig<-getSimsNational(wdir=wdir)#If called with default parameters, use saved object to speed things up.
+
+  }else{
+    simBig<-getSimsNational(quants=quants,Anthro=Anthro)#If called with default parameters, use saved object to speed things up.
+
+  }
 
   setwd(wdir)
-  scResults = runScnSet(scns,eParsIn,simBig)
+  scResults = runScnSet(scns,eParsIn,simBig,survAnalysisMethod)
   setwd(retdir)
   return(scResults)
 }

### 2022-12-12 analysis/scripts/calibratePriors.R

@@ -1,0 +1,285 @@
+#app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
+
+
+#vary lse, sse, and ssv in factorial array. Outcome of interest is KS distance
+lse=seq(1,10,by=2);sse=0.08696*seq(0.2,1,by=0.2);ssv=seq(0.01,0.09,by=0.02)
+
+#Looking for low KS distance from full range of input sims when given almost no info
+numObsYrs=c(1);startsByYr = 1;J=1
+scns=expand.grid(P=numObsYrs,aSf=0,J=J,st=startsByYr,lse=lse,sse=sse,ssv=ssv)
+scResults = runScenario(scns,survAnalysisMethod = "Exponential")
+
+KSAll = subset(scResults$ksDists,(Year==2019)&(Parameter=="Adult female survival"))
+addBit = paste0("sQStarts",startsByYr)
+
+makeInterceptPlots(scResults,addBit=addBit,facetVars=c("lse","sse"),loopVars="ssv",whichPlots=c("Adult female survival"))
+
+#Also looking for low KS distance from range of input sims that match the observed data when given lots of observed data
+#Case 1 - low survival/recruitment quantile
+#Case 2 - high survival/recruitment quantile
+
+numObsYrs=c(15);startsByYr = 300;J=1;cw=200;sQ=0.025;rQ=0.025
+scns=expand.grid(P=numObsYrs,aSf=0,J=J,sQ=sQ,rQ=rQ,st=startsByYr,cw=cw,lse=lse,sse=sse,ssv=ssv)
+scResultsLow = runScenario(scns,quants=c(0.025,0.025),Anthro=0,survAnalysisMethod = "KaplanMeier")
+addBit = paste0("sQStarts",startsByYr,"low")
+KSLow = subset(scResultsLow$ksDists,(Year==2019)&(Parameter=="Adult female survival"))
+makeInterceptPlots(scResultsLow,addBit=addBit,facetVars=c("lse","sse"),loopVars="ssv",whichPlots=c("Adult female survival"))
+
+numObsYrs=c(15);startsByYr = 300;J=1;cw=200;sQ=0.975;rQ=0.975
+scns=expand.grid(P=numObsYrs,aSf=0,J=J,sQ=sQ,rQ=rQ,st=startsByYr,cw=cw,lse=lse,sse=sse,ssv=ssv)
[…]

### 2022-12-12 analysis/scripts/formatPriors.R

@@ -5,8 +5,6 @@ source(paste0(app_path,"/CaribouDemoFns.R"))
 
 rawPriors<-getPriors(returnValues=F)
 
-
-str(rawPriors)
 #tbl <- data.frame()
 "Parameter|Description|Mean|Standard Deviation
 "
@@ -15,7 +13,6 @@ nameSet <- gsub(".Prior1","",names(rawPriors)[grepl("Prior1",names(rawPriors))],
 
 start=T
 for(n in nameSet){
-  print(n)
   crow = data.frame(name=n,Mean=rawPriors[[paste0(n,".Prior1")]],SD=rawPriors[[paste0(n,".Prior2")]])
 
   if(start){
@@ -26,14 +23,12 @@ for(n in nameSet){
   }
 }
 
-tbl
-
-pars <- data.frame(name="l.R",Parameter="$\beta^R_0$",Description="R intercept")
-pars<-rbind(pars,data.frame(name="beta.Rec.anthro",Parameter="$\beta^R_a$",Description="R anthro slope"))
-pars<-rbind(pars,data.frame(name="beta.Rec.fire",Parameter="$\beta^R_f$",Description="R fire slope"))
+pars <- data.frame(name="l.R",Parameter="$\\beta^R_0$",Description="R intercept")
+pars<-rbind(pars,data.frame(name="beta.Rec.anthro",Parameter="$\\beta^R_a$",Description="R anthro slope"))
+pars<-rbind(pars,data.frame(name="beta.Rec.fire",Parameter="$\\beta^R_f$",Description="R fire slope"))
 pars<-rbind(pars,data.frame(name="sig.R",Parameter="$\\sigma^2_{R}$",Description="R random effect"))
-pars <- rbind(pars,data.frame(name="l.Saf",Parameter="$\beta^S_0$",Description="S intercept"))
-pars<-rbind(pars,data.frame(name="beta.Saf",Parameter="$\beta^S_a$",Description="S anthro slope"))
+pars <- rbind(pars,data.frame(name="l.Saf",Parameter="$\\beta^S_0$",Description="S intercept"))
+pars<-rbind(pars,data.frame(name="beta.Saf",Parameter="$\\beta^S_a$",Description="S anthro slope"))
 pars<-rbind(pars,data.frame(name="sig.Saf",Parameter="$\\sigma^2_{S}$",Description="S random effect"))
 
 tbl<-merge(pars,tbl);tbl$name=NULL

### 2022-12-13 analysis/paper/paper.Rmd
More work on prior calibration

@@
 
-runScenarios = T 
-doPriorCalibration=T
+runScenarios = F 
+doPriorCalibration=F #Note the prior calibration is time consuming. 
 ```
@@
   KS<-read.csv(here::here("tabs/SurvKSAll.csv"))
+
 }
@@
 Low values of the mean and standard deviations of the prior distribution of the random effect of year on slope (sse=0.034784, ssv=0.01, $\\sigma^2_S$ in Table @ref(tab:priors)) meet this condition (Fig @ref(fig:priorAllFits).
-Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 300 collars per year from 97.5% and 2.5% quantiles of the national model), posterior survival distributions reflect local survival, rather than priors from the national model.
-When the multiplier on the prior standard deviation of the survival intercept (lse: $\\beta^S_0$ in Table @ref(tab:priors)) is < 5, the prior is not sufficiently vague to meet the latter condition." , cache=TRUE}
+Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 45 collars per year from 97.5% and 2.5% quantiles of the national model), posterior survival distributions reflect local survival, rather than priors from the national model.
+When the multiplier on the prior standard deviation of the survival intercept (lse: $\\beta^S_0$ in Table @ref(tab:priors)) is < 5, the prior is not sufficiently vague to meet the latter condition.
+Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
 knitr::include_graphics(here::here("figs/PriorKSDists.png"))
@@
 ```{r PriorAllFits, fig.cap="Differences between posterior distributions from the Bayesian model and simulated outcomes from the national model, assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
-The multiplier on the prior standard deviation of the survival intercept ($\\beta^S_0$ in Table @ref(tab:priors)) is set to 5 in these examples.", cache=TRUE}
+In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.01, and  anthropogenic disturbance is 0.", cache=TRUE}
 knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.01.png"))
@@
 
-```{r PriorHighFits, fig.cap="Differences between posterior distributions from the Bayesian model and simulated outcomes from the national model, assuming intensive monitoring (15 years of 300 collars per year) of populations from the 97.5% quantile of the national model.
-The multiplier on the prior standard deviation of the survival intercept ($\\beta^S_0$ in Table @ref(tab:priors)) is set to 5 in these examples.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts300highlse5.png"))
+```{r PriorHighFits, fig.cap="Differences between posterior distributions from the Bayesian model and simulated outcomes from the national model, assuming intensive monitoring (15 years of 45 collars per year) of populations from the 97.5% quantile of the national model.
+In these examples, the multiplier on the prior standard deviation of the survival intercept (lse:$\\beta^S_0$ in Table @ref(tab:priors)) is 5 and  anthropogenic disturbance is 0.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvsQStarts45highlse5.png"))
 
+```
+
+```{r PriorAllFitsAnthro90, fig.cap="Differences between posterior distributions from the Bayesian model and simulated outcomes from the national model, assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
+In these examples, anthropogenic disturbance is 90%.
+When the multiplier on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) is 3, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvsQStarts1Anthro90.png"))
+```
+
+```{r PriorHighFitsAnthro90, fig.cap="Differences between posterior distributions from the Bayesian model and simulated outcomes from the national model, assuming intensive monitoring (15 years of 45 collars per year) of populations from the 97.5% quantile of the national model.
+In these examples, anthropogenic disturbance is 90%, and the effect of anthropogenic disturbance varies from none (sS=0) through expected (sS=1) to double the expected slope (sS=2).
+All considered multipliers on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvsQStarts45Anthro90high.png"))
 ```

### 2022-12-13 R/CaribouDemographicFunctions.R

@@ -6,7 +6,7 @@
 # functions for caribou demographics paper
 
 # run scenario - temporary wrapper
-runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMeier"){
+runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMeier",popGrowthTable=NULL){
   retdir <-getwd()
   wdir <- app_path
 
@@ -39,12 +39,11 @@ runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMei
 
   ##########
   #Get full set of sims for comparison
+
   if(is.null(quants)){
-    simBig<-getSimsNational(wdir=wdir)#If called with default parameters, use saved object to speed things up.
-
+    simBig<-getSimsNational(wdir=wdir,popGrowthTable = NULL)#If called with default parameters, use saved object to speed things up.
   }else{
-    simBig<-getSimsNational(quants=quants,Anthro=Anthro)#If called with default parameters, use saved object to speed things up.
-
+    simBig<-getSimsNational(quants=quants,Anthro=Anthro,popGrowthTable=NULL)#If called with default parameters, use saved object to speed things up.
   }
 
   setwd(wdir)

### 2022-12-13 analysis/scripts/calibratePriors.R

@@ -1,5 +1,5 @@
 #app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
-
+#devtools::load_all(here::here())
 
 #vary lse, sse, and ssv in factorial array. Outcome of interest is KS distance
 lse=seq(1,10,by=2);sse=0.08696*seq(0.2,1,by=0.2);ssv=seq(0.01,0.09,by=0.02)
@@ -18,21 +18,23 @@ makeInterceptPlots(scResults,addBit=addBit,facetVars=c("lse","sse"),loopVars="ss
 #Case 1 - low survival/recruitment quantile
 #Case 2 - high survival/recruitment quantile
 
-numObsYrs=c(15);startsByYr = 300;J=1;cw=200;sQ=0.025;rQ=0.025
+numObsYrs=c(15);startsByYr = 45;J=1;cw=200;sQ=0.025;rQ=0.025
 scns=expand.grid(P=numObsYrs,aSf=0,J=J,sQ=sQ,rQ=rQ,st=startsByYr,cw=cw,lse=lse,sse=sse,ssv=ssv)
 scResultsLow = runScenario(scns,quants=c(0.025,0.025),Anthro=0,survAnalysisMethod = "KaplanMeier")
 addBit = paste0("sQStarts",startsByYr,"low")
 KSLow = subset(scResultsLow$ksDists,(Year==2019)&(Parameter=="Adult female survival"))
-makeInterceptPlots(scResultsLow,addBit=addBit,facetVars=c("lse","sse"),loopVars="ssv",whichPlots=c("Adult female survival"))
+makeInterceptPlots(scResultsLow,addBit=addBit,facetVars=c("lse","sse"),loopVars="ssv",
+                   whichPlots=c("Adult female survival"),useNational=F)
 
-numObsYrs=c(15);startsByYr = 300;J=1;cw=200;sQ=0.975;rQ=0.975
+numObsYrs=c(15);startsByYr = 45;J=1;cw=200;sQ=0.975;rQ=0.975
 scns=expand.grid(P=numObsYrs,aSf=0,J=J,sQ=sQ,rQ=rQ,st=startsByYr,cw=cw,lse=lse,sse=sse,ssv=ssv)
 scResultsHigh = runScenario(scns,quants=c(0.975,0.975),Anthro=0,survAnalysisMethod = "KaplanMeier")
 addBit = paste0("sQStarts",startsByYr,"high")
[…]

### 2022-12-14 analysis/paper/paper.Rmd
Ongoing prior calibration work: recruitment

--- 2022-12-12 analysis/paper/paper.Rmd
@@
   KS<-read.csv(here::here("tabs/SurvKSAll.csv"))
+  KSRec<-read.csv(here::here("tabs/RecKSAll.csv"))
+
 }
@@
     base = ggplot(KS,aes(x=lse,y=KSDistance,color=sse))+
-    geom_point()+facet_grid(scn~ssv,labeller="label_both",scales="free_y")+theme_bw()
+    geom_point()+facet_grid(scn~ssv,labeller="label_both",scales="free_y")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("lse: multiplier on SD of survival intercept")
     print(base)
@@
 
+  png(here::here("figs/PriorKSDistsRec.png"),
+    height = 4, width = fig_widths["two"], units = "in",res=600)
+    base = ggplot(KSRec,aes(x=lre,y=KSDistance,color=sre))+
+    geom_point()+facet_grid(scn~srv,labeller="label_both",scales="free_y")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("lre: multiplier on SD of recruitment intercept")
+    print(base)
+  dev.off()
 ```
@@
 ```{r PriorKSDists, fig.cap="Prior parameter values should be set so that, given minimal monitoring data ('All' scenario: 1 year of 1 collar per year), the posterior survival distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022].
-Low values of the mean and standard deviations of the prior distribution of the random effect of year on slope (sse=0.034784, ssv=0.01, $\\sigma^2_S$ in Table @ref(tab:priors)) meet this condition (Fig @ref(fig:priorAllFits).
-Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 300 collars per year from 97.5% and 2.5% quantiles of the national model), posterior survival distributions reflect local survival, rather than priors from the national model.
-When the multiplier on the prior standard deviation of the survival intercept (lse: $\\beta^S_0$ in Table @ref(tab:priors)) is < 5, the prior is not sufficiently vague to meet the latter condition." , cache=TRUE}
+Low values of the mean and standard deviations of the prior distribution of the random effect of year on survival (sse=0.034784, ssv=0.01, $\\sigma^2_S$ in Table @ref(tab:priors)) meet this condition (Fig @ref(fig:priorAllFits).
+Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 45 collars per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local survival, rather than priors from the national model.
+When the multiplier on the prior standard deviation of the survival intercept (lse: $\\beta^S_0$ in Table @ref(tab:priors)) is < 5, the prior is not sufficiently vague to meet the latter condition.
+Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
 knitr::include_graphics(here::here("figs/PriorKSDists.png"))
@@
 #TO DO: figure out what to do about citations and references in figure captions. Or just fill them in manually at a later stage.
-
 ```
 
-```{r PriorAllFits, fig.cap="Differences between posterior distributions from the Bayesian model and simulated outcomes from the national model, assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
-The multiplier on the prior standard deviation of the survival intercept ($\\beta^S_0$ in Table @ref(tab:priors)) is set to 5 in these examples.", cache=TRUE}
+```{r PriorAllFits, fig.cap="Differences between posterior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
+In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.01, and  anthropogenic disturbance is 0.
+Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.",  cache=TRUE}
 knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.01.png"))
-
 ```
 
-```{r PriorHighFits, fig.cap="Differences between posterior distributions from the Bayesian model and simulated outcomes from the national model, assuming intensive monitoring (15 years of 300 collars per year) of populations from the 97.5% quantile of the national model.
-The multiplier on the prior standard deviation of the survival intercept ($\\beta^S_0$ in Table @ref(tab:priors)) is set to 5 in these examples.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts300highlse5.png"))
+```{r PriorHighFits, fig.cap="Posterior predictive survival distributions from the Bayesian model ('local') assuming intensive monitoring (15 years of 45 collars per year) of populations from the 97.5% quantile of the national model.
+In these examples, the multiplier on the prior standard deviation of the survival intercept (lse:$\\beta^S_0$ in Table @ref(tab:priors)) is 5 and  anthropogenic disturbance is 0.
+Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvsQStarts45highlse5.png"))
+```
 
+```{r PriorAllFitsAnthro90, fig.cap="Differences between posterior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
+In these examples, anthropogenic disturbance is 90%.
+When the multiplier on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) is 3, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model.
+Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvsQStarts1Anthro90.png"))
+```
+
+```{r PriorHighFitsAnthro90, fig.cap="Posterior predictive survival distributions from the Bayesian model, assuming intensive monitoring (15 years of 45 collars per year) of populations from the 97.5% quantile of the national model.
+In these examples, anthropogenic disturbance is 90%, and the effect of anthropogenic disturbance varies from none (sS=0) through expected (sS=1) to double the expected slope (sS=2).
+All considered multipliers on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.
+Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvsQStarts45Anthro90high.png"))
+```
+
+```{r PriorKSDistsRec, fig.cap="Prior parameter values should be set so that, given no monitoring data ('All' scenario), the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022].
+XX values of the mean and standard deviations of the prior distribution of the random effect of year on recruitment (sre=XX, srv=XX, $\\sigma^2_R$ in Table @ref(tab:priors)) meet this condition (Fig @ref(fig:priorAllFits).
+Priors should also be vague enough so that, given intensive monitoring of populations where recruitment is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 135 cows per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local recruitment, rather than priors from the national model.
+When the multiplier on the prior standard deviation of the recruitment intercept (lre: $\\beta^R_0$ in Table @ref(tab:priors)) is XX, the prior is not sufficiently vague to meet the latter condition.
+Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
+knitr::include_graphics(here::here("figs/PriorKSDistsRec.png"))
+```
+
+```{r PriorAllFitsRec, fig.cap="Differences between posterior predictive recruitment distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no monitoring.
+In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is XX, and  anthropogenic disturbance is 0.
+Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.",  cache=TRUE}
+knitr::include_graphics(here::here("figs/RecrQStarts1srv0.22.png"))
+```
+
+```{r PriorHighFitsRec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model ('local') assuming intensive monitoring (135 cows per year for 15 years) of populations from the 97.5% quantile of the national model.
+In these examples, the multiplier on the prior standard deviation of the recruitment intercept (lre:$\\beta^R_0$ in Table @ref(tab:priors)) is XX and anthropogenic disturbance is 0.
+Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
+knitr::include_graphics(here::here("figs/RecrQStarts45highsrv0.22.png"))
+```
+
+```{r PriorAllFitsAnthro90Rec, fig.cap="Differences between posterior recruitment predictive distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no monitoring.
+In these examples, anthropogenic disturbance is 90%.
+When the multiplier on the prior standard deviation of the slope of the relationship between recruitment and anthropogenic disturbance (bre: $\\beta^R_a$ in Table @ref(tab:priors)) is XX, the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model.
+Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvsQStarts1Anthro90.png"))
+```
+
+```{r PriorHighFitsAnthro90Rec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model, assuming intensive monitoring (15 years of 135 collars per year) of populations from the 97.5% quantile of the national model.
+In these examples, anthropogenic disturbance is 90%, and the effect of anthropogenic disturbance varies from none (rS=0) through expected (rS=1) to double the expected slope (rS=2).
+XX multipliers on the prior standard deviation of the survival slope (bre: $\\beta^R_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.
+Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvsQStarts45Anthro90high.png"))
 ```

### 2022-12-14 analysis/scripts/calibratePriors.R

@@ -1,6 +1,8 @@
 #app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
 #devtools::load_all(here::here())
 
+########################
+#Survival
 #vary lse, sse, and ssv in factorial array. Outcome of interest is KS distance
 lse=seq(1,10,by=2);sse=0.08696*seq(0.2,1,by=0.2);ssv=seq(0.01,0.09,by=0.02)
 
@@ -8,15 +10,9 @@ lse=seq(1,10,by=2);sse=0.08696*seq(0.2,1,by=0.2);ssv=seq(0.01,0.09,by=0.02)
 numObsYrs=c(1);startsByYr = 1;J=1
 scns=expand.grid(P=numObsYrs,aSf=0,J=J,st=startsByYr,lse=lse,sse=sse,ssv=ssv)
 scResults = runScenario(scns,survAnalysisMethod = "Exponential")
-
 KSAll = subset(scResults$ksDists,(Year==2019)&(Parameter=="Adult female survival"))
 addBit = paste0("sQStarts",startsByYr)
-
 makeInterceptPlots(scResults,addBit=addBit,facetVars=c("lse","sse"),loopVars="ssv",whichPlots=c("Adult female survival"))
-
-#Also looking for low KS distance from range of input sims that match the observed data when given lots of observed data
-#Case 1 - low survival/recruitment quantile
-#Case 2 - high survival/recruitment quantile
 
 numObsYrs=c(15);startsByYr = 45;J=1;cw=200;sQ=0.025;rQ=0.025
 scns=expand.grid(P=numObsYrs,aSf=0,J=J,sQ=sQ,rQ=rQ,st=startsByYr,cw=cw,lse=lse,sse=sse,ssv=ssv)
@@ -31,38 +27,22 @@ scns=expand.grid(P=numObsYrs,aSf=0,J=J,sQ=sQ,rQ=rQ,st=startsByYr,cw=cw,lse=lse,s
 scResultsHigh = runScenario(scns,quants=c(0.975,0.975),Anthro=0,survAnalysisMethod = "KaplanMeier")
 addBit = paste0("sQStarts",startsByYr,"high")
 KSHigh = subset(scResultsHigh$ksDists,(Year==2019)&(Parameter=="Adult female survival"))
-#source("CaribouDemoFns.R")
-
[…]

### 2022-12-15 analysis/paper/paper.Rmd
tweaking priors

@@
 ```{r PriorKSDists, fig.cap="Prior parameter values should be set so that, given minimal monitoring data ('All' scenario: 1 year of 1 collar per year), the posterior survival distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022].
-Low values of the mean and standard deviations of the prior distribution of the random effect of year on survival (sse=0.034784, ssv=0.01, $\\sigma^2_S$ in Table @ref(tab:priors)) meet this condition (Fig @ref(fig:priorAllFits).
 Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 45 collars per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local survival, rather than priors from the national model.
-When the multiplier on the prior standard deviation of the survival intercept (lse: $\\beta^S_0$ in Table @ref(tab:priors)) is < 5, the prior is not sufficiently vague to meet the latter condition.
+When the multiplier on the prior standard deviation of the survival intercept (lse: $\\beta^S_0$ in Table @ref(tab:priors)) is 5, the prior is informative enough to meet the former condition, and vague enough to meet the latter.
+Given lse=5, low values of the mean standard deviations of the prior distribution of the random effect of year on survival (sse=0.034784, ssv=0.01, $\\sigma^2_S$ in Table @ref(tab:priors)) yield priors that meet both conditions (Figs @ref(fig:priorAllFits) and @ref(fig:priorLowFits)).
 Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
@@
 
-```{r PriorHighFits, fig.cap="Posterior predictive survival distributions from the Bayesian model ('local') assuming intensive monitoring (15 years of 45 collars per year) of populations from the 97.5% quantile of the national model.
-In these examples, the multiplier on the prior standard deviation of the survival intercept (lse:$\\beta^S_0$ in Table @ref(tab:priors)) is 5 and  anthropogenic disturbance is 0.
+```{r PriorLowFits, fig.cap="Posterior predictive survival distributions from the Bayesian model ('local') assuming intensive monitoring (15 years of 45 collars per year) of populations from the 2.5% quantile of the national model.
+In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.01, and  anthropogenic disturbance is 0.
 Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts45highlse5.png"))
+knitr::include_graphics(here::here("figs/SurvsQStarts45lowssv0.01.png"))
 ```
@@
 ```{r PriorKSDistsRec, fig.cap="Prior parameter values should be set so that, given no monitoring data ('All' scenario), the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022].
-XX values of the mean and standard deviations of the prior distribution of the random effect of year on recruitment (sre=XX, srv=XX, $\\sigma^2_R$ in Table @ref(tab:priors)) meet this condition (Fig @ref(fig:priorAllFits).
+Setting the multiplier on the prior standard deviation of the recruitment intercept (lre: $\\beta^R_0$ in Table @ref(tab:priors)) to 3, and values of the mean and standard deviations of the prior distribution of the random effect of year on recruitment to sre=0.23 and srv=0.22, respectively ($\\sigma^2_R$ in Table @ref(tab:priors)) allows us to meet this condition (Fig @ref(fig:priorAllFitsRec)).
 Priors should also be vague enough so that, given intensive monitoring of populations where recruitment is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 135 cows per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local recruitment, rather than priors from the national model.
-When the multiplier on the prior standard deviation of the recruitment intercept (lre: $\\beta^R_0$ in Table @ref(tab:priors)) is XX, the prior is not sufficiently vague to meet the latter condition.
+When the multiplier on the prior standard deviation of the recruitment intercept (lre: $\\beta^R_0$ in Table @ref(tab:priors)) is 1, the prior is not sufficiently vague to meet the latter condition (Fig @ref(fig:PriorLowFitsRec)).
 Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
@@
 ```{r PriorAllFitsRec, fig.cap="Differences between posterior predictive recruitment distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no monitoring.
-In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is XX, and  anthropogenic disturbance is 0.
+In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0.
 Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.",  cache=TRUE}
@@
 
-```{r PriorHighFitsRec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model ('local') assuming intensive monitoring (135 cows per year for 15 years) of populations from the 97.5% quantile of the national model.
-In these examples, the multiplier on the prior standard deviation of the recruitment intercept (lre:$\\beta^R_0$ in Table @ref(tab:priors)) is XX and anthropogenic disturbance is 0.
+```{r PriorLowFitsRec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model ('local') assuming intensive monitoring (135 cows per year for 15 years) of populations from the 2.5% quantile of the national model.
+In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0.
 Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/RecrQStarts45highsrv0.22.png"))
+knitr::include_graphics(here::here("figs/RecrQStarts45lowsrv0.22.png"))
 ```

### 2022-12-15 analysis/paper/paper.Rmd
setting up to knit on Anna's machine

@@
 data_path_drvd <- "analysis/data/derived_data"
-app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
+
+cd <- here::here()
+app_path <- gsub("Caribou-Demographic-Projection-Paper","BayesianCaribouDemographicProjection",cd,fixed=T)
 

### 2022-12-15 analysis/paper/paper.Rmd
Finishing prior calibration

@@
 In these examples, anthropogenic disturbance is 90%.
-When the multiplier on the prior standard deviation of the slope of the relationship between recruitment and anthropogenic disturbance (bre: $\\beta^R_a$ in Table @ref(tab:priors)) is XX, the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model.
+When the multiplier on the prior standard deviation of the slope of the relationship between recruitment and anthropogenic disturbance (bre: $\\beta^R_a$ in Table @ref(tab:priors)) is 4, the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model.
 Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts1Anthro90.png"))
+knitr::include_graphics(here::here("figs/RecrQStarts1Anthro90.png"))
 ```
@@
 In these examples, anthropogenic disturbance is 90%, and the effect of anthropogenic disturbance varies from none (rS=0) through expected (rS=1) to double the expected slope (rS=2).
-XX multipliers on the prior standard deviation of the survival slope (bre: $\\beta^R_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.
+All multipliers on the prior standard deviation of the survival slope (bre: $\\beta^R_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.
 Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts45Anthro90high.png"))
+knitr::include_graphics(here::here("figs/RecrQStarts45Anthro90high.png"))
 ```

### 2022-12-15 analysis/scripts/calibratePriors.R

@@ -110,10 +110,10 @@ write.csv(KS,here::here(paste0("tabs/RecKSAll.csv")))
 
 ###############
 #vary bse to get slope.
-bre=seq(1,10,by=2);rS=c(1);iA=90
+bre=seq(1,7,by=1.5);rS=c(1);iA=90
 
-numObsYrs=c(1);startsByYr = 1;J=1
-scns=expand.grid(P=numObsYrs,aSf=0,J=J,st=startsByYr,bre=bre,rS=rS,iA=iA)
+numObsYrs=c(1);startsByYr = 1;J=1;cw=0
+scns=expand.grid(P=numObsYrs,aSf=0,J=J,st=startsByYr,cw=cw,bre=bre,rS=rS,iA=iA)
 scResults = runScenario(scns,survAnalysisMethod = "Exponential")
 KSAll = subset(scResults$ksDists,(Year==2019)&(Parameter=="Recruitment"))
 addBit = paste0("rQStarts",startsByYr,"Anthro",iA)
@@ -142,235 +142,3 @@ KSHigh$scn = "High"
 sCols = c("KSDistance","bre","rS","scn")
 KSAnthro90 = rbind(subset(KSAll,select=sCols),subset(KSLow,select=sCols),subset(KSHigh,select=sCols))
 write.csv(KSAnthro90,here::here(paste0("tabs/RecKSAllAnthro90.csv")))
-
-
-if(0){
-  #Next - survival slope parameter sensitivity
-  #CaribouDemo_v1.15Reduce.R
-
-  numObsYrs=c(2,5,20);startsByYr = 25; iA=90
-  scns=expand.grid(P=numObsYrs,sQ=c(0.025,0.5,0.975),iA=iA,st=startsByYr)
-  scResults = runScnSet(scns,eParsIn,simBig)
-  addBit = paste0("sQStarts",startsByYr,"iA",iA)
-  makeInterceptPlots(scResults,addBit=addBit,facetVars=c("P","sQ"))
-
-  numObsYrs=c(2,5,20);startsByYr = 25; iA=90; bse=1
-  scns=expand.grid(P=numObsYrs,sQ=c(0.025,0.5,0.975),iA=iA,st=startsByYr,bse=bse)
-  scResults = runScnSet(scns,eParsIn,simBig)
-  addBit = paste0("sQStarts",startsByYr,"iA",iA,"bse",bse)
[…]

### 2022-12-16 analysis/paper/paper.Rmd
more prior calibration tweaks.

@@
 ```{r PriorAllFits, fig.cap="Differences between posterior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
-In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.01, and  anthropogenic disturbance is 0.
+In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0.
 Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.",  cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.01.png"))
+knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.03.png"))
 ```
@@
 ```{r PriorLowFits, fig.cap="Posterior predictive survival distributions from the Bayesian model ('local') assuming intensive monitoring (15 years of 45 collars per year) of populations from the 2.5% quantile of the national model.
-In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.01, and  anthropogenic disturbance is 0.
+In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0.
 Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts45lowssv0.01.png"))
+knitr::include_graphics(here::here("figs/SurvsQStarts45lowssv0.03.png"))
 ```

### 2022-12-16 R/CaribouDemographicFunctions.R

@@ -7,6 +7,7 @@
 
 # run scenario - temporary wrapper
 runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMeier",popGrowthTable=NULL){
+
   retdir <-getwd()
   wdir <- app_path
 
@@ -14,6 +15,8 @@ runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMei
   dir.create("tabs")
 
   source(paste0(wdir,"/CaribouDemoFns.R"))
+
+  scns<-fillDefaults(scns)
 
   # usage
   packages <- c("shiny", "R2jags","gdata","mcmcplots",
@@ -41,9 +44,9 @@ runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMei
   #Get full set of sims for comparison
 
   if(is.null(quants)){
-    simBig<-getSimsNational(wdir=wdir,popGrowthTable = NULL)#If called with default parameters, use saved object to speed things up.
+    simBig<-getSimsNational(wdir=wdir,popGrowthTable = NULL,adjustR=unique(scns$adjustR))#If called with default parameters, use saved object to speed things up.
   }else{
-    simBig<-getSimsNational(quants=quants,Anthro=Anthro,popGrowthTable=NULL)#If called with default parameters, use saved object to speed things up.
+    simBig<-getSimsNational(quants=quants,Anthro=Anthro,popGrowthTable=NULL,adjustR=unique(scns$adjustR))#If called with default parameters, use saved object to speed things up.
   }
 
   setwd(wdir)

### 2022-12-16 analysis/paper/paper.Rmd
some thoughts on Discussion

@@
 Dehcho North and South, Hay River Lowlands with 13+yrs of S and R) -could also use other MB ranges as examples for fairly sparse data (though not as bad as ON), and QC or LAB ranges as examples of good data (e.g.
-how about Charlevoix for a very well monitored but tiny pop, or Labrador pops have long term monitoring data but uncertain status...)
+how about Charlevoix for a very well monitored but tiny pop, northern QC for really broad-expanse pop, or Labrador pops have long term monitoring data but uncertain status...)
 
@@
 TO DO: fill in set of monitoring scenarios when we have decided what they are.
-*from Anna: happy to chat about what these could be! (e.g. see suggestions above for examples of other ranges...)
+*Anna thinking about this...
 
@@
 ```
+
+# Discussion
+
+Will need a section on the fundamental assumptions of these models/aspatial nature of them, e.g.
+that collared cows  and cows observed in surveys are a random representative sample of the larger population, and that there is full mixing in the population so that estimated survival/recruitment rates apply equally to all animals in the population.
+We can discuss this with some contrasting examples, e.g.
+collaring 30 animals from a very small/restricted population (like Little Smoky, 308,606ha, pop size estimate 78 as of COSEWIC 2014) would likely be much more representative of the population vs.
+collaring 30 animals from a relatively large/expansive population (northern Quebec range Q6, 62,156,186 ha, pop size 9000 as of COSEWIC 2014).
+But because of the way this model is built, we assume that the sample is equally representative in either of those cases, so changing the initial starting N won't affect model performance/predictions much here...
+
+We should likely also have a pgph about costs of monitoring - it'll be pretty short, cause basically we have no idea what it costs :) Through the NBCKC MOnitoring Working Group work, we tried over and over to get some basic cost estimates from P/T reps and practitioners, and hit a wall - nobody wanted to tell us a thing.
+Don't know if we'd now find anyone is more open to sharing, but don't think we should get our hopes up.
+But at the very least we'll want to acknowledge that we haven't accounted for relative costs of monitoring...
 

### 2022-12-16 analysis/paper/paper.Rmd
more notes and thoughts from Anna...

@@
 
-Table 2. Monitoring scenarios adapted from [@rettie_summary_2017].
+Table 2.
+Monitoring scenarios adapted from [@rettie_summary_2017].
+*maybe we could remove the last couple columns if they're not things we're going to be varying...?
+(same for min collars if the same as init?)
 ```{r}

### 2022-12-16 analysis/paper/paper.Rmd
reconciling

--- 2022-12-16 analysis/paper/paper.Rmd
@@
 In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0.
-Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.",  cache=TRUE}
+Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
 knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.03.png"))
@@
 When the multiplier on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) is 3, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model.
-Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.", cache=TRUE}
+Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
 knitr::include_graphics(here::here("figs/SurvsQStarts1Anthro90.png"))
@@
 In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0.
-Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.",  cache=TRUE}
+Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
 knitr::include_graphics(here::here("figs/RecrQStarts1srv0.22.png"))
@@
 When the multiplier on the prior standard deviation of the slope of the relationship between recruitment and anthropogenic disturbance (bre: $\\beta^R_a$ in Table @ref(tab:priors)) is 4, the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model.
-Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 500 sample populations from the national model.", cache=TRUE}
+Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
 knitr::include_graphics(here::here("figs/RecrQStarts1Anthro90.png"))

### 2022-12-23 analysis/scripts/sensitivityMinimal.R
cloud setup

@@ -1,11 +1,16 @@
+#!/usr/bin/env Rscript
+
 #Run scenario batches on cloud machine
+#assuming workingDir is where we are, cmd line invocation is
+#Rscript --vanilla ./analysis/scripts/sensitivityMinimal.R 3
 
-workingDir = "C:/Users/HughesJo/Documents/gitprojects/Caribou-Demographic-Projection-Paper"
-toolDir = "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
-libDir = .libPaths()[1]
-cpageId=2 #which batch?
+args = commandArgs(trailingOnly=TRUE)
+
+workingDir = "C:/Users/HughesJo/Documents/cloudDeploymentSandbox/Caribou-Demographic-Projection-Paper"
+toolDir = "C:/Users/HughesJo/Documents/cloudDeploymentSandbox/BayesianCaribouDemographicProjection"
+libDir = "C:/Users/HughesJo/Documents/cloudDeploymentSandbox/Rpackages"
+cpageId=args[1] #which batch?
 setName = "s1"
-
 
 #######################
 setwd(workingDir)
@@ -14,8 +19,8 @@ dir.create(paste0("tabs/",setName),recursive=T)
 dir.create(paste0("results/",setName),recursive=T)
 
 packages <- c("R2jags","gdata","mcmcplots",
-              "ggplot2","RODBC","plyr","survival","gdata","data.table",
-              "tidyr","dplyr","caribouMetrics")
+              "ggplot2","RODBC","plyr","dplyr","survival","gdata","data.table",
+              "tidyr","caribouMetrics")
 
 for(p in packages){library(p,lib.loc=libDir,character.only=T)}
 

### 2022-12-23 analysis/scripts/sensitivity1.R
Refining figures and anthropogenic disturbance scenarios.

@@ -12,7 +12,12 @@ setName = "s1"
 minimalScn = expand.grid(P=1,st=1,cmult=0,ri=1,assessmentYrs=1)
 
 monitoringScns = expand.grid(P=c(1,2,4,8,16),st=c(30,60),cmult=c(2,3,4),ri=c(1,2,4),assessmentYrs=c(1,3))
-stateScns = expand.grid(iA=c(0),rep=seq(1:100))
+stateScns = data.frame(tA=c(0,20,40,60,0,20,40,60),aS=c(0,1,1,1,0,0,0,0),aSf=c(1,1,1,1,0,0,0,0))
+stateScns = merge(stateScns,data.frame(rep=seq(1:100)))
+
+monitoringScns = expand.grid(P=c(1,4,16),st=c(60),cmult=c(3),ri=c(1),assessmentYrs=c(3))
+stateScns = data.frame(tA=c(0,20,40,60),aS=c(0,1,1,1),aSf=c(1,1,1,1))
+stateScns = merge(stateScns,data.frame(rep=seq(1:10)))
 
 #monitoringScns = expand.grid(P=c(1,4,8,16),st=c(30,60),cmult=c(3),ri=c(1,2,4),assessmentYrs=1)
 #stateScns = expand.grid(iA=c(0),rep=seq(1:30))
@@ -22,9 +27,10 @@ stateScns$rQ = runif(nrow(stateScns),min=0.01,max=0.99)
 monitoringScns=rbind(monitoringScns,minimalScn)
 scns=merge(monitoringScns,stateScns)
 
+scns$iA = scns$tA-scns$P*scns$aS+1
+
 scns$N0 = 10000
 nrow(scns)
-str(scns)
 #scns=scns[10,]
 #scResults = readRDS("temp.Rds")
 
@@ -37,7 +43,7 @@ if(0){
   runTimeFull # of processor days, roughly.
 }else{runTime10=1.27}
 
-scns$pageLab = paste0("cmult",scns$cmult,"ay",scns$assessmentYrs)
+scns$pageLab = paste0("cmult",scns$cmult,"ay",scns$assessmentYrs,"aSf",scns$aSf)
 scns$pageId = as.numeric(as.factor(scns$pageLab))
 unique(scns$pageId)
 
@@ -60,9 +66,30 @@ for(pp in pages){
 
   saveRDS(scResults,paste0("results/r",pp,".Rds"))
 
[…]

### 2022-12-23 analysis/paper/paper.Rmd

@@
 Monitoring scenarios adapted from [@rettie_summary_2017].
-*maybe we could remove the last couple columns if they're not things we're going to be varying...?
-(same for min collars if the same as init?)
+*Anna will remove the last couple columns since we won't be varying those (same for min collars since it'll be the same as init collars)
 ```{r}

### 2022-12-23 analysis/scripts/sensitivity1.R
Setup for cloud

@@ -15,9 +15,9 @@ monitoringScns = expand.grid(P=c(1,2,4,8,16),st=c(30,60),cmult=c(2,3,4),ri=c(1,2
 stateScns = data.frame(tA=c(0,20,40,60,0,20,40,60),aS=c(0,1,1,1,0,0,0,0),aSf=c(1,1,1,1,0,0,0,0))
 stateScns = merge(stateScns,data.frame(rep=seq(1:100)))
 
-monitoringScns = expand.grid(P=c(1,4,16),st=c(60),cmult=c(3),ri=c(1),assessmentYrs=c(3))
-stateScns = data.frame(tA=c(0,20,40,60),aS=c(0,1,1,1),aSf=c(1,1,1,1))
-stateScns = merge(stateScns,data.frame(rep=seq(1:10)))
+#monitoringScns = expand.grid(P=c(1,4,16),st=c(60),cmult=c(3),ri=c(1),assessmentYrs=c(3))
+#stateScns = data.frame(tA=c(0,20,40,60),aS=c(0,1,1,1),aSf=c(1,1,1,1))
+#stateScns = merge(stateScns,data.frame(rep=seq(1:10)))
 
 #monitoringScns = expand.grid(P=c(1,4,8,16),st=c(30,60),cmult=c(3),ri=c(1,2,4),assessmentYrs=1)
 #stateScns = expand.grid(iA=c(0),rep=seq(1:30))
@@ -29,7 +29,7 @@ scns=merge(monitoringScns,stateScns)
 
 scns$iA = scns$tA-scns$P*scns$aS+1
 
-scns$N0 = 10000
+scns$N0 = 1000
 nrow(scns)
 #scns=scns[10,]
 #scResults = readRDS("temp.Rds")

### 2022-12-23 analysis/scripts/sensitivityMinimal.R
Working on cloud deployment

@@ -2,13 +2,14 @@
 
 #Run scenario batches on cloud machine
 #assuming workingDir is where we are, cmd line invocation is
-#Rscript --vanilla ./analysis/scripts/sensitivityMinimal.R 3
+#Rscript --vanilla ./cloudDeploymentSandbox/Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R 1
 
 args = commandArgs(trailingOnly=TRUE)
 
-workingDir = "C:/Users/HughesJo/Documents/cloudDeploymentSandbox/Caribou-Demographic-Projection-Paper"
-toolDir = "C:/Users/HughesJo/Documents/cloudDeploymentSandbox/BayesianCaribouDemographicProjection"
-libDir = "C:/Users/HughesJo/Documents/cloudDeploymentSandbox/Rpackages"
+baseDir = getwd()
+workingDir = paste0(baseDir,"/cloudDeploymentSandbox/Caribou-Demographic-Projection-Paper")
+toolDir = paste0(baseDir,"/cloudDeploymentSandbox/BayesianCaribouDemographicProjection")
+libDir = paste0(baseDir,"/cloudDeploymentSandbox/Rpackages")
 cpageId=args[1] #which batch?
 setName = "s1"
 

### 2023-01-13 analysis/paper/paper.Rmd
Anna adding some more stuff to various bits of the paper...

@@
 
-In Missisa (and other places) available local data is old and limited.
+In Missisa (and other places, esp Ontario) available local data are old and limited.
 Not clear how indicative older data is of state of population.
@@
 Can't monitor in great detail everywhere all the time.
-More generally, need for tools to help assess the value of additional monitoring, given the info that is already available.
+More generally, need for tools to help assess the value of additional monitoring and the best way to invest additional monitoring resources, given the info that is already available.
 
@@
 * How get a demographic forecast informed by local demographic data and national disturbance-demographic relationships?
-* How to assess the value of collecting additional demographic data? 
-* maybe also something like "How much of a bias/misinterpretation can be (has been?) incurred if we don't recognize local differences"?
+* How to assess the value of collecting additional demographic data? How to prioritize whether to e.g. monitor many areas for a short time or fewer areas over the long term? 
+* How much of a bias/misinterpretation can be (has been?) incurred if we don't recognize local differences?
 * How does that value depend on current status of the range, available existing demographic data, and projected future disturbance.
@@
 
-Add in some brief description of the traditional approaches to estimating caribou survival (telemetry) and recruitment (aerial surveys/calf-cow ratios) - and show how these have been key for estimating population growth and status in most major boreal caribou projections/status documents.
-This could be useful background for defining terms below e.g.
-Rt, ct, at
+Marking of caribou cows with telemetry GPS/VHF collars is the traditional approach to estimating demographic parameters (refs/examples, e.g.
+key boreal caribou projections and status documents reliant on these ests, also methods papers like Courtois et al.
+2003, RettieMessier 1998).
+
+Adult survival is estimated by tracking the fate of marked individuals (Kaplan-Meier ests) [brief description of field/logistical details - can get details from NBCKC Telemetry writeup (https://www.cclmportal.ca/sites/default/files/2021-10/Telemetry%20draft%20October%2027%202021.pdf), e.g.
+GPS allow for remote detection vs VHF-only require physical relocation of individual; collar sends out a
+mortality signal when no longer active after a predetermined period of time; males only occasionally collared; age of collared individuals is usually unknown]
+
+Collared females are also used in traditional estimates of recruitment (refs/examples): collared individuals are located by aircraft and then the entire group of caribou is surveyed to estimate the ratio of cows:calves in the group...
+Because this estimate is based on more than just the collared sample of cows, the sample size for recruitment ests is typically larger than for survival (introduce the multiplier parameter).
 
@@
 Monitoring scenarios adapted from [@rettie_summary_2017].
-*Anna will remove the last couple columns since we won't be varying those (same for min collars since it'll be the same as init collars)
+*removed the second collar column (assuming minimum=initial) and removed collaring month and duration, but added renewal interval, and hoping to also be able to provide some sort of example of real ranges that are similar to the scenario.
+Anna waiting to hear back from Monitoring WG folks to get a sense of what values they've been using so that we can make these values realistic.
 ```{r}
-tbl <- "Years|Initial collars|Min collars|cow multiplier|collaring month|collar duration
-3|25|25|3|1|3
+tbl <- "Years|Collars|cow multiplier|collar renewal interval|example
+3|30|3|1|ex range?
 "
@@
 ```
-TO DO: fill in set of monitoring scenarios when we have decided what they are.
-*Anna thinking about this...
 
@@
 
-We should likely also have a pgph about costs of monitoring - it'll be pretty short, cause basically we have no idea what it costs :) Through the NBCKC MOnitoring Working Group work, we tried over and over to get some basic cost estimates from P/T reps and practitioners, and hit a wall - nobody wanted to tell us a thing.
+We should likely also have a pgph about costs of monitoring - it'll be pretty short, cause basically we have no idea what it costs since nobody will tell us :) Through the NBCKC Monitoring Working Group work, we tried over and over to get some basic cost estimates from P/T reps and practitioners, and hit a wall - nobody wanted to tell us a thing.
 Don't know if we'd now find anyone is more open to sharing, but don't think we should get our hopes up.
 But at the very least we'll want to acknowledge that we haven't accounted for relative costs of monitoring...
+[e.g.
+costs outlined in the Telemetry guidance report from the MonitoringWG: "Expenses to consider in radio-collared monitoring will vary, based on: helicopter (or other aircraft) use, purchase of tracking collars, data fees associated with the collars (the cost of which varies by operating system, recording frequency and data transmission frequency), and live-capture equipment (3 or more sets of capture nets and associated buckets, a netting gun, and ankle hobbles)"]
 
@@
 #TO DO: figure out what to do about citations and references in figure captions. Or just fill them in manually at a later stage.
+
+#Note: file wouldn't knit for Anna without access to all the newest png files below, so just commenting out the problem ones for now
 ```
@@
 Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.03.png"))
+# COMMENTED OUT FOR NOW CAUSE I DIDN"T HAVE ACCESS
+#knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.03.png"))
 ```
@@
 Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts45lowssv0.03.png"))
+# COMMENTED OUT FOR NOW CAUSE I DIDN"T HAVE ACCESS
+#knitr::include_graphics(here::here("figs/SurvsQStarts45lowssv0.03.png"))
 ```

### 2023-01-18 analysis/paper/paper.Rmd
updates to paper draft based on MonitoringWG contacts' feedback re monitoring parameters

@@
 Table 1. Example ranges. Ontario demographic information from [@mnrf_state_2014]. If two numbers given for Anthropogenic disturbance they are BEAD 2010 and 2015.
+*note that we've had some input on monitoring numbers from NL, BC, ON (and MB pending), so some of them may also be willing to give us case study examples to use here?
 
@@
 NOTE: would be great to info on number of collars per year in each example range. Though note MNRF recruitment estimates in some years came from aerial distribution surveys.
-*from Anna: the monitoring data summary I collated a few years ago doesn't include numbers of collars, just numbers of years of data of each type...
-I imagine some provinces would be more willing than others to share those details?
-(we tried for ages to get more details and eventually gave up)
+*feedback from MonitoringWG contacts includes details on collar numbers for BC, ON, NL, (MB pending); e.g.
+in Ontario they aim for 20-30 collars in smaller pops, 35+ in bigger pops; Labrador aims to have between 5-25 collars monitored at any one time, depending on pop size; in BC, real numbers of collared animals have ranged been between 3-59 per pop/year, varying both across-yrs-within-pop and across-pop-within-yr [with COVID yrs by far the lowest].
 
@@
 *removed the second collar column (assuming minimum=initial) and removed collaring month and duration, but added renewal interval, and hoping to also be able to provide some sort of example of real ranges that are similar to the scenario.
-Anna waiting to hear back from Monitoring WG folks to get a sense of what values they've been using so that we can make these values realistic.
+*Monitoring WG input from BC, ON, NL, (MB pending): lots of detail provided but generally those numbers suggest we could try values of ~3-10 [e.g.
+3,6,9?] for the cow multiplier, 5-50 for collars [e.g.
+5,10,20,40,60?], [not sure about renewal interval, but see SaraM email for some insights].
+We know (e.g.
+from the MonitoringWG perspectives report/update to Rettie summary) that number of years of monitoring can be as low as 2 or as many as 20+
 ```{r}
@@
 Get Churchill scenario from Colin's Lac Seul model with roads - can lift it from the Churchill full model if not easy to sub in Lac Seul.
+*and can maybe ask MonitoringWG contacts who seemed keen whether they'd have case studies to offer too...?
 
@@
 
+On a related topic to population size; Note that the stuff we got from ArtR regarding simulations for sample size does acknowledge that pops of different sizes have different monitoring requirements, and Sara's Labrador numbers also show targets relative to pop size...
+Since this kind of aspatial model will mostly be using small population sizes in order to see any effects, we'll jsut have to acknowledge that a caveat of this is that we're not fully able to explore all the population size effects with this model, but generally Josie has found little impact of initial population size to the results.
+
 We should likely also have a pgph about costs of monitoring - it'll be pretty short, cause basically we have no idea what it costs since nobody will tell us :) Through the NBCKC Monitoring Working Group work, we tried over and over to get some basic cost estimates from P/T reps and practitioners, and hit a wall - nobody wanted to tell us a thing.
@@
 costs outlined in the Telemetry guidance report from the MonitoringWG: "Expenses to consider in radio-collared monitoring will vary, based on: helicopter (or other aircraft) use, purchase of tracking collars, data fees associated with the collars (the cost of which varies by operating system, recording frequency and data transmission frequency), and live-capture equipment (3 or more sets of capture nets and associated buckets, a netting gun, and ankle hobbles)"]
+-->perhaps the keen MonitoringWG contacts would be willing to provide even really rough cost estimates, and/or even jsut a relative sense of how different it would be to e.g.
+monitor a few pops for many yrs vs lots of pops for a short time...?
 

### 2023-01-19 analysis/paper/paper.Rmd
minor edits

@@
     correspondence: false
-  - name: Anna Calvert
+  - name: Anna M Calvert
     email: anna.calvert@ec.gc.ca

### 2023-01-19 analysis/paper/paper.Rmd
tiny edits

@@
 TO DO: id examples from elsewhere to complement Ontario examples. Would be good to include an example where status is ok. What else, and why? WBI NWT ranges?
-*from Anna: i don't have access to actual estimates of R and S for different ranges, but some potentially interesting other ranges to compare to could be an Alberta range or two, where there are ++years of R and S estimates but really low lambda (E.g.
+*from Anna: withouot access to actual estimates of R and S for different ranges, some potentially interesting other ranges to compare to could be an Alberta range or two, where there are ++years of R and S estimates but really low lambda (E.g.
 ESAR and WSAR have 27+ yrs consecutive monitoring but declining numbers and high disturbance), some sections of NWT1 that are thought to be self-sustaining and have pretty good data sets (e.g.

### 2023-01-20 analysis/paper/paper.Rmd
work on initial sensitivity analysis in progress.

@@
 * add references to documentation of popGrowthJohnson and other caribouMetrics functions.
+* consider & comment on alternative choices made in this example https://poissonconsulting.shinyapps.io/bboushiny/. Follow Eacker example in all particulars, change some aspects, or do comparisons?
 
@@
 *note that we've had some input on monitoring numbers from NL, BC, ON (and MB pending), so some of them may also be willing to give us case study examples to use here?
+Note can get details about SK1 from Phil's report.
 
@@
 
-Table 2.
-Monitoring scenarios adapted from [@rettie_summary_2017].
-*removed the second collar column (assuming minimum=initial) and removed collaring month and duration, but added renewal interval, and hoping to also be able to provide some sort of example of real ranges that are similar to the scenario.
+relevant refs: [@rettie_summary_2017];?
+
+```{r monitoringScenarios, echo=FALSE,message=FALSE}
+
+source(here::here("analysis/scripts/formatScenarios.R"))
+
+knitr::kable(tbl,caption="Monitoring scenarios parameters and values (example values to stimulate discussion - what should these numbers be and why?)")
+```
+
 *Monitoring WG input from BC, ON, NL, (MB pending): lots of detail provided but generally those numbers suggest we could try values of ~3-10 [e.g.
@@
 from the MonitoringWG perspectives report/update to Rettie summary) that number of years of monitoring can be as low as 2 or as many as 20+
-```{r}
-tbl <- "Years|Collars|cow multiplier|collar renewal interval|example
-3|30|3|1|ex range?
-"
-df <- read.delim(textConnection(tbl),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
-names(df) <- unname(as.list(df[1,]))
-df <- df[-1,] # remove first row
-row.names(df)<-NULL
-knitr::kable(df)
-```
 
@@
 
-TO DO: Fire scenarios?
+TO DO: Fire scenarios? Add high fire scenario in appendix of one by one sensitivity checks.
+
+TO DO: run low initial population size scenario, put in appendix of one by one sensitivity checks.
+
+TO DO: calculate change in anthro disturbance over 5 yrs across ranges to get realistic distribution of numbers of this & justify choice of parameters.
+run alternative scenario to put in one by one sensitivity appendix.
+
+```{r fig1, fig.cap="Anthropogenic disturbance and monitoring scenarios (examples to stimulate discussion).", cache=TRUE}
+knitr::include_graphics(here::here("figs/forPaper/distScnscmult3ay1aSf1.png"))
+
+```
 
@@
 
-## Example figure
+TO DO: consider showing results in terms of average number of collars deployed per year rather than target # and renewal interval.
 
@@
 }
-
-```
-
-```{r fig1, fig.cap="Example figure.", cache=TRUE}
-#knitr::include_graphics(here::here("figs/Figure1_Example.png"))
 

### 2023-01-20 R/CaribouDemographicFunctions.R

@@ -39,7 +39,7 @@ runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMei
                                          numStarts = 30)
   eParsIn$collarOnTime=1
   eParsIn$collarOffTime=12
-  eParsIn$collarNumYears=3
+  eParsIn$collarNumYears=6
 
   ##########
   #Get full set of sims for comparison

### 2023-01-20 analysis/scripts/formatScenarios.R

@@ -1,0 +1,29 @@
+
+#app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
+library(caribouMetrics)
+
+rawScns<-read.csv(here::here("tabs/s2.csv"))
+nrow(rawScns)
+rawScns = subset(rawScns,st!=1)
+
+monitoringPars = list(P=list(description = "monitoring duration",unit="years"),
+                      st=list(description="target # of collars", unit="cows"),
+                      cmult=list(description="calf:cow survey multiplier",unit = "cows per collared cow"),
+                      ri = list(description="time between collar deployments", unit="years"),
+                      assessmentYrs = list(description="assessment duration",unit="years"))
+
+start=T
+for(par in names(monitoringPars)){
+  #par="st"
+  crow = data.frame(name=par,description=monitoringPars[[par]]$description,
+                    units=monitoringPars[[par]]$unit,
+                    values = paste(unique(rawScns[,par]),collapse=","))
+
+  if(start){
+    tbl = crow
+    start=F
+  }else{
+    tbl=rbind(tbl,crow)
+  }
+}
+

### 2023-01-20 analysis/scripts/sensitivity1.R

@@ -8,12 +8,11 @@ scn_defaults <- c(eval(formals(fillDefaults)$defList),
 
 ########################
 #sensitivity
-setName = "s1"
-minimalScn = expand.grid(P=1,st=1,cmult=0,ri=1,assessmentYrs=1)
-
-monitoringScns = expand.grid(P=c(1,2,4,8,16),st=c(30,60),cmult=c(2,3,4),ri=c(1,2,4),assessmentYrs=c(1,3))
-stateScns = data.frame(tA=c(0,20,40,60,0,20,40,60),aS=c(0,1,1,1,0,0,0,0),aSf=c(1,1,1,1,0,0,0,0))
-stateScns = merge(stateScns,data.frame(rep=seq(1:100)))
+#setName = "s1"
+#minimalScn = expand.grid(P=1,st=1,cmult=0,ri=1,assessmentYrs=1)
+#monitoringScns = expand.grid(P=c(1,2,4,8,16),st=c(30,60),cmult=c(2,3,4),ri=c(1,2,4),assessmentYrs=c(1,3))
+#stateScns = data.frame(tA=c(0,20,40,60,0,20,40,60),aS=c(0,1,1,1,0,0,0,0),aSf=c(1,1,1,1,0,0,0,0))
+#stateScns = merge(stateScns,data.frame(rep=seq(1:100)))
 
 #monitoringScns = expand.grid(P=c(1,4,16),st=c(60),cmult=c(3),ri=c(1),assessmentYrs=c(3))
 #stateScns = data.frame(tA=c(0,20,40,60),aS=c(0,1,1,1),aSf=c(1,1,1,1))
@@ -22,9 +21,15 @@ stateScns = merge(stateScns,data.frame(rep=seq(1:100)))
 #monitoringScns = expand.grid(P=c(1,4,8,16),st=c(30,60),cmult=c(3),ri=c(1,2,4),assessmentYrs=1)
 #stateScns = expand.grid(iA=c(0),rep=seq(1:30))
 
+setName = "s2"
+monitoringScns = expand.grid(P=c(1,2,4,8,16,24),st=c(15,30,60),cmult=c(3,6,9),ri=c(1,4),assessmentYrs=c(3))
+stateScns = data.frame(tA=c(0,20,40,60,0,20,40,60),aS=c(0,1,1,1,0,0,0,0),aSf=c(1,1,1,1,0,0,0,0))
+stateScns = merge(stateScns,data.frame(rep=seq(1:100)))
+
+
[…]

### 2023-01-20 analysis/scripts/sensitivityMinimal.R

@@ -1,20 +1,29 @@
 #!/usr/bin/env Rscript
 
 #Run scenario batches on cloud machine
-#assuming workingDir is where we are, cmd line invocation is
+#assuming workingDir is where we are, cmd li{ne invocation is
 #Rscript --vanilla ./cloudDeploymentSandbox/Caribou-Demographic-Projection-Paper/Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R 1
+#Rscript --vanilla ./analysis/scripts/sensitivityMinimal.R 1 "local"
+#nohup Rscript --vanilla ./analysis/scripts/sensitivityMinimal.R 1 "local" &
 
 args = commandArgs(trailingOnly=TRUE)
 #args=1
-baseDir = getwd()
-print(baseDir)
-workingDir = paste0(baseDir,"/cloudDeploymentSandbox/Caribou-Demographic-Projection-Paper/Caribou-Demographic-Projection-Paper")
-#baseDir = "C:/Users/HughesJo/Documents/"; workingDir = paste0(baseDir,"/cloudDeploymentSandbox/Caribou-Demographic-Projection-Paper")
+if(args[2]!="local"){
+  baseDir = getwd()
+  print(baseDir)
+  workingDir = paste0(baseDir,"/cloudDeploymentSandbox/Caribou-Demographic-Projection-Paper/Caribou-Demographic-Projection-Paper")
+  toolDir = paste0(baseDir,"/cloudDeploymentSandbox/BayesianCaribouDemographicProjection")
+  libDir = paste0(baseDir,"/cloudDeploymentSandbox/Rpackages")
 
-toolDir = paste0(baseDir,"/cloudDeploymentSandbox/BayesianCaribouDemographicProjection")
-libDir = paste0(baseDir,"/cloudDeploymentSandbox/Rpackages")
+}else{
+  baseDir = "C:/Users/HughesJo/Documents"
+  workingDir = paste0(baseDir,"/gitprojects/Caribou-Demographic-Projection-Paper")
+  toolDir = paste0(baseDir,"/gitprojects/BayesianCaribouDemographicProjection")
+  libDir = paste0(baseDir,"/cloudDeploymentSandbox/Rpackages")
+}
+
 cpageId=args[1] #which batch?
-setName = "s1"
+setName = "s2"
 
 #######################
 setwd(workingDir)
@@ -49,7 +58,7 @@ setwd(toolDir)
 source("CaribouDemoFns.R")
[…]

### 2023-01-20 analysis/scripts/viewSensitivityResults.R

@@ -1,0 +1,112 @@
+#app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
+#devtools::load_all(here::here())
+library(tidyverse)
+library(ggplot2)
+source( paste0(app_path,"/CaribouDemoFns.R"))
+scn_defaults <- c(eval(formals(fillDefaults)$defList),
+                  curYear = eval(formals(fillDefaults)$curYear))
+
+########################
+#sensitivity
+setName = "s1"
+scns = read.csv(here::here("tabs/s1.csv"))
+scns = subset(scns,cmult!=0)
+pages=unique(scns$pageId)
+unique(scns$pageLab)
+
+for(cpageId in pages){
+  #cpageId=pages[1]
+  p = scns$pageLab[scns$pageId==cpageId]
+
+  if(!file.exists(paste0("results/",setName,"/r",cpageId,".Rds"))){next}
+  scResults = readRDS(paste0("results/",setName,"/r",cpageId,".Rds"))
+
+  probs = subset(scResults$rr.summary.all,(Parameter=="Population growth rate"))
+  probs$projectionTime = probs$Year-2023
+  probs$Anthro2023 = probs$tA
+  probs$AssessmentYear = probs$Year
+
+  #See disturbance scenarios
+  distScns = subset(probs,is.element(projectionTime,c(0,5,20))|(Year==iYr))
+  distScns$grp = paste0(distScns$Anthro2023,distScns$aSf)
+  distScns$Timeline[distScns$Year==2023]="Final monitoring year"
+  distScns$Timeline[distScns$projectionTime==5]="AssessmentYear 2028"
+  distScns$Timeline[distScns$projectionTime==20]="AssessmentYear 2043"
+  distScns$Timeline[distScns$Year<2023]= paste("Start",distScns$P[distScns$Year<2023],"yrs of monitoring")
+
+  startLevels = unique(distScns$Timeline[distScns$Year<2023])
+  distScns$Timeline = factor(distScns$Timeline,levels=c(startLevels,"Final monitoring year","AssessmentYear 2028","AssessmentYear 2043"))
+
+  png(here::here(paste0("figs/",setName,"/distScns",p,".png")),
+      height = 4, width = 5.51, units = "in",res=600)
[…]

### 2023-02-10 analysis/paper/paper.Rmd
ongoing.

@@
 ---
-title: "Caribou Demographic Projection - interesting & informative title TBD"
+title: "National demographic-disturbance relationships and local monitoring both reduce uncertainty in projections of boreal caribou population viability: A Bayesian integrated population model, demographic simulations and a Shiny interface show how"
 author:
@@
 <!-- The actual document text starts here: -->
+# Abstract
+Across fifty-eight boreal caribou ranges in Canada, both survival and recruitment decrease with the percentage of the range that is within five hundred metres of anthropogenic disturbance.
+However, there is also variation in demographic rates among ranges particularly where anthropogenic disturbance is low, and this variation leads to uncertainty in demographic projections derived from national disturbance-demographic relationships.
+Demographic projections derived from local range-specific demographic data are also uncertain because few places have enough long-term monitoring data to characterize local demographic-disturbance relationships.
+We present a simple Bayesian integrated population model to integrate prior information from national analysis of demographic-disturbance relationships with available local demographic data to reduce uncertainty in population viability projections.
+We combine this model with methods for simulating local population dynamics and monitoring to show how monitoring requirements depend on landscape condition.
+Where projected anthropogenic disturbance is very high or very low, reasonably accurate status projections can be made using only national demographic-disturbance relationships.
+Where uncertainty is higher, deploying few collars for more years may be a lower cost option for improving the reliability of status projections.
+We provide an R Shiny interface to enable others to explore the implications of alternative monitoring scenarios, disturbance trajectories, priors, and other model parameters for boreal caribou demographic projections.
 
@@
 
-knitr::kable(tbl,caption="Monitoring scenarios parameters and values (example values to stimulate discussion - what should these numbers be and why?)")
+knitr::kable(tbl,caption="Monitoring scenarios parameters and values.")
 ```
@@
 
-```{r fig1, fig.cap="Anthropogenic disturbance and monitoring scenarios (examples to stimulate discussion).", cache=TRUE}
-knitr::include_graphics(here::here("figs/forPaper/distScnscmult3ay1aSf1.png"))
+```{r fig1, fig.cap="Anthropogenic disturbance and monitoring scenarios.", cache=TRUE}
+knitr::include_graphics(here::here("figs/forPaper/distScnscmult3ay3aSf1.png"))
 
@@
 
+```
+
+```{r fig2, fig.cap="Power analysis example result.", cache=TRUE}
+knitr::include_graphics(here::here("figs/forPaper/powercmult3ay3aSf1.png"))
+knitr::include_graphics(here::here("figs/forPaper/powerEffortcmult3ay3aSf1.png"))
+
+#TO DO: figure out how to put no monitoring results on these graphs.
 ```

### 2023-02-10 analysis/scripts/sensitivity1.R

@@ -1,15 +1,10 @@
 #app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
 #devtools::load_all(here::here())
 
-library(ggplot2)
-source( paste0(app_path,"/CaribouDemoFns.R"))
-scn_defaults <- c(eval(formals(fillDefaults)$defList),
-                  curYear = eval(formals(fillDefaults)$curYear))
-
 ########################
 #sensitivity
 #setName = "s1"
-#minimalScn = expand.grid(P=1,st=1,cmult=0,ri=1,assessmentYrs=1)
+#minimalScn = expand.grid(P=1,st=1,cmult=0,ri=c(1,4),assessmentYrs=3)
 #monitoringScns = expand.grid(P=c(1,2,4,8,16),st=c(30,60),cmult=c(2,3,4),ri=c(1,2,4),assessmentYrs=c(1,3))
 #stateScns = data.frame(tA=c(0,20,40,60,0,20,40,60),aS=c(0,1,1,1,0,0,0,0),aSf=c(1,1,1,1,0,0,0,0))
 #stateScns = merge(stateScns,data.frame(rep=seq(1:100)))
@@ -21,138 +16,30 @@ scn_defaults <- c(eval(formals(fillDefaults)$defList),
 #monitoringScns = expand.grid(P=c(1,4,8,16),st=c(30,60),cmult=c(3),ri=c(1,2,4),assessmentYrs=1)
 #stateScns = expand.grid(iA=c(0),rep=seq(1:30))
 
-setName = "s2"
-monitoringScns = expand.grid(P=c(1,2,4,8,16,24),st=c(15,30,60),cmult=c(3,6,9),ri=c(1,4),assessmentYrs=c(3))
+setName = "s3"
+monitoringScns = expand.grid(P=c(1,2,4,8,16,24),st=c(1,15,30,60),cmult=c(3,6,9),ri=c(1,4),assessmentYrs=c(3))
 stateScns = data.frame(tA=c(0,20,40,60,0,20,40,60),aS=c(0,1,1,1,0,0,0,0),aSf=c(1,1,1,1,0,0,0,0))
-stateScns = merge(stateScns,data.frame(rep=seq(1:100)))
-
+stateScns = merge(stateScns,data.frame(rep=seq(1:200)))
 
 stateScns$sQ=runif(nrow(stateScns),min=0.01,max=0.99)
[…]

### 2023-02-10 analysis/scripts/sensitivityMinimalTest.R

@@ -4,7 +4,7 @@
 #assuming workingDir is where we are, cmd li{ne invocation is
 #Rscript --vanilla ./cloudDeploymentSandbox/Caribou-Demographic-Projection-Paper/Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R 1
 #Rscript --vanilla ./analysis/scripts/sensitivityMinimal.R 1 "local"
-#nohup Rscript --vanilla ./analysis/scripts/sensitivityMinimal.R 1 "local" &
+#nohup Rscript --vanilla ./analysis/scripts/sensitivityMinimalTest.R 1 "local" &
 
 args = commandArgs(trailingOnly=TRUE)
 #args=1
@@ -23,7 +23,7 @@ if(args[2]!="local"){
 }
 
 cpageId=args[1] #which batch?
-setName = "s3"
+setName = "s2"
 
 #######################
 setwd(workingDir)
@@ -58,8 +58,8 @@ setwd(toolDir)
 source("CaribouDemoFns.R")
 simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
 
-scResults = runScnSet(scns,eParsIn,simBig,getKSDists=F,printProgress=F)
+scResults = runScnSet(scns[1,],eParsIn,simBig,getKSDists=F,printProgress=F)
 setwd(workingDir)
 
-saveRDS(scResults,paste0("results/",setName,"/r",cpageId,".Rds"))
+saveRDS(scResults,paste0("results/",setName,"/rTest",cpageId,".Rds"))
 

### 2023-02-10 analysis/scripts/viewSensitivityResults.R

@@ -27,20 +27,27 @@ for(cpageId in pages){
   probs$AssessmentYear = probs$Year
 
   #See disturbance scenarios
-  distScns = subset(probs,is.element(projectionTime,c(0,5,20))|(Year==iYr))
+  #distScns = subset(probs,is.element(projectionTime,c(0,5,20))|(Year==iYr))
+  distScns = subset(probs,is.element(projectionTime,c(0,5,20))|(Year<2023))
   distScns$grp = paste0(distScns$Anthro2023,distScns$aSf)
-  distScns$Timeline[distScns$Year==2023]="Final monitoring year"
-  distScns$Timeline[distScns$projectionTime==5]="AssessmentYear 2028"
-  distScns$Timeline[distScns$projectionTime==20]="AssessmentYear 2043"
-  distScns$Timeline[distScns$Year<2023]= paste("Start",distScns$P[distScns$Year<2023],"yrs of monitoring")
+
+  distScns$Timeline[distScns$Year==2023]="Finish monitoring"
+  distScns$Timeline[distScns$projectionTime==5]="Assessment 2028"
+  distScns$Timeline[distScns$projectionTime==20]="Assessment 2043"
+  distScns$Timeline[distScns$Year<2023]= paste("Start",distScns$P[distScns$Year<2023],"yrs monitoring")
 
   startLevels = unique(distScns$Timeline[distScns$Year<2023])
-  distScns$Timeline = factor(distScns$Timeline,levels=c(startLevels,"Final monitoring year","AssessmentYear 2028","AssessmentYear 2043"))
+  #distScns$Timeline = factor(distScns$Timeline,levels=c(startLevels[length(startLevels):1],"Finish monitoring","Assessment 2028","Assessment 2043"))
+  distScns$Anthro2023=as.character(distScns$Anthro2023)
 
+  timelineLabs = unique(subset(distScns,(Year==iYr)|(Year>=2023),select=c(Year,Timeline,Anthro,Anthro2023,grp)))
+  timelineLabs = timelineLabs[order(timelineLabs$Year),]
+  distScns$Anthro=pmax(0,distScns$Anthro) #temp: remove this is next round of analysis. Problem fixed where it should be.
   png(here::here(paste0("figs/",setName,"/distScns",p,".png")),
       height = 4, width = 5.51, units = "in",res=600)
[…]

### 2023-02-16 analysis/scripts/sensitivityMinimalTest.R
fix lib

@@ -30,7 +30,7 @@ if(args[2]=="local"){
   print(workingDir)
   toolDir = file.path(baseDir,"BayesianCaribouDemographicProjection")
   print(toolDir)
-  libDir = .libPaths()
+  libDir = "R-packages"
   print(libDir)
 }
 

### 2023-02-16 make.R
remove caribouMEtrics from DESC and is install_gitHub instead

@@ -14,6 +14,7 @@ install.packages("devtools", dependencies = TRUE)
 
 devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never")
 
+devtools::install_github("LandSciTech/caribouMetrics")
 
 ## Load Project Addins (R Functions and Packages) ----
 

### 2023-02-16 analysis/scripts/sensitivityMinimalTest.R
pkg install issue again

@@ -32,6 +32,7 @@ if(args[2]=="local"){
   print(toolDir)
   libDir = "R-packages"
   print(libDir)
+  print(installed.packages()[,1:2])
 }
 
 cpageId=args[1] #which batch?

### 2023-02-17 analysis/scripts/sensitivityMinimalTest.R
more lib fix

@@ -31,7 +31,10 @@ if(args[2]=="local"){
   toolDir = file.path(baseDir,"BayesianCaribouDemographicProjection")
   print(toolDir)
   libDir = "R-packages"
+  cat("libDir")
   print(libDir)
+  cat(".lipPaths")
+  print(.libPaths())
   print(installed.packages()[,1:2])
 }
 

### 2023-02-17 make.R

@@ -12,9 +12,9 @@
 install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
 
-devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never")
+devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never", lib = "R-packages")
 
-devtools::install_github("LandSciTech/caribouMetrics")
+devtools::install_github("LandSciTech/caribouMetrics", lib = "R-packages")
 
 ## Load Project Addins (R Functions and Packages) ----
 

### 2023-02-17 make.R
fix lib again

@@ -12,9 +12,9 @@
 install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
 
-devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never", lib = "R-packages")
+devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never")
 
-devtools::install_github("LandSciTech/caribouMetrics", lib = "R-packages")
+devtools::install_github("LandSciTech/caribouMetrics")
 
 ## Load Project Addins (R Functions and Packages) ----
 

### 2023-02-17 make.R
add enable bspm

@@ -8,7 +8,7 @@
 #'
 #' @date 2023/02/16
 
-
+bspm::enable()
 install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
 

### 2023-02-17 make.R
try to use binary

@@ -8,7 +8,11 @@
 #'
 #' @date 2023/02/16
 
+# I copied these from the rocker/r-bspm Dockerfile because it sets them in the
+# RProfile but it wasn't working
 bspm::enable()
+options(pkgType="binary-source")
+
 install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
 

### 2023-02-17 make.R
fix options

@@ -11,7 +11,7 @@
 # I copied these from the rocker/r-bspm Dockerfile because it sets them in the
 # RProfile but it wasn't working
 bspm::enable()
-options(pkgType="binary-source")
+options(pkgType="binary", install.packages.check.source = "no")
 
 install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----

### 2023-02-17 analysis/scripts/sensitivityMinimalTest.R
use default library

@@ -30,7 +30,7 @@ if(args[2]=="local"){
   print(workingDir)
   toolDir = file.path(baseDir,"BayesianCaribouDemographicProjection")
   print(toolDir)
-  libDir = "R-packages"
+  libDir = NULL
   cat("libDir")
   print(libDir)
   cat(".lipPaths")

### 2023-02-17 make.R
fix cM install

@@ -18,7 +18,7 @@ install.packages("devtools", dependencies = TRUE)
 
 devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never")
 
-devtools::install_github("LandSciTech/caribouMetrics")
+devtools::install_github("LandSciTech/caribouMetrics", type = "source")
 
 ## Load Project Addins (R Functions and Packages) ----
 

### 2023-02-17 make.R
cM install again

@@ -13,11 +13,11 @@
 bspm::enable()
 options(pkgType="binary", install.packages.check.source = "no")
 
-install.packages("devtools", dependencies = TRUE)
+utils::install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
-
+print("install deps")
 devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never")
-
+warning("install caribouMetrics from GitHub")
 devtools::install_github("LandSciTech/caribouMetrics", type = "source")
 
 ## Load Project Addins (R Functions and Packages) ----

### 2023-02-17 make.R
same

@@ -17,7 +17,8 @@ utils::install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
 print("install deps")
 devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never")
-warning("install caribouMetrics from GitHub")
+message("install caribouMetrics from GitHub")
+options(pkgType = "both")
 devtools::install_github("LandSciTech/caribouMetrics", type = "source")
 
 ## Load Project Addins (R Functions and Packages) ----

### 2023-02-17 make.R
try pkgType both

@@ -11,7 +11,7 @@
 # I copied these from the rocker/r-bspm Dockerfile because it sets them in the
 # RProfile but it wasn't working
 bspm::enable()
-options(pkgType="binary", install.packages.check.source = "no")
+options(pkgType="both", install.packages.check.source = "no")
 
 utils::install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----

### 2023-02-17 make.R
back to binary, both seems to use source

@@ -11,12 +11,12 @@
 # I copied these from the rocker/r-bspm Dockerfile because it sets them in the
 # RProfile but it wasn't working
 bspm::enable()
-options(pkgType="both", install.packages.check.source = "no")
+options(pkgType="binary", install.packages.check.source = "no")
 
 utils::install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
 print("install deps")
-devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never")
+try(devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never"))
 message("install caribouMetrics from GitHub")
 options(pkgType = "both")
 devtools::install_github("LandSciTech/caribouMetrics", type = "source")

### 2023-02-17 make.R
instaling cM was not the issue

@@ -18,8 +18,7 @@ utils::install.packages("devtools", dependencies = TRUE)
 print("install deps")
 try(devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never"))
 message("install caribouMetrics from GitHub")
-options(pkgType = "both")
-devtools::install_github("LandSciTech/caribouMetrics", type = "source")
+devtools::install_github("LandSciTech/caribouMetrics")
 
 ## Load Project Addins (R Functions and Packages) ----
 

### 2023-02-17 make.R
type does need to be source for github but still want binaries for dependencies...

@@ -18,7 +18,7 @@ utils::install.packages("devtools", dependencies = TRUE)
 print("install deps")
 try(devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never"))
 message("install caribouMetrics from GitHub")
-devtools::install_github("LandSciTech/caribouMetrics")
+devtools::install_github("LandSciTech/caribouMetrics", type = "source")
 
 ## Load Project Addins (R Functions and Packages) ----
 

### 2023-02-17 make.R
remove bspm stuff bc wasn't working well now trying with rocker/verse which uses RSPM

@@ -8,17 +8,13 @@
 #'
 #' @date 2023/02/16
 
-# I copied these from the rocker/r-bspm Dockerfile because it sets them in the
-# RProfile but it wasn't working
-bspm::enable()
-options(pkgType="binary", install.packages.check.source = "no")
-
 utils::install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
 print("install deps")
 try(devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never"))
+
 message("install caribouMetrics from GitHub")
-devtools::install_github("LandSciTech/caribouMetrics", type = "source")
+devtools::install_github("LandSciTech/caribouMetrics")
 
 ## Load Project Addins (R Functions and Packages) ----
 

### 2023-02-17 make.R
install and enable rspm

@@ -8,6 +8,10 @@
 #'
 #' @date 2023/02/16
 
+utils::install.packages("rspm")
+
+rspm::enable()
+
 utils::install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
 print("install deps")

### 2023-02-17 make.R
go back to bspm but try doing cM deps separately

@@ -8,9 +8,10 @@
 #'
 #' @date 2023/02/16
 
-utils::install.packages("rspm")
-
-rspm::enable()
+# I copied these from the rocker/r-bspm Dockerfile because it sets them in the
+# RProfile but it wasn't working
+bspm::enable()
+options(pkgType="binary", install.packages.check.source = "no")
 
 utils::install.packages("devtools", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
@@ -18,7 +19,11 @@ print("install deps")
 try(devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never"))
 
 message("install caribouMetrics from GitHub")
-devtools::install_github("LandSciTech/caribouMetrics")
+try(devtools::install_github("LandSciTech/caribouMetrics",
+                             type = "source", dependencies = FALSE))
+
+# do this separately so they can be installed from binaries
+try(devtools::install_deps(file.path(.libPaths(), "caribouMetrics")))
 
 ## Load Project Addins (R Functions and Packages) ----
 

### 2023-02-17 make.R
install cM deps separately again

@@ -19,11 +19,17 @@ print("install deps")
 try(devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never"))
 
 message("install caribouMetrics from GitHub")
+# do this separately so they can be installed from binaries
+try(
+  remotes::install_deps(
+    remotes::remote_download(
+      remotes::github_remote("LandSciTech/caribouMetrics"))))
+
+
 try(devtools::install_github("LandSciTech/caribouMetrics",
                              type = "source", dependencies = FALSE))
 
-# do this separately so they can be installed from binaries
-try(devtools::install_deps(file.path(.libPaths(), "caribouMetrics")))
+
 
 ## Load Project Addins (R Functions and Packages) ----
 

### 2023-02-17 make.R
also install deps from source

@@ -13,27 +13,27 @@
 bspm::enable()
 options(pkgType="binary", install.packages.check.source = "no")
 
-utils::install.packages("devtools", dependencies = TRUE)
+utils::install.packages("remotes", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
 print("install deps")
-try(devtools::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never"))
+try(remotes::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never"))
 
 message("install caribouMetrics from GitHub")
 # do this separately so they can be installed from binaries
-try(
-  remotes::install_deps(
-    remotes::remote_download(
-      remotes::github_remote("LandSciTech/caribouMetrics"))))
+cm_download <- remotes::remote_download(remotes::github_remote("LandSciTech/caribouMetrics"))
 
+try(remotes::install_deps(cm_download))
+# this should do ones that were not available from binary
+try(remotes::install_deps(cm_download), type = "source")
 
-try(devtools::install_github("LandSciTech/caribouMetrics",
+try(remotes::install_github("LandSciTech/caribouMetrics",
                              type = "source", dependencies = FALSE))
 
 
 
 ## Load Project Addins (R Functions and Packages) ----
 
-devtools::load_all(here::here())
+# devtools::load_all(here::here())
 
 
 ## Global Variables ----

### 2023-02-17 make.R
more install fix

@@ -17,17 +17,20 @@ utils::install.packages("remotes", dependencies = TRUE)
 ## Install Dependencies (listed in DESCRIPTION) ----
 print("install deps")
 try(remotes::install_deps("Caribou-Demographic-Projection-Paper", upgrade = "never"))
+try(remotes::install_deps("Caribou-Demographic-Projection-Paper", type = "source",
+                          upgrade = "never"))
 
 message("install caribouMetrics from GitHub")
-# do this separately so they can be installed from binaries
+
+# do deps separately so they can be installed from binaries
 cm_download <- remotes::remote_download(remotes::github_remote("LandSciTech/caribouMetrics"))
 
-try(remotes::install_deps(cm_download))
+try(remotes::install_deps(cm_download), upgrade = "never")
 # this should do ones that were not available from binary
-try(remotes::install_deps(cm_download), type = "source")
+try(remotes::install_deps(cm_download), type = "source", upgrade = "never")
 
 try(remotes::install_github("LandSciTech/caribouMetrics",
-                             type = "source", dependencies = FALSE))
+                             type = "source", dependencies = FALSE, upgrade = "never"))
 
 
 

### 2023-02-17 make.R
fix above

@@ -25,9 +25,9 @@ message("install caribouMetrics from GitHub")
 # do deps separately so they can be installed from binaries
 cm_download <- remotes::remote_download(remotes::github_remote("LandSciTech/caribouMetrics"))
 
-try(remotes::install_deps(cm_download), upgrade = "never")
+try(remotes::install_deps(cm_download, upgrade = "never"))
 # this should do ones that were not available from binary
-try(remotes::install_deps(cm_download), type = "source", upgrade = "never")
+try(remotes::install_deps(cm_download, type = "source", upgrade = "never"))
 
 try(remotes::install_github("LandSciTech/caribouMetrics",
                              type = "source", dependencies = FALSE, upgrade = "never"))

### 2023-02-17 analysis/paper/paper.Rmd
Set up cloud batch.

--- 2023-01-20 analysis/paper/paper.Rmd
@@
 ---
-title: "Caribou Demographic Projection - interesting & informative title TBD"
+title: "National demographic-disturbance relationships and local monitoring both reduce uncertainty in projections of boreal caribou population viability: tools to support monitoring decions"
 author:
@@
 <!-- The actual document text starts here: -->
-
+# Abstract
+Across fifty-eight boreal caribou ranges in Canada, both survival and recruitment decrease with the percentage of the range that is disturbed.
+However, there is also variation in demographic rates among ranges, particularly where anthropogenic disturbance is low, and this variation leads to uncertainty in projections derived from national disturbance-demographic relationships.
+Demographic projections derived from local range-specific data are also uncertain because few places have enough long-term monitoring data to characterize local demographic-disturbance relationships.
+We present a simple Bayesian integrated population model that integrates prior information from national analysis of demographic-disturbance relationships with available local demographic data to reduce uncertainty in population viability projections.
+We combine this model with methods for simulating local population dynamics and monitoring to show how monitoring requirements depend on landscape condition.
+Where anthropogenic disturbance is very high or very low, reasonably accurate status projections can be made using only national demographic-disturbance relationships.
+At intermediate disturbance levels where uncertainty is higher, deploying few collars for more years may be a lower cost option for improving the reliability of status projections relative to more intensive short-term monitoring.
+We provide an R Shiny interface to enable others to explore the implications of alternative monitoring scenarios, disturbance trajectories, priors, and other model parameters for boreal caribou demographic projections.
+This tool can help users assess monitoring needs and strategies.
 # Introduction
@@
 
+```
+
+```{r fig2, fig.cap="Example projections, assuming a target of 30 collars, renewed every 4 years, with 3 cows per collared cow included in aerial surveys, and a 1% increase in anthropogenic disturbance per year (as in Figure 1).
+The number of monitoring years varies from P=0 (top row) to P=24 (bottom row), and the anthropogenic disturbance footprint in 2023 varies from 0% (left column) to 60% (right column).
+A status assessment error occurs when the projected population growth rate is above one and the true rate is below one or vice versa.
+When disturbance is very high (e.g.
+Anthro2023 = 60) populations that differ substantially from one another (compare 'true' lines for high and low quantile populations) share the same status, and a projection informed only by national demographic disturbance relationships (P=0) is accurate for both.
+When disturbance is intermediate, populations that differ substantially from one another are more likely to differ in their status.
+In these cases, integrating local monitoring data improves projection accuracy (i.e.
+smaller distances between projected and true lines) and reduces status assessment errors.", cache=TRUE}
+knitr::include_graphics(here::here("figs/forPaper/examples.png"))
+```
+
+```{r fig3, fig.cap="Power analysis example result.", cache=TRUE}
+knitr::include_graphics(here::here("figs/forPaper/powercmult3ay3aSf1.png"))
+knitr::include_graphics(here::here("figs/forPaper/powerEffortcmult3ay3aSf1.png"))
+
+#TO DO: figure out how to put no monitoring results on these graphs.
 ```

### 2023-02-17 analysis/scripts/sensitivity1.R

@@ -16,6 +16,8 @@
 #monitoringScns = expand.grid(P=c(1,4,8,16),st=c(30,60),cmult=c(3),ri=c(1,2,4),assessmentYrs=1)
 #stateScns = expand.grid(iA=c(0),rep=seq(1:30))
 
+#TO DO: adjust method so all
+
 setName = "s3"
 monitoringScns = expand.grid(P=c(1,2,4,8,16,24),st=c(1,15,30,60),cmult=c(3,6,9),ri=c(1,4),assessmentYrs=c(3))
 stateScns = data.frame(tA=c(0,20,40,60,0,20,40,60),aS=c(0,1,1,1,0,0,0,0),aSf=c(1,1,1,1,0,0,0,0))

### 2023-02-17 analysis/scripts/sensitivityMinimal.R

@@ -50,7 +50,7 @@ eParsIn$freqStartsByYear <- data.frame(Year = 1981:2023,
                                        numStarts = 30)
 eParsIn$collarOnTime=1
 eParsIn$collarOffTime=12
-eParsIn$collarNumYears=3
+eParsIn$collarNumYears=6
 
 
 setwd(toolDir)

### 2023-02-17 analysis/scripts/viewSensitivityResults.R

@@ -21,6 +21,37 @@ for(cpageId in pages){
   if(!file.exists(paste0("results/",setName,"/r",cpageId,".Rds"))){next}
   scResults = readRDS(paste0("results/",setName,"/r",cpageId,".Rds"))
 
+  #show examples projections
+  exResults = subset(scResults$rr.summary.all,(st==30)&(ri==4)&(Parameter=="Population growth rate"))
+  exResults$meanQ = (exResults$rQ+exResults$sQ)/2
+  grpID = subset(exResults,select=c(tA,P,meanQ)) %>% group_by(tA,P) %>%
+    summarise(minQ = min(meanQ),maxQ=max(meanQ))
+
+  exResults=merge(exResults,grpID)
+  exResults$quantile[exResults$meanQ == exResults$maxQ]="high"
+  exResults$quantile[exResults$meanQ == exResults$minQ]="low"
+  exResults = subset(exResults,!is.na(quantile))
+  exResults$type[exResults$Year<=2023]="estimated"
+  exResults$type[exResults$Year>2023]="projected"
+  exResults$grp = paste(exResults$type,exResults$quantile, exResults$tA,exResults$P)
+  exResults$Anthro2023 = exResults$tA
+
+  obs = subset(scResults$obs.all,(st==30)&(ri==4)&(parameter=="Population growth rate"))
+  obs = merge(obs,unique(subset(exResults,select=c(tA,P,rQ,sQ,quantile,grp,Anthro2023))))
+  obs$type = "true"
+
+  png(here::here(paste0("figs/",setName,"/examples.png")),
+      height = 6, width = 5.56, units = "in",res=600)
+  base=ggplot(exResults,aes(x=Year,y=Mean,col=quantile,group=grp,linetype=type))+geom_line(show.legend=T)+
+    geom_ribbon(aes(ymin = `Lower 95% CRI`, ymax = `Upper 95% CRI`,fill=quantile),
+                show.legend = FALSE, alpha = 0.25,col=NA)+
+    geom_line(data=obs,aes(x=Year,y=Mean, col=quantile,group=grp,linetype=type),show.legend=T)+
+    theme_bw()+facet_grid(P~Anthro2023,labeller = "label_both")+ylab("Population growth rate")+
+    theme(axis.text.x = element_text(angle=90,vjust=0.5,hjust=1,size=8))+
+    geom_hline(yintercept=1, color = "black",size=0.7)
+  print(base)
+  dev.off()
+
[…]

### 2023-02-17 analysis/scripts/sensitivityMinimal.R
copy dirs for cloud to sensMinimal

@@ -11,9 +11,16 @@ args = commandArgs(trailingOnly=TRUE)
 if(args[2]!="local"){
   baseDir = getwd()
   print(baseDir)
-  workingDir = paste0(baseDir,"/cloudDeploymentSandbox/Caribou-Demographic-Projection-Paper/Caribou-Demographic-Projection-Paper")
-  toolDir = paste0(baseDir,"/cloudDeploymentSandbox/BayesianCaribouDemographicProjection")
-  libDir = paste0(baseDir,"/cloudDeploymentSandbox/Rpackages")
+  workingDir = file.path(baseDir,"Caribou-Demographic-Projection-Paper")
+  print(workingDir)
+  toolDir = file.path(baseDir,"BayesianCaribouDemographicProjection")
+  print(toolDir)
+  libDir = NULL
+  cat("libDir")
+  print(libDir)
+  cat(".lipPaths")
+  print(.libPaths())
+  print(installed.packages()[,1:2])
 
 }else{
   baseDir = "C:/Users/HughesJo/Documents"

### 2023-02-17 make.R
comments

@@ -48,3 +48,9 @@ try(remotes::install_github("LandSciTech/caribouMetrics",
 
 # List all R scripts in a sequential order and using the following form:
 # source(here::here("analyses", "script_X.R"))
+
+# code to make commands for all batches to use in bash script
+# cat(paste0('nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" ',
+#        1:24,
+#        ' "cloud"', sep = "\n"))
+

### 2023-02-17 cloud/run_caribou.sh
add cloud files I have removed PAT and SAS tokens. There is probably a better way to do this...

@@ -1,0 +1,43 @@
+#!/bin/bash
+#############################################
+#####installing git and jags onto node##################
+apt-get update
+apt-get install -y --no-install-recommends git-all
+apt-get install -y --no-install-recommends jags
+echo "Finished installing Git, jags, and updating Node"
+
+#############################################
+#####clone repo to get code needed##################
+git clone https://[PAT]@github.com/LandSciTech/Caribou-Demographic-Projection-Paper.git
+git clone https://[PAT]@github.com/LandSciTech/BayesianCaribouDemographicProjection.git
+
+mkdir R-packages
+export R_LIBS=R-packages
+#############################################
+#####Run R script##################
+echo "Running the scripts"
+Rscript "Caribou-Demographic-Projection-Paper/make.R"
+nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" 1 "cloud"
+nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" 2 "cloud"
+nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" 3 "cloud"
+nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" 4 "cloud"
+nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" 5 "cloud"
+nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" 6 "cloud"
+nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" 7 "cloud"
+nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" 8 "cloud"
+nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" 9 "cloud"
+nohup Rscript "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" 10 "cloud"
[…]

### 2023-04-25 R/CaribouDemographicFunctions.R
move makeInterceptPlots from app repo to here

@@ -55,3 +55,111 @@ runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMei
   setwd(retdir)
   return(scResults)
 }
+
+makeInterceptPlots <- function(scResults, addBit = "", facetVars = c("P", "sQ"),
+                               loopVars = NULL,
+                               whichPlots = c("Adult female survival",
+                                              "Population growth rate",
+                                              "Recruitment",
+                                              "Female population size"),
+                               survLow = 0.6, type = "png", useNational = T) {
+  # facetVars=c("lre","sre");loopVars="srv";scResults=scResultsHigh
+
+  if (!is.null(loopVars)) {
+    loopSet <- unique(subset(scResults$rr.summary.all, select = loopVars))
+    loopSet$dummy <- 1
+  } else {
+    loopSet <- data.frame(dummy = 1)
+  }
+
+
+  for (l in 1:nrow(loopSet)) {
+    # l = 1
+    crow <- loopSet[l, ]
+
+    aa <- ""
+    for (n in names(crow)) {
+      if (n == "dummy") {
+        next
+      }
+      aa <- paste0(aa, n, crow[[n]])
+    }
+
+    addBitO <- paste0(addBit, aa)
+
+    if (useNational) {
+      simRange <- merge(scResults$sim.all, crow)
+    } else {
+      simRange <- NULL
+    }
+
+    if (is.element("Adult female survival", whichPlots)) {
+      if (type == "png") {
+        png(here::here(paste0("figs/Surv", addBitO, ".png")),
+            height = 6, width = 7.48, units = "in", res = 600
+        )
+      } else {
+        pdf(paste0("figs/Surv", addBitO, ".pdf"), width = 10, height = 7)
+      }
+      print(plotRes(merge(scResults$rr.summary.all, crow), "Adult female survival",
+                    obs = merge(scResults$obs.all, crow),
+                    lowBound = survLow, simRange = simRange, facetVars = facetVars
+      ))
+      dev.off()
+    }
+
+    if (is.element("Population growth rate", whichPlots)) {
+      if (type == "png") {
+        png(here::here(paste0("figs/Lambda", addBitO, ".png")),
+            height = 6, width = 7.48, units = "in", res = 600
+        )
+      } else {
+        pdf(paste0("figs/Lambda", addBitO, ".pdf"), width = 10, height = 7)
+      }
[…]

### 2023-04-25 analysis/scripts/sensitivityMinimalTestPar.R
Make minimal test work with updated package and using future for parellel

@@ -1,38 +1,12 @@
 #!/usr/bin/env Rscript
 
-# Run batches from Rscript that uses parallel backend
-
-args = commandArgs(trailingOnly=TRUE)
-#args=1
-if(args[2]=="local"){
-  baseDir = "C:/Users/HughesJo/Documents"
-  workingDir = paste0(baseDir,"/gitprojects/Caribou-Demographic-Projection-Paper")
-  toolDir = paste0(baseDir,"/gitprojects/BayesianCaribouDemographicProjection")
-  libDir = paste0(baseDir,"")
-}else if (args[2]=="local_SE"){
-  baseDir = "C:/Users/EndicottS/Documents"
-  workingDir = paste0(baseDir,"/gitprojects/Caribou-Demographic-Projection-Paper")
-  toolDir = paste0(baseDir,"/gitprojects/BayesianCaribouDemographicProjection")
-  libDir = paste0("C:/Users/endicotts/AppData/Local/Programs/R/R-4.2.1/library")
-}else if (args[2]=="docker"){
-  baseDir = "/"
-  workingDir = paste0(baseDir,"Caribou-Demographic-Projection-Paper")
-  toolDir = paste0(baseDir,"BayesianCaribouDemographicProjection")
-  libDir = NULL
-}else{
-  baseDir = getwd()
-  workingDir = file.path(baseDir,"Caribou-Demographic-Projection-Paper")
-  toolDir = file.path(baseDir,"BayesianCaribouDemographicProjection")
-  libDir = NULL
-}
-
+# Run batches from Rscript that uses parallel backend and new caribouMetrics functions
+n_batches <- 6
+n_scns <- 2
 
 setName = "s2"
 
-
-
 #######################
-setwd(workingDir)
 dir.create(paste0("figs/",setName),recursive=T)
 dir.create(paste0("tabs/",setName),recursive=T)
 dir.create(paste0("results/",setName),recursive=T)
@@ -41,14 +15,11 @@ packages <- c("R2jags","gdata","mcmcplots",
               "ggplot2","RODBC","plyr","dplyr","survival","gdata","data.table",
               "tidyr","caribouMetrics")
 
-for(p in packages){library(p,lib.loc=libDir,character.only=T)}
+for(p in packages){library(p,character.only=T)}
 
-setwd(toolDir)
-
-source("CaribouDemoFns.R")
[…]

### 2023-04-25 R/CaribouDemographicFunctions.R
Make paper and scripts called by it work with updated package

@@ -9,27 +9,16 @@
 runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMeier",getKSDists=T){
   #quants=NULL;Anthro=NULL;survAnalysisMethod="KaplanMeier";getKSDists=F
   retdir <-getwd()
-  wdir <- app_path
 
   dir.create("figs")
   dir.create("tabs")
   dir.create("results")
 
-  source(paste0(wdir,"/CaribouDemoFns.R"))
-
-  scns<-fillDefaults(scns)
-
-  # usage
-  packages <- c("shiny", "R2jags","gdata","mcmcplots",
-                "ggplot2", "rmarkdown", "knitr",
-                "RODBC","plyr","survival","gdata","data.table")
-
-  ipak(packages)
-  load.module("glm")
-
   library(caribouMetrics)
   library(tidyr)
   library(dplyr)
+
+  scns<-fillDefaults(scns)
 
   eParsIn = list()
   eParsIn$cowCounts <- data.frame(Year = 1981:2023,
@@ -45,14 +34,12 @@ runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMei
   #Get full set of sims for comparison
 
   if(is.null(quants)){
-    simBig<-getSimsNational(wdir=wdir,adjustR=unique(scns$adjustR))#If called with default parameters, use saved object to speed things up.
+    simBig<-getSimsNational(adjustR=unique(scns$adjustR))#If called with default parameters, use saved object to speed things up.
   }else{
-    simBig<-getSimsNational(quants=quants,Anthro=Anthro,adjustR=unique(scns$adjustR))#If called with default parameters, use saved object to speed things up.
+    simBig<-getSimsNational(useQuantiles = quants,Anthro=Anthro,adjustR=unique(scns$adjustR))#If called with default parameters, use saved object to speed things up.
   }
 
-  setwd(wdir)
-  scResults = runScnSet(scns,eParsIn,simBig,survAnalysisMethod,getKSDists=getKSDists,printProgress=T)
-  setwd(retdir)
+  scResults = caribouMetrics:::runScnSet(scns,eParsIn,simBig,survAnalysisMethod,getKSDists=getKSDists,printProgress=T)
   return(scResults)
 }
 
@@ -101,7 +88,7 @@ makeInterceptPlots <- function(scResults, addBit = "", facetVars = c("P", "sQ"),
[…]

### 2023-04-25 analysis/scripts/calibratePriors.R

@@ -1,4 +1,3 @@
-#app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
 #devtools::load_all(here::here())
 
 ########################

### 2023-04-25 analysis/scripts/formatPriors.R

@@ -1,27 +1,13 @@
 
-#app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
 library(caribouMetrics)
-source(paste0(app_path,"/CaribouDemoFns.R"))
 
 rawPriors<-getPriors(returnValues=F)
 
-#tbl <- data.frame()
-"Parameter|Description|Mean|Standard Deviation
-"
-
 nameSet <- gsub(".Prior1","",names(rawPriors)[grepl("Prior1",names(rawPriors))],fixed=T)
 
-start=T
-for(n in nameSet){
-  crow = data.frame(name=n,Mean=rawPriors[[paste0(n,".Prior1")]],SD=rawPriors[[paste0(n,".Prior2")]])
-
-  if(start){
-    tbl = crow
-    start=F
-  }else{
-    tbl=rbind(tbl,crow)
-  }
-}
+tbl <- data.frame(name = nameSet,
+                  Mean = unlist(rawPriors[paste0(nameSet,".Prior1")], use.names = FALSE),
+                  SD = unlist(rawPriors[paste0(nameSet,".Prior2")], use.names = FALSE))
 
 pars <- data.frame(name="l.R",Parameter="$\\beta^R_0$",Description="R intercept")
 pars<-rbind(pars,data.frame(name="beta.Rec.anthro",Parameter="$\\beta^R_a$",Description="R anthro slope"))

### 2023-04-25 analysis/scripts/formatScenarios.R

@@ -1,5 +1,4 @@
 
-#app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
 library(caribouMetrics)
 
 rawScns<-read.csv(here::here("tabs/s2.csv"))

### 2023-04-25 analysis/scripts/viewSensitivityResults.R

@@ -1,10 +1,8 @@
-#app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
 #devtools::load_all(here::here())
 library(tidyverse)
 library(ggplot2)
-source( paste0(app_path,"/CaribouDemoFns.R"))
-scn_defaults <- c(eval(formals(fillDefaults)$defList),
-                  curYear = eval(formals(fillDefaults)$curYear))
+
+scn_defaults <- eval(formals(fillDefaults))
 
 ########################
 #sensitivity

### 2023-04-25 analysis/scripts/viewSensitivityResults.R
switching to dev branch

@@ -8,18 +8,58 @@ scn_defaults <- c(eval(formals(fillDefaults)$defList),
 
 ########################
 #sensitivity
-setName = "s2"
-scns = read.csv(here::here("tabs/s2.csv"))
+setName = "s3"
+scns = read.csv(here::here("tabs/s3.csv"))
 scns = subset(scns,cmult!=0)
-pages=unique(scns$pageId)
-unique(scns$pageLab)
+pagesa=unique(scns$pageId)
 
-for(cpageId in pages){
-  #cpageId=pages[1]
-  p = scns$pageLab[scns$pageId==cpageId]
+pages = pagesa
+for (p in pagesa){
+  if(!file.exists(paste0("results/",setName,"/r",p,".Rds"))){pages=pages[pages!=p]}
+}
 
-  if(!file.exists(paste0("results/",setName,"/r",cpageId,".Rds"))){next}
-  scResults = readRDS(paste0("results/",setName,"/r",cpageId,".Rds"))
+batchStrip<-function(l,batches=seq(1:4)){
+  for(b in batches){
+    l=gsub(paste0("repBatch",b),"",l,fixed=T)
+  }
+  return(l)
+}
+
+pages=sort(pages)
+
+combine=T
+for(i in 1:length(pages)){
+  #combine=F;i=5
+  cpageId=pages[i]
+
+  if(i==length(pages)){
+    nextP = "Batch1"
+  }else{
+    nextP = unique(scns$pageLab[scns$pageId==pages[i+1]])
+  }
+
+  p = unique(scns$pageLab[scns$pageId==cpageId])
+
+  if(combine&(cpageId>1)){
+    scNew = readRDS(paste0("results/",setName,"/r",cpageId,".Rds"))
+    for(n in names(scNew)){
+      if(n=="errorLog"){
+        scResults[[n]]=c(scResults[[n]],scNew[[n]])
+      }else{
+        scResults[[n]]=rbind(scResults[[n]],scNew[[n]])
+      }
+    }
+  }else{
+    scResults = readRDS(paste0("results/",setName,"/r",cpageId,".Rds"))
+  }
+
+  if(as.numeric(strsplit(nextP,"repBatch")[[1]][2])<=as.numeric(strsplit(p,"repBatch")[[1]][2])){
+    combine=F
+  }else{combine=T;next}
+
+  print(paste(i,p))
+
+  #figure out how to count out errors.
 
   #show examples projections
[…]

### 2023-04-25 analysis/scripts/sensitivityMinimal.R
update sensitivityMinimal

@@ -6,39 +6,20 @@
 #Rscript --vanilla ./analysis/scripts/sensitivityMinimal.R 1 "local"
 #nohup Rscript --vanilla ./analysis/scripts/sensitivityMinimal.R 1 "local" &
 
-args = commandArgs(trailingOnly=TRUE)
-#args=1
-if(args[2]!="local"){
-  baseDir = getwd()
-  workingDir = file.path(baseDir,"Caribou-Demographic-Projection-Paper")
-  toolDir = file.path(baseDir,"BayesianCaribouDemographicProjection")
-  libDir = NULL
-  print(workingDir)
-}else{
-  baseDir = "C:/Users/HughesJo/Documents"
-  workingDir = paste0(baseDir,"/gitprojects/Caribou-Demographic-Projection-Paper")
-  toolDir = paste0(baseDir,"/gitprojects/BayesianCaribouDemographicProjection")
-  libDir = paste0(baseDir,"/cloudDeploymentSandbox/Rpackages")
-}
+# Run batches from Rscript that uses parallel backend and new caribouMetrics functions
+library(doFuture)
+library(caribouMetrics)
 
-cpageId=args[1] #which batch?
 setName = "s3"
-message("batch ", args[1], " started")
+
 #######################
-setwd(workingDir)
 dir.create(paste0("figs/",setName),recursive=T)
 dir.create(paste0("tabs/",setName),recursive=T)
 dir.create(paste0("results/",setName),recursive=T)
 
-packages <- c("R2jags","gdata","mcmcplots",
-              "ggplot2","RODBC","plyr","dplyr","survival","gdata","data.table",
-              "tidyr","caribouMetrics")
-
-for(p in packages){library(p,lib.loc=libDir,character.only=T)}
+simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
 
 allScns = read.csv(paste0("tabs/",setName,".csv"))
-scns = subset(allScns,pageId==cpageId)
-nrow(allScns)
 
 ####################
 eParsIn = list()
@@ -51,15 +32,19 @@ eParsIn$collarOnTime=1
 eParsIn$collarOffTime=12
 eParsIn$collarNumYears=6
 
+future::plan("multisession")
 
-setwd(toolDir)
+foreach(cpageId = 1:n_batches, .options.future = list(seed = TRUE)) %dofuture% {
 
-source("CaribouDemoFns.R")
[…]

### 2023-04-25 analysis/scripts/sensitivityMinimalTestPar.R

@@ -1,55 +1,48 @@
 #!/usr/bin/env Rscript
 
+
 # Run batches from Rscript that uses parallel backend and new caribouMetrics functions
-n_batches <- 6
+n_batches <- 1
 n_scns <- 2
 
-setName = "s2"
+library(doFuture)
+library(caribouMetrics)
+
+setName = "s3"
 
 #######################
 dir.create(paste0("figs/",setName),recursive=T)
 dir.create(paste0("tabs/",setName),recursive=T)
 dir.create(paste0("results/",setName),recursive=T)
 
-packages <- c("R2jags","gdata","mcmcplots",
-              "ggplot2","RODBC","plyr","dplyr","survival","gdata","data.table",
-              "tidyr","caribouMetrics")
-
-for(p in packages){library(p,character.only=T)}
-
 simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
-
-
 
 allScns = read.csv(paste0("tabs/",setName,".csv"))
 
-library(doFuture)
+####################
+eParsIn = list()
+eParsIn$cowCounts <- data.frame(Year = 1981:2023,
+                                Count = 100,
+                                Class = "cow")
+eParsIn$freqStartsByYear <- data.frame(Year = 1981:2023,
+                                       numStarts = 30)
+eParsIn$collarOnTime=1
+eParsIn$collarOffTime=12
+eParsIn$collarNumYears=3
 
 future::plan("multisession")
 
-foreach(cpageId = 1:n_batches) %dofuture% {
+foreach(cpageId = 1:n_batches, .options.future = list(seed = TRUE)) %dofuture% {
 
-  scns = subset(allScns,pageId==cpageId)
-  nrow(allScns)
+  scns = subset(allScns, pageId==cpageId)
 
   message("batch ", cpageId, " started")
 
-  ####################
-  eParsIn = list()
-  eParsIn$cowCounts <- data.frame(Year = 1981:2023,
-                                  Count = 100,
-                                  Class = "cow")
-  eParsIn$freqStartsByYear <- data.frame(Year = 1981:2023,
-                                         numStarts = 30)
-  eParsIn$collarOnTime=1
-  eParsIn$collarOffTime=12
-  eParsIn$collarNumYears=3
-
   scResults = caribouMetrics:::runScnSet(scns[1:n_scns,],eParsIn,simBig,getKSDists=F,printProgress=F)
-
 
[…]

### 2023-04-25 make.R
use branch in cloud

@@ -29,7 +29,7 @@ try(remotes::install_deps(cm_download, upgrade = "never"))
 # this should do ones that were not available from binary
 try(remotes::install_deps(cm_download, type = "source", upgrade = "never"))
 
-try(remotes::install_github("LandSciTech/caribouMetrics",
+try(remotes::install_github("LandSciTech/caribouMetrics", ref = "add-bayes-demog",
                              type = "source", dependencies = FALSE, upgrade = "never"))
 
 

### 2023-04-26 cloud/run_caribou.sh
update docker and cloud files for new version of caribouMetrics and parallel processing with future

@@ -1,29 +1,10 @@
 #!/bin/bash
 #############################################
-#####installing git and jags onto node##################
-apt-get update
-apt-get install -y --no-install-recommends git-all
-apt-get install -y --no-install-recommends jags
-echo "Finished installing Git, jags, and updating Node"
-
-#############################################
-#####clone repo to get code needed##################
-git clone https://<PAT>@github.com/LandSciTech/Caribou-Demographic-Projection-Paper.git
-git clone https://<PAT>@github.com/LandSciTech/BayesianCaribouDemographicProjection.git
-
-mkdir R-packages
-export R_LIBS=R-packages
-
-mv "/usr/local/lib/R/site-library/bspm" "/usr/lib/R/library"
-
-#############################################
 #####Run R script##################
+echo "updating repo code"
+cd /Caribou-Demographic-Projection-Paper
+git pull https://<GITHUBPAT>@github.com/LandSciTech/Caribou-Demographic-Projection-Paper.git
 
 echo "Running the scripts"
-nohup Rscript --vanilla "Caribou-Demographic-Projection-Paper/make.R"
-for i in 15 16 18 19 20 21 24
-do
-	echo "Running batch" $i
-	nohup Rscript --vanilla "Caribou-Demographic-Projection-Paper/analysis/scripts/sensitivityMinimal.R" $i "cloud" > nohup_$i.out 2>&1 &
-done
-wait
+nohup Rscript --vanilla "make.R"
+nohup Rscript --vanilla "analysis/scripts/sensitivityMinimalTestPar.R" > $AZ_BATCH_TASK_WORKING_DIR/nohup.out 2>&1

### 2023-05-08 analysis/scripts/formatPriors.R
undo some changes

@@ -1,13 +1,26 @@
 
+#app_path <- "C:/Users/HughesJo/Documents/gitprojects/BayesianCaribouDemographicProjection"
 library(caribouMetrics)
 
 rawPriors<-getPriors(returnValues=F)
 
+#tbl <- data.frame()
+"Parameter|Description|Mean|Standard Deviation
+"
+
 nameSet <- gsub(".Prior1","",names(rawPriors)[grepl("Prior1",names(rawPriors))],fixed=T)
 
-tbl <- data.frame(name = nameSet,
-                  Mean = unlist(rawPriors[paste0(nameSet,".Prior1")], use.names = FALSE),
-                  SD = unlist(rawPriors[paste0(nameSet,".Prior2")], use.names = FALSE))
+start=T
+for(n in nameSet){
+  crow = data.frame(name=n,Mean=rawPriors[[paste0(n,".Prior1")]],SD=rawPriors[[paste0(n,".Prior2")]])
+
+  if(start){
+    tbl = crow
+    start=F
+  }else{
+    tbl=rbind(tbl,crow)
+  }
+}
 
 pars <- data.frame(name="l.R",Parameter="$\\beta^R_0$",Description="R intercept")
 pars<-rbind(pars,data.frame(name="beta.Rec.anthro",Parameter="$\\beta^R_a$",Description="R anthro slope"))

### 2023-05-08 analysis/scripts/sensitivityMinimalTestPar.R
reduce globals passed to each worker and try to reduce memory a bit

@@ -2,8 +2,8 @@
 
 
 # Run batches from Rscript that uses parallel backend and new caribouMetrics functions
-n_batches <- 10
-n_scns <- 2
+n_batches <- 24
+n_scns <- 4
 
 library(doFuture)
 library(caribouMetrics)
@@ -15,26 +15,28 @@ dir.create(paste0("figs/",setName),recursive=T)
 dir.create(paste0("tabs/",setName),recursive=T)
 dir.create(paste0("results/",setName),recursive=T)
 
-simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
-
-allScns = read.csv(paste0("tabs/",setName,".csv"))
-
-####################
-eParsIn = list()
-eParsIn$cowCounts <- data.frame(Year = 1981:2023,
-                                Count = 100,
-                                Class = "cow")
-eParsIn$freqStartsByYear <- data.frame(Year = 1981:2023,
-                                       numStarts = 30)
-eParsIn$collarOnTime=1
-eParsIn$collarOffTime=12
-eParsIn$collarNumYears=3
-
 future::plan("multisession")
 
 foreach(cpageId = 1:n_batches, .options.future = list(seed = TRUE)) %dofuture% {
 
+  simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
+
+  allScns = read.csv(paste0("tabs/",setName,".csv"))
+
+  ####################
+  eParsIn = list()
+  eParsIn$cowCounts <- data.frame(Year = 1981:2023,
+                                  Count = 100,
+                                  Class = "cow")
+  eParsIn$freqStartsByYear <- data.frame(Year = 1981:2023,
+                                         numStarts = 30)
+  eParsIn$collarOnTime=1
+  eParsIn$collarOffTime=12
+  eParsIn$collarNumYears=3
+
   scns = subset(allScns, pageId==cpageId)
+
+  rm(allScns)
 
   message("batch ", cpageId, " started")
 
@@ -46,3 +48,7 @@ foreach(cpageId = 1:n_batches, .options.future = list(seed = TRUE)) %dofuture% {
 
 }
 future::plan("sequential")
+
+# mem allocated to run 24 scenarios 35.9 GB not necessarily used simultaneously
+# while running 24 scenarios memory usage was up to 3.6 GB at one point
+# D16s_v3 VM has 16 cores and twice as much memory 4 GB per core so might work better.

### 2023-05-08 analysis/paper/paper.Rmd
Make paper and scripts called by it work with updated package

@@
 ---
-title: "National demographic-disturbance relationships and local monitoring both reduce uncertainty in projections of boreal caribou population viability: tools to support monitoring decions"
+title: "National demographic-disturbance relationships and local monitoring both reduce uncertainty in projections of boreal caribou population viability: tools to support monitoring decisions"
 author:
@@
 abstract: |
-  Text of abstract
+  Across fifty-eight boreal caribou ranges in Canada, both survival and recruitment decrease with the percentage of the range that is disturbed.
+However, there is also variation in demographic rates among ranges, particularly where anthropogenic disturbance is low, and this variation leads to uncertainty in projections derived from national disturbance-demographic relationships.
+Demographic projections derived from local range-specific data are also uncertain because few places have enough long-term monitoring data to characterize local demographic-disturbance relationships.
+We present a simple Bayesian integrated population model that integrates prior information from national analysis of demographic-disturbance relationships with available local demographic data to reduce uncertainty in population viability projections.
+We combine this model with methods for simulating local population dynamics and monitoring to show how monitoring requirements depend on landscape condition.
+Where anthropogenic disturbance is very high or very low, reasonably accurate status projections can be made using only national demographic-disturbance relationships.
+At intermediate disturbance levels where uncertainty is higher, deploying few collars for more years may be a lower cost option for improving the reliability of status projections relative to more intensive short-term monitoring.
+We provide an R Shiny interface to enable others to explore the implications of alternative monitoring scenarios, disturbance trajectories, priors, and other model parameters for boreal caribou demographic projections.
+This tool can help users assess monitoring needs and strategies.
+
 keywords: |
@@
 <!-- The actual document text starts here: -->
-# Abstract
-Across fifty-eight boreal caribou ranges in Canada, both survival and recruitment decrease with the percentage of the range that is disturbed.
-However, there is also variation in demographic rates among ranges, particularly where anthropogenic disturbance is low, and this variation leads to uncertainty in projections derived from national disturbance-demographic relationships.
-Demographic projections derived from local range-specific data are also uncertain because few places have enough long-term monitoring data to characterize local demographic-disturbance relationships.
-We present a simple Bayesian integrated population model that integrates prior information from national analysis of demographic-disturbance relationships with available local demographic data to reduce uncertainty in population viability projections.
-We combine this model with methods for simulating local population dynamics and monitoring to show how monitoring requirements depend on landscape condition.
-Where anthropogenic disturbance is very high or very low, reasonably accurate status projections can be made using only national demographic-disturbance relationships.
-At intermediate disturbance levels where uncertainty is higher, deploying few collars for more years may be a lower cost option for improving the reliability of status projections relative to more intensive short-term monitoring.
-We provide an R Shiny interface to enable others to explore the implications of alternative monitoring scenarios, disturbance trajectories, priors, and other model parameters for boreal caribou demographic projections.
-This tool can help users assess monitoring needs and strategies.
 # Introduction
+
+NOTE: find working version of intro and discussion [here:] (https://docs.google.com/document/d/1RD7bJ6yhtBJULjJFRJfHzObQscteyObM/edit?usp=sharing&ouid=111541730342945675365&rtpof=true&sd=true).
 
@@
 
-cd <- here::here()
-app_path <- gsub("Caribou-Demographic-Projection-Paper","BayesianCaribouDemographicProjection",cd,fixed=T)
-
 pal_nm <- "Redon"
@@
 ```
-TO DO: build large table of all parameters as the various sections are written. Adjust UI for consistent terminology along the way.
 
@@
 
-## A simple Bayesian IPM to integrate local demographic data and national disturbance-demographic relationships
+## Integration of local demographic data and national disturbance-demographic relationships in Bayesian population model
 
-Following [@eacker_web-based_2019], the observed number of calves $c_t$ is a binomially distributed function of the observed number of adult female caribou $a_t$ and the estimated recruitment probability $R_t$: $$c_t \sim R_t^{c_t}(1-R_t)^{a_t-c_t}.$$
-We model recruitment probability as a function of anthropogenic disturbance and fire [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022], with a log link and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a anthro_t+\beta^R_f fireExclAnthro_t+\epsilon^R_t; \epsilon^R_t \sim \text{normal}(0,\sigma^2_{R}).$$
-Estimated recruitment probability is adjusted for sex ratio and delayed age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012] to get expected recruits per female: $$\tilde{R}_t=\frac{R_t/2}{1+R_t/2}.$$
+Following Eacker et al.
+[-@eacker_web-based_2019], the observed number of calves $\hat{C}_t$ is a binomially distributed function of the observed number of adult female caribou $\hat{W}_t$ and the estimated recruitment rate $R_t$: $$\hat{C}_t \sim \text{Binomial}(\hat{W}_t,R_t).$$
+
+We model recruitment probability as a function of anthropogenic disturbance and fire with a log link [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a a_t+\beta^R_f f_t+\epsilon^R_t; \epsilon^R_t \sim \text{Normal}(0,\sigma^2_{R}).$$
+Estimated recruitment probability is adjusted for sex ratio to get expected recruits per female: $\tilde{R}_t=R_t/2$.
+We include an option to also adjust for delayed age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012], but to maintain consistency with Johnson et al.
+[-@johnson_science_2020] we do not apply this adjustment by default: $$\tilde{R}_t=\frac{R_t/2}{1+R_t/2}.$$
 
 Also following [@eacker_web-based_2019], Kaplan-Meier estimates of observed adult female survival probability $\hat{S}_t$ and associated precisions $\tau_t$ are estimated from known-fate radio collar data using the survival R package [@therneau_modeling_2000], with precisions for years with no observed mortality estimated using a separate Bayesian model.
-Observed estimated survival is a Gaussian distributed function of the true survival probability: $$\hat{S_t} \sim \text{normal}(S_t,1/\tau_t).$$
-We model true survival probability as an adjusted function of anthropogenic disturbance  [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022], with a log link and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a anthro_t+\epsilon^S_t; \epsilon^S_t \sim \text{normal}(0,\sigma^2_{S}).$$
-To assess sensitivity of results to the survival estimation method, we also implemented an alternative parametric exponential survival model, wherein the state of an individual animal in year $t$ and month $m$ is Bernoulli distributed (ref IPM book): $$A_{t,m} \sim \text{Bernoulli}(S^{1/12}_t A_{t,m-1}).$$
+Observed estimated survival is a Gaussian distributed function of the true survival probability: $$\hat{S_t} \sim \text{Normal}(S_t,1/\tau_t).$$
+We model true survival probability as an adjusted function of buffered anthropogenic disturbance $a_t$ with a log link [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a a_t+\epsilon^S_t; \epsilon^S_t \sim \text{Normal}(0,\sigma^2_{S}).$$
+To assess sensitivity of results to the survival estimation method, we also implemented an alternative parametric exponential survival model, wherein the observed state (alive or dead, $\hat{I}_{i,t,m}$) of an individual animal $i$ in year $t$ and month $m$ is Bernoulli distributed (ref IPM book): $$\hat{I}_{i,t,m} \sim \text{Bernoulli}(S^{1/12}_t\hat{I}_{i,t,m-1}).$$
 
 All regression coefficients are assumed to be Gaussian distributed, with priors derived from the expected values and 95% confidence intervals estimated by Johnson et al (2020) (Table \@ref(tab:priors)).
-We calibrated the prior distributions so that the 95% prediction interval from the IPM matches the range between the 2.5% and 97.5% quantiles of 1000 simulated recruitment (Supplement X).
+We calibrated the prior distributions so that the 95% prediction interval from the IPM matches the range between the 2.5% and 97.5% quantiles of 1000 simulated recruitment (Appendix A).
 
@@
 
-knitr::kable(tbl,caption="Model priors parameters are Gaussian distributed with expected values ")
+knitr::kable(tbl,caption="Prior means and standard deviations of Gaussian distributed survival and reproduction model parameters.")
+#TO DO: indicate which of these are from Johnson, and which are calibrated.
 ```
 
+We assume a simple two-class demographic model, with a census in late winter of surviving females and new recruits to the post-juvenile (aged 1+) class.
+The number of post-juvenile females that survive from year $t$ to the census $A_t$ is binomially distributed with survival probability $S_t$: $A_{t} \sim \text{Binomial}(N_t,S_t)$.
+The number of juveniles recruiting to the post-juvenile class at the census is a Poisson distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate $\tilde{R}_t$: $J_{t} \sim \text{Poisson}(\tilde{R}_tA_t)$.
+The post-juvenile female population in the next year includes both survivors and new recruits: $N_{t+1}=A_t+J_t$.
+Annual population growth rate $\lambda_t$ is set to 0 when $N_t$ is 0, and is otherwise $\lambda_t=N_{t+1}/N_t$.
+For more stable estimates of population status, we report average population growth over a period of $y$ years, calculated as $\bar{\lambda}_{y,t}=\sum_{x=t-y+1}^{x=t}\lambda_x$.
+The initial post-juvenile population size $N_0$ is 1000 animals.
+For reference, all variables and parameters are defined in Table A1.
+
+<!--
+Annotated equations for poster.
+$$\underbrace{N_{t+1}}_{\text{Adult females in year t+1}}=\underbrace{A_t}_{\text{Surviving adults}}+\underbrace{J_t}_{\text{New recruits to adult age class}}$$
+$$A_{t} \sim \text{Binomial}(N_t,\underbrace{S_t}_{\text{Adult survival probability}})$$
+$$S_t=\underbrace{\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})}_{\text{Adjustment as in in Johnson et al.
+2020}}=\beta^S_0+ \underbrace{\beta^S_aa_t}_{\text{Anthropogenic disturbance effect}}+\underbrace{\epsilon^S_t}_{\text{Random effect of year}}.$$
+$$J_{t} \sim \text{Poisson}(A_t\underbrace{R_t/2}_{\text{Recruitment rate}})$$
+$$\log(R_t)=\beta^R_0+\underbrace{\beta^R_a a_t}_{\text{Anthropogenic disturbance effect}}+ \underbrace{\beta^R_ff_t}_{\text{Fire effect}}+\underbrace{\epsilon^R_t}_{\text{Random effect of year}}.$$
+$$\underbrace{\hat{C}_t}_{\text{Observed calves}} \sim \text{Binomial}(\underbrace{\hat{W}_t}_{\text{Observed cows}},R_t)$$
+$$\underbrace{\hat{S_t}}_{\text{Observed survival (Kaplan-Meier estimate)}} \sim \text{Normal}(S_t,1/\tau_t).$$ -->
+
 TO DO: 
-
-* ensure that INF_saf_INF_r_JAGS_extentTime.txt and INF_saf_INF_r_JAGS_expSurvival2.txt are identical except for the survival observation model.
-Or better yet, write code to sub in correct survival model more elegantly.
-* note various bits of truncation in the JAGS code to ensure R and S remain between 0 and 1 (not guaranteed given log link).
-Not all documented in this explanation.
-Bayesian wizard might know a better way.
-Or at least know how to explain.
+* note distribution of random effect parameters are truncated to ensure R and S remain within reasonable bounds (not guaranteed given log link). Not yet documented here.
 * note Eacker survival method gives biased results in years when <12 months of survival data is available.
@@
 * consider & comment on alternative choices made in this example https://poissonconsulting.shinyapps.io/bboushiny/. Follow Eacker example in all particulars, change some aspects, or do comparisons?
+* Ensure notation consistency amongst manuscript, UI, documentaiton and code where appropriate.
 
-## A method for simulating example data
+## A method for simulating example local demographic data
 
@@
 # Appendix
+##Definitions of variables and parameters in the Bayesian integrated population model
+
+Table A1. Variables and parameters in the Bayesian integrated population model.
+
+|**Symbol**|**Units**|**Description**|
+|:-----|-----|:-----------|
+|**Indices & constants**|
+|$t$|year|Year.|
+|$m$|month|Month.|
+|$i$|ID|Individual animal identifier.|
+|$y$|years|Assessment period for population growth rate.|
+|$N_0$|animals|Initial adult population size.|
+|**Observed**|
+|$a_t$|%|Buffered anthropogenic disturbance.|
+|$f_t$|%|Fire disturbance.| 
+|$\hat{C}_t$|animals|Calves observed in calf:cow survey.|
+|$\hat{W}_t$|animals|Cows observed in calf:cow survey.|
+|$\hat{S}_t$|probability|Kaplan-Meier estimate of observed adult female survival.|
+|$\tau_t$||Precision of Kaplan-Meier survival estimate.|
+|$\hat{I}_{i,t,m}$|0 or 1|Indicator of whether animal $i$ is dead or alive.|
+|**Reported state**|
+|$R_t$|rate|Recruitment (calves per cow).|
+|$S_t$|probability|Adult female survival.|
+|$N_t$|animals|Adult female population size.|
+|$\bar{\lambda}_{y,t}$|rate|Average population growth over $y$ years.|
+|**Other state**|
+|$\tilde{R}_t$|rate|Recruitment adjusted for sex ratio and (optionally) delayed age of first reproduction.|
+|$A_t$|animals|Surviving adult females before recruitment.|
+|$J_t$|animals|New recruits to adult age class.|
+|$\lambda_t$|rate|Annual population growth.|
+|**Parameters**|
+|$\beta^R_0$||Recruitment intercept.|
+|$\beta^R_a$||Anthropogenic disturbance slope for recruitment.|
+|$\beta^R_f$||Fire slope for recruitment.|
+|$\sigma^2_{R}$||Variance in random effect of year on recruitment.|
+|$\beta^S_0$||Survival intercept.|
+|$\beta^S_a$||Anthropogenic disturbance slope for survival.|
+|$\sigma^2_{S}$||Variance in random effect of year on survival.|
 

### 2023-05-08 analysis/scripts/sensitivityMinimalTestPar.R
try using sink to keep output

@@ -19,6 +19,8 @@ future::plan("multisession")
 
 foreach(cpageId = 1:n_batches, .options.future = list(seed = TRUE)) %dofuture% {
 
+  sink(paste0("outbatch_", cpageId, ".txt"))
+
   simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
 
   allScns = read.csv(paste0("tabs/",setName,".csv"))
@@ -45,6 +47,7 @@ foreach(cpageId = 1:n_batches, .options.future = list(seed = TRUE)) %dofuture% {
   saveRDS(scResults,paste0("results/",setName,"/rTest",cpageId,".Rds"))
 
   message("batch ", cpageId, " complete")
+  sink()
 
 }
 future::plan("sequential")

### 2023-05-08 analysis/scripts/sensitivityMinimal.R
do same for main script

@@ -17,27 +17,31 @@ dir.create(paste0("figs/",setName),recursive=T)
 dir.create(paste0("tabs/",setName),recursive=T)
 dir.create(paste0("results/",setName),recursive=T)
 
-simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
 
-allScns = read.csv(paste0("tabs/",setName,".csv"))
-n_batches <- length(unique(allScns$pageId))
-
-####################
-eParsIn = list()
-eParsIn$cowCounts <- data.frame(Year = 1981:2023,
-                                Count = 100,
-                                Class = "cow")
-eParsIn$freqStartsByYear <- data.frame(Year = 1981:2023,
-                                       numStarts = 30)
-eParsIn$collarOnTime=1
-eParsIn$collarOffTime=12
-eParsIn$collarNumYears=6
 
 future::plan("multisession")
 
 foreach(cpageId = 1:n_batches, .options.future = list(seed = TRUE)) %dofuture% {
 
+  simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
+
+  allScns = read.csv(paste0("tabs/",setName,".csv"))
+  n_batches <- length(unique(allScns$pageId))
+
+  ####################
+  eParsIn = list()
+  eParsIn$cowCounts <- data.frame(Year = 1981:2023,
+                                  Count = 100,
+                                  Class = "cow")
+  eParsIn$freqStartsByYear <- data.frame(Year = 1981:2023,
+                                         numStarts = 30)
+  eParsIn$collarOnTime=1
+  eParsIn$collarOffTime=12
+  eParsIn$collarNumYears=6
+
   scns = subset(allScns, pageId==cpageId)
+
+  rm(allScns)
 
   message("batch ", cpageId, " started")
 

### 2023-05-08 analysis/scripts/sensitivityMinimal.R
fix

@@ -21,7 +21,7 @@ dir.create(paste0("results/",setName),recursive=T)
 
 future::plan("multisession")
 
-foreach(cpageId = 1:n_batches, .options.future = list(seed = TRUE)) %dofuture% {
+foreach(cpageId = 1:24, .options.future = list(seed = TRUE)) %dofuture% {
 
   simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
 

### 2023-05-15 analysis/paper/paper.Rmd
Fixing scenario set 4

@@
 runScenarios = F 
-doPriorCalibration=F #Note the prior calibration is time consuming. 
+doPriorCalibration=T #Note the prior calibration is time consuming. 
 ```
@@
 
-We model recruitment probability as a function of anthropogenic disturbance and fire with a log link [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a a_t+\beta^R_f f_t+\epsilon^R_t; \epsilon^R_t \sim \text{Normal}(0,\sigma^2_{R}).$$
-Estimated recruitment probability is adjusted for sex ratio to get expected recruits per female: $\tilde{R}_t=R_t/2$.
+We model recruitment probability as a function of anthropogenic disturbance and fire with a log link [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a A_t+\beta^R_f F_t+\epsilon^R_t; \epsilon^R_t \sim \text{Normal}(0,\sigma^2_{R}).$$
+Estimated recruitment probability is adjusted for sex ratio to get expected recruits per female: $X_t=R_t/2$.
 We include an option to also adjust for delayed age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012], but to maintain consistency with Johnson et al.
-[-@johnson_science_2020] we do not apply this adjustment by default: $$\tilde{R}_t=\frac{R_t/2}{1+R_t/2}.$$
+[-@johnson_science_2020] we do not apply this adjustment by default: $$X_t=\frac{R_t/2}{1+R_t/2}.$$
 
@@
 Observed estimated survival is a Gaussian distributed function of the true survival probability: $$\hat{S_t} \sim \text{Normal}(S_t,1/\tau_t).$$
-We model true survival probability as an adjusted function of buffered anthropogenic disturbance $a_t$ with a log link [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a a_t+\epsilon^S_t; \epsilon^S_t \sim \text{Normal}(0,\sigma^2_{S}).$$
+We model true survival probability as an adjusted function of buffered anthropogenic disturbance $A_t$ with a log link [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a A_t+\epsilon^S_t; \epsilon^S_t \sim \text{Normal}(0,\sigma^2_{S}).$$
 To assess sensitivity of results to the survival estimation method, we also implemented an alternative parametric exponential survival model, wherein the observed state (alive or dead, $\hat{I}_{i,t,m}$) of an individual animal $i$ in year $t$ and month $m$ is Bernoulli distributed (ref IPM book): $$\hat{I}_{i,t,m} \sim \text{Bernoulli}(S^{1/12}_t\hat{I}_{i,t,m-1}).$$
@@
 We assume a simple two-class demographic model, with a census in late winter of surviving females and new recruits to the post-juvenile (aged 1+) class.
-The number of post-juvenile females that survive from year $t$ to the census $A_t$ is binomially distributed with survival probability $S_t$: $A_{t} \sim \text{Binomial}(N_t,S_t)$.
-The number of juveniles recruiting to the post-juvenile class at the census is a Poisson distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate $\tilde{R}_t$: $J_{t} \sim \text{Poisson}(\tilde{R}_tA_t)$.
-The post-juvenile female population in the next year includes both survivors and new recruits: $N_{t+1}=A_t+J_t$.
+The number of post-juvenile females that survive from year $t$ to the census $W_t$ is binomially distributed with survival probability $S_t$: $W_{t} \sim \text{Binomial}(N_t,S_t)$.
+The number of juveniles recruiting to the post-juvenile class at the census is a Poisson distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate $X_t$: $J_{t} \sim \text{Poisson}(X_tW_t)$.
+The post-juvenile female population in the next year includes both survivors and new recruits: $N_{t+1}=W_t+J_t$.
 Annual population growth rate $\lambda_t$ is set to 0 when $N_t$ is 0, and is otherwise $\lambda_t=N_{t+1}/N_t$.
@@
 Annotated equations for poster.
-$$\underbrace{N_{t+1}}_{\text{Adult females in year t+1}}=\underbrace{A_t}_{\text{Surviving adults}}+\underbrace{J_t}_{\text{New recruits to adult age class}}$$
-$$A_{t} \sim \text{Binomial}(N_t,\underbrace{S_t}_{\text{Adult survival probability}})$$
+$$\underbrace{N_{t+1}}_{\text{Adult females in year t+1}}=\underbrace{W_t}_{\text{Surviving adults}}+\underbrace{J_t}_{\text{New recruits to adult age class}}$$
+$$W_{t} \sim \text{Binomial}(N_t,\underbrace{S_t}_{\text{Adult survival probability}})$$
 $$S_t=\underbrace{\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})}_{\text{Adjustment as in in Johnson et al.
-2020}}=\beta^S_0+ \underbrace{\beta^S_aa_t}_{\text{Anthropogenic disturbance effect}}+\underbrace{\epsilon^S_t}_{\text{Random effect of year}}.$$
-$$J_{t} \sim \text{Poisson}(A_t\underbrace{R_t/2}_{\text{Recruitment rate}})$$
-$$\log(R_t)=\beta^R_0+\underbrace{\beta^R_a a_t}_{\text{Anthropogenic disturbance effect}}+ \underbrace{\beta^R_ff_t}_{\text{Fire effect}}+\underbrace{\epsilon^R_t}_{\text{Random effect of year}}.$$
+2020}}=\beta^S_0+ \underbrace{\beta^S_aA_t}_{\text{Anthropogenic disturbance effect}}+\underbrace{\epsilon^S_t}_{\text{Random effect of year}}.$$
+$$J_{t} \sim \text{Poisson}(W_t\underbrace{R_t/2}_{\text{Recruitment rate}})$$
+$$\log(R_t)=\beta^R_0+\underbrace{\beta^R_a A_t}_{\text{Anthropogenic disturbance effect}}+ \underbrace{\beta^R_fF_t}_{\text{Fire effect}}+\underbrace{\epsilon^R_t}_{\text{Random effect of year}}.$$
 $$\underbrace{\hat{C}_t}_{\text{Observed calves}} \sim \text{Binomial}(\underbrace{\hat{W}_t}_{\text{Observed cows}},R_t)$$
@@
 ### True population dynamics
-Write for now assuming that caribouMetrics paper will be published somewhere.
+To simulate plausible plausible demographic trajectories for example populations we used a national model described by Dyson et al.
+[-@dyson_existing_2022] and Stewart et al.
+[-@stewart_climate-informed_2023].
+Expected recruitment ($\bar{R}_t$) and survival ($\bar{S}_t$) vary with disturbance according the beta regression models estimated by Johnson et al.
+[@johnson_science_2020]:  :$$\bar{R}_t~Beta(\mu^R_t,\phi^R);log(\mu^R_t)=\dot{\beta^R_0}+\dot{\beta^R_a}A_t+\dot{\beta}^R_fF_t,$$
+$$\bar{S}_t~(46\times Beta(\mu^S_t,\phi^S)-0.5)/45;log(\mu^S_t)=\dot{\beta^S_0}+\dot{\beta^S_a}A_t.$$
+$\phi^R~Normal(19.862,2.229)$ and $\phi^S~Normal(63.733,8.311)$ are precisions of the Beta distributed errors [@ferrari_beta_2004].
+At the beginning of a simulation for an example population, regression coefficient values are sampled from Gaussian distributions (see Table \@ref(tab:priors)) and the population is assigned to quantiles of the Beta error distributions for survival and recruitment.
+The population remains in these quantiles as disturbance changes over time, so there is substantial persistent variation in recruitment and survival among example among populations (example figure).
+There is also interannual variation in recruitment for each population, modelled using truncated beta distributions (rtrunc function [@novomestky_package_2016]): $$\dot{R}_t~$$
 
@@
 
-Table A1. Variables and parameters in the Bayesian integrated population model.
+Table A1. Variables and parameters in the Bayesian integrated population model and the national demographic model.
 
@@
 |**Observed**|
-|$a_t$|%|Buffered anthropogenic disturbance.|
-|$f_t$|%|Fire disturbance.| 
+|$A_t$|%|Buffered anthropogenic disturbance.|
+|$F_t$|%|Fire disturbance.| 
 |$\hat{C}_t$|animals|Calves observed in calf:cow survey.|
@@
 |$\hat{I}_{i,t,m}$|0 or 1|Indicator of whether animal $i$ is dead or alive.|
-|**Reported state**|
+|**Estimated state**|
 |$R_t$|rate|Recruitment (calves per cow).|
@@
 |$\bar{\lambda}_{y,t}$|rate|Average population growth over $y$ years.|
-|**Other state**|
-|$\tilde{R}_t$|rate|Recruitment adjusted for sex ratio and (optionally) delayed age of first reproduction.|
-|$A_t$|animals|Surviving adult females before recruitment.|
+|$X_t$|rate|Recruitment adjusted for sex ratio and (optionally) delayed age of first reproduction.|
+|$W_t$|animals|Surviving adult females before recruitment.|
 |$J_t$|animals|New recruits to adult age class.|
 |$\lambda_t$|rate|Annual population growth.|
-|**Parameters**|
+|**Estimated Parameters**|
 |$\beta^R_0$||Recruitment intercept.|
@@
 # COMMENTED OUT FOR NOW CAUSE I DIDN"T HAVE ACCESS
-#knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.03.png"))
+knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.03.png"))
 ```
@@
 # COMMENTED OUT FOR NOW CAUSE I DIDN"T HAVE ACCESS
-#knitr::include_graphics(here::here("figs/SurvsQStarts45lowssv0.03.png"))
+knitr::include_graphics(here::here("figs/SurvsQStarts45lowssv0.03.png"))
 ```

### 2023-05-15 R/CaribouDemographicFunctions.R

@@ -18,7 +18,7 @@ runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMei
   library(tidyr)
   library(dplyr)
 
-  scns<-fillDefaults(scns)
+  scns<-getScenarioDefaults(scns)
 
   eParsIn = list()
   eParsIn$cowCounts <- data.frame(Year = 1981:2023,

### 2023-05-15 analysis/scripts/sensitivity1.R

@@ -19,25 +19,27 @@
 #TO DO: adjust method so all
 
 setName = "s4"
-monitoringScns = expand.grid(P=c(1,2,4,8,16,24),st=c(1,15,30,60),cmult=c(3,6,9),ri=c(1,4),assessmentYrs=c(3))
-stateScns = data.frame(tA=c(0,20,40,60,0,20,40,60),aS=c(0,1,1,1,0,0,0,0),aSf=c(1,1,1,1,0,0,0,0))
+monitoringScns = expand.grid(obsYears=c(1,2,4,8,16,24),collarCount=c(1,15,30,60),cowMult=c(3,6,9),collarInterval=c(1,4),assessmentYrs=c(3))
+stateScns = data.frame(tA=c(0,20,40,60,0,20,40,60),obsAnthroSlope=c(0,1,1,1,0,0,0,0),projAnthroSlope=c(1,1,1,1,0,0,0,0))
 stateScns = merge(stateScns,data.frame(rep=seq(1:200)))
 
-stateScns$sQ=runif(nrow(stateScns),min=0.01,max=0.99)
-stateScns$rQ = runif(nrow(stateScns),min=0.01,max=0.99)
+stateScns$sQuantile=runif(nrow(stateScns),min=0.01,max=0.99)
+stateScns$rQuantile = runif(nrow(stateScns),min=0.01,max=0.99)
 #monitoringScns=rbind(monitoringScns,minimalScn)
 scns=merge(monitoringScns,stateScns)
 
-scns$iA = scns$tA-(scns$P-1)*scns$aS
-unique(scns$iA)
+scns$iAnthro = scns$tA-(scns$obsYears-1)*scns$obsAnthroSlope
+scns$projYears = 20
+unique(scns$iAnthro)
 scns$repBatch = ceiling(scns$rep/50)
 table(scns$repBatch)
 scns$N0 = 1000
 nrow(scns)/6
 #scns=scns[10,]
 #scResults = readRDS("temp.Rds")
+head(scns)
 
-scns$pageLab = paste0("cmult",scns$cmult,"ay",scns$assessmentYrs,"aSf",scns$aSf,"repBatch",scns$repBatch)
+scns$pageLab = paste0("cmult",scns$cowMult,"ay",scns$assessmentYrs,"aSf",scns$projAnthroSlope,"repBatch",scns$repBatch)
 scns$pageId = as.numeric(as.factor(scns$pageLab))
 unique(scns$pageId)
 

### 2023-05-15 analysis/scripts/sensitivityMinimalTest.R

@@ -1,9 +1,9 @@
 #!/usr/bin/env Rscript
-n_scns <- 2
+n_reps <- 3
 
 # Run batches from Rscript that uses parallel backend and new caribouMetrics functions
 cpageId <- commandArgs(trailingOnly = TRUE)
-# cpageId <- 1
+# cpageId <- 5
 
 library(caribouMetrics)
 
@@ -18,7 +18,7 @@ dir.create(paste0("results/",setName),recursive=T)
 simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
 
 allScns = read.csv(paste0("tabs/",setName,".csv"))
-
+unique(allScns$pageId)
 ####################
 eParsIn = list()
 eParsIn$cowCounts <- data.frame(Year = 1981:2023,
@@ -28,15 +28,14 @@ eParsIn$freqStartsByYear <- data.frame(Year = 1981:2023,
                                        numStarts = 30)
 eParsIn$collarOnTime=1
 eParsIn$collarOffTime=12
-eParsIn$collarNumYears=3
+eParsIn$collarNumYears=6
 
 scns = subset(allScns, pageId==cpageId)
 
-rm(allScns)
-
+nrow(subset(scns,rep<=n_reps))
 message("batch ", cpageId, " started")
 
-scResults = caribouMetrics:::runScnSet(scns[1:n_scns,],eParsIn,simBig,getKSDists=F,printProgress=F)
+scResults = caribouMetrics:::runScnSet(subset(scns,rep<=n_reps),eParsIn,simBig,getKSDists=F,printProgress=F)
 
 saveRDS(scResults,paste0("results/",setName,"/rTest",cpageId,".Rds"))
 

### 2023-05-15 analysis/scripts/sensitivityMinimalTestPar.R

@@ -34,7 +34,7 @@ foreach(cpageId = 1:n_batches, .options.future = list(seed = TRUE)) %dofuture% {
                                          numStarts = 30)
   eParsIn$collarOnTime=1
   eParsIn$collarOffTime=12
-  eParsIn$collarNumYears=3
+  eParsIn$collarNumYears=6
 
   scns = subset(allScns, pageId==cpageId)
 

### 2023-05-15 analysis/scripts/viewSensitivityResults.R

@@ -11,14 +11,16 @@ scn_defaults <- eval(formals(getScenarioDefaults))
 
 ########################
 #sensitivity
-setName = "s3"
-scns = read.csv(here::here("tabs/s3.csv"))
+setName = "s4"
+dir.create(paste0("figs/",setName),recursive=T)
+scns = read.csv(here::here("tabs/s4.csv"))
 scns = subset(scns,cmult!=0)
 pagesa=unique(scns$pageId)
 
 pages = pagesa
 for (p in pagesa){
-  if(!file.exists(paste0("results/",setName,"/r",p,".Rds"))){pages=pages[pages!=p]}
+  #p=pagesa[1]
+  if(!file.exists(paste0("results/",setName,"/rTest",p,".Rds"))){pages=pages[pages!=p]}
 }
 
 batchStrip<-function(l,batches=seq(1:4)){
@@ -44,7 +46,7 @@ for(i in 1:length(pages)){
   p = unique(scns$pageLab[scns$pageId==cpageId])
 
   if(combine&(cpageId>1)){
-    scNew = readRDS(paste0("results/",setName,"/r",cpageId,".Rds"))
+    scNew = readRDS(paste0("results/",setName,"/rTest",cpageId,".Rds"))
     for(n in names(scNew)){
       if(n=="errorLog"){
         scResults[[n]]=c(scResults[[n]],scNew[[n]])
@@ -53,7 +55,7 @@ for(i in 1:length(pages)){
       }
     }
   }else{
-    scResults = readRDS(paste0("results/",setName,"/r",cpageId,".Rds"))
+    scResults = readRDS(paste0("results/",setName,"/rTest",cpageId,".Rds"))
   }
 
   if(as.numeric(strsplit(nextP,"repBatch")[[1]][2])<=as.numeric(strsplit(p,"repBatch")[[1]][2])){
@@ -64,10 +66,12 @@ for(i in 1:length(pages)){
 
   #figure out how to count out errors.
 
+  head(scResults$rr.summary.all)
+  unique(scResults$rr.summary.all$collarInterval)
   #show examples projections
-  exResults = subset(scResults$rr.summary.all,(st==30)&(ri==4)&(Parameter=="Population growth rate"))
-  exResults$meanQ = (exResults$rQ+exResults$sQ)/2
-  grpID = subset(exResults,select=c(tA,P,meanQ)) %>% group_by(tA,P) %>%
[…]

### 2023-06-01 analysis/paper/paper.Rmd
notation work  in progress.

@@
 Following Eacker et al.
-[-@eacker_web-based_2019], the observed number of calves $\hat{C}_t$ is a binomially distributed function of the observed number of adult female caribou $\hat{W}_t$ and the estimated recruitment rate $R_t$: $$\hat{C}_t \sim \text{Binomial}(\hat{W}_t,R_t).$$
+[-@eacker_web-based_2019], the observed number of calves $\hat{J}_t$ is a binomially distributed function of the observed number of adult female caribou $\hat{W}_t$ and the estimated recruitment rate $R_t$: $$\hat{J}_t \sim \text{Binomial}(\hat{W}_t,R_t).$$
 
@@
 
+$$J_{t} \sim \text{Poisson}(W_tR_t/2)$$
 <!--
@@
 
-### True population dynamics
-To simulate plausible plausible demographic trajectories for example populations we used a national model described by Dyson et al.
-[-@dyson_existing_2022] and Stewart et al.
-[-@stewart_climate-informed_2023].
+### Population dynamics
+To simulate plausible plausible demographic trajectories for example populations we used the slightly modified version of Johnson et al's [-@johnson_science_2020] two-stage demographic model described by Dyson et al.
+[-@dyson_existing_2022].
+The true number of post-juvenile females that survive from year $t$ to the census $\dot{W}_t$ is binomially distributed with true survival probability $\dot{S}_t$: $\dot{W}_{t} \sim \text{Binomial}(\dot{N}_t,\dot{S}_t)$.
+As in the IPM, maximum potential recruitment rate is adjusted for sex ratio and (optionally) delayed age at first reproduction $$\dot{X}_t=\frac{\dot{R}_t/2}{1+\dot{R}_t/2}.$$ Realized recruitment rate varies with population density [@lacy_vortex_2017], and the number of juveniles recruiting to the post-juvenile class at the census is a binomially distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate: $$\dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{N_0k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}).$$ Given parameters in Table X, recruitment rate is lowest $(0.5\dot{X}_t)$ when $\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$ at intermediate population sizes, and declines to $0.6\dot{X}_t$ as the population reaches carrying capacity of $k=100$ times the initial population size.
+The post-juvenile female population in the next year includes both survivors and new recruits: $\dot{N}_{t+1}=\text{min}(\dot{W}_t+\dot{J}_t,r_{max}\dot{N}_t)$.
+
+Interannual variation in survival and recruitment is modelled using truncated beta distributions [rtrunc function; @novomestky_package_2016]: $\dot{R}_t \sim \text{TruncatedBeta}(\bar{R}_t,\nu_R,l_R,h_R); \dot{S}_t \sim \text{TruncatedBeta}(\bar{S}_t,\nu_S,l_S,h_S)$.
+Coefficients of variation among years $(\nu_R,\nu_S)$ and maximum/minimum values $l_R,h_R,l_S,h_S$ for recruitment and survival are given in Table X.
 Expected recruitment ($\bar{R}_t$) and survival ($\bar{S}_t$) vary with disturbance according the beta regression models estimated by Johnson et al.
-[@johnson_science_2020]:  :$$\bar{R}_t~Beta(\mu^R_t,\phi^R);log(\mu^R_t)=\dot{\beta^R_0}+\dot{\beta^R_a}A_t+\dot{\beta}^R_fF_t,$$
-$$\bar{S}_t~(46\times Beta(\mu^S_t,\phi^S)-0.5)/45;log(\mu^S_t)=\dot{\beta^S_0}+\dot{\beta^S_a}A_t.$$
-$\phi^R~Normal(19.862,2.229)$ and $\phi^S~Normal(63.733,8.311)$ are precisions of the Beta distributed errors [@ferrari_beta_2004].
+[@johnson_science_2020]: $$ \bar{R}_t \sim Beta(\mu^R_t,\phi^R);log(\mu^R_t)=\dot{\beta^R_0}+\dot{\beta^R_a}A_t+\dot{\beta}^R_fF_t,$$
+$$\bar{S}_t \sim (46\times Beta(\mu^S_t,\phi^S)-0.5)/45;log(\mu^S_t)=\dot{\beta^S_0}+\dot{\beta^S_a}A_t.$$
+$\phi^R \sim \text{Normal}(19.862,2.229)$ and $\phi^S \sim \text{Normal}(63.733,8.311)$ are precisions of the Beta distributed errors [@ferrari_beta_2004].
 At the beginning of a simulation for an example population, regression coefficient values are sampled from Gaussian distributions (see Table \@ref(tab:priors)) and the population is assigned to quantiles of the Beta error distributions for survival and recruitment.
 The population remains in these quantiles as disturbance changes over time, so there is substantial persistent variation in recruitment and survival among example among populations (example figure).
-There is also interannual variation in recruitment for each population, modelled using truncated beta distributions (rtrunc function [@novomestky_package_2016]): $$\dot{R}_t~$$
 
@@
 |$\sigma^2_{S}$||Variance in random effect of year on survival.|
+|**True state**|
+|$\dot{R}_t$|rate|Recruitment (calves per cow).|
+|$\dot{S}_t$|probability|Adult female survival.|
+|$\dot{N}_t$|animals|Adult female population size.|
+|$\dot{X}_t$|rate|Recruitment adjusted for sex ratio and (optionally) delayed age of first reproduction.|
+|$\dot{W}_t$|animals|Surviving adult females before recruitment.|
+|$\dot{J}_t$|animals|New recruits to adult age class.|
+|**True parameters**|
+|$\dot{\beta}^R_0$||Recruitment intercept.|
+|$\dot{\beta}^R_a$||Anthropogenic disturbance slope for recruitment.|
+|$\dot{\beta}^R_f$||Fire slope for recruitment.|
+|$\dot{\beta}^S_0$||Survival intercept.|
+|$\dot{\beta}^S_a$||Anthropogenic disturbance slope for survival.|
+|$p_0=1$||Maximum recruitment multiplier.|
+|$p_k=0.6$||Recruitment multiplier at carrying capacity.|
+|$k=100$||Carrying capacity multiplier.|
+|$b=4$||Density dependence shape parameter.|
+|$a=1$||Allee effect parameter.|
+|$r_{max}=1.3$||Maximum population growth rate.|
+|$\nu_R=0.46$||Coefficient of variation in $\dot{R}_t$ among years.|
+|$\nu_S=0.08696$||Coefficient of variation in $\dot{S}_t$ among years.|
+|$l_R=0$||Minimum value for $\dot{R}_t$.|
+|$l_S=0$||Minimum value for $\dot{S}_t$.|
+|$h_R=0$||Maximum value for $\dot{R}_t$.|
+|$h_S=0$||Maximum value for $\dot{S}_t$.|
 

### 2023-06-01 analysis/scripts/formatScenarios.R

@@ -1,14 +1,14 @@
 
 library(caribouMetrics)
 
-rawScns<-read.csv(here::here("tabs/s2.csv"))
+rawScns<-read.csv(here::here("tabs/s5.csv"))
 nrow(rawScns)
-rawScns = subset(rawScns,st!=1)
+rawScns = subset(rawScns,collarCount!=1)
 
-monitoringPars = list(P=list(description = "monitoring duration",unit="years"),
-                      st=list(description="target # of collars", unit="cows"),
-                      cmult=list(description="calf:cow survey multiplier",unit = "cows per collared cow"),
-                      ri = list(description="time between collar deployments", unit="years"),
+monitoringPars = list(obsYears=list(description = "monitoring duration",unit="years"),
+                      collarCount=list(description="target # of collars", unit="cows"),
+                      cowMult=list(description="calf:cow survey multiplier",unit = "cows per collared cow"),
+                      collarInterval = list(description="time between collar deployments", unit="years"),
                       assessmentYrs = list(description="assessment duration",unit="years"))
 
 start=T

### 2023-06-01 analysis/scripts/viewSensitivityResults.R

@@ -7,6 +7,8 @@ library(RColorBrewer)
 pal2 = brewer.pal(7,"RdBu")[c(2,6)]
 pal4 = brewer.pal(5,"RdBu")[c(1,2,4,5)]
 
+pal4b = brewer.pal(5,"Purples")[c(2,3,4,5)]
+
 scn_defaults <- eval(formals(getScenarioDefaults))
 
 ########################
@@ -135,7 +137,7 @@ for(i in 1:length(pages)){
   base=ggplot(distScns,aes(x=Year,y=Anthro,col=DisturbanceScn,group=grp))+geom_line()+geom_point(data=timelineLabs)+
     theme_bw()+theme(axis.text.x = element_text(angle=90,vjust=0.5,hjust=1,size=8)) + labs(color="Anthropogenic\nDisturbance\nScenario")+
     scale_x_continuous(name="Timeline", breaks=timelineLabs$Year, labels=timelineLabs$Timeline)+ylab("Anthropogenic Disturbance")+
-    scale_color_discrete(type=rev(pal4))
+    scale_color_discrete(type=(pal4b))
   print(base)
   dev.off()
 

### 2023-07-12 analysis/paper/paper.Rmd
In progress - bias model description and integration

@@
 The true number of post-juvenile females that survive from year $t$ to the census $\dot{W}_t$ is binomially distributed with true survival probability $\dot{S}_t$: $\dot{W}_{t} \sim \text{Binomial}(\dot{N}_t,\dot{S}_t)$.
-As in the IPM, maximum potential recruitment rate is adjusted for sex ratio and (optionally) delayed age at first reproduction $$\dot{X}_t=\frac{\dot{R}_t/2}{1+\dot{R}_t/2}.$$ Realized recruitment rate varies with population density [@lacy_vortex_2017], and the number of juveniles recruiting to the post-juvenile class at the census is a binomially distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate: $$\dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{N_0k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}).$$ Given parameters in Table X, recruitment rate is lowest $(0.5\dot{X}_t)$ when $\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$ at intermediate population sizes, and declines to $0.6\dot{X}_t$ as the population reaches carrying capacity of $k=100$ times the initial population size.
+As in the IPM, maximum potential recruitment rate is adjusted for sex ratio, misidentification biases (Appendix X) and (optionally) delayed age at first reproduction $$\dot{X}_t=\frac{c\dot{R}_t/2}{1+c\dot{R}_t/2};c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}, $$
+where $w$ is a multiplier that defines the apparent number of adult females in the composition survey as a function of the number of collared animals, $q$ the ratio of young bulls to adult females, $u$ is the probability of misidentifying young bulls as adult females and vice versa, and $z$ is the probability of missing a calf.
+
+Realized recruitment rate varies with population density [@lacy_vortex_2017], and the number of juveniles recruiting to the post-juvenile class at the census is a binomially distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate: $$\dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{N_0k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}).$$ Given parameters in Table X, recruitment rate is lowest $(0.5\dot{X}_t)$ when $\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$ at intermediate population sizes, and declines to $0.6\dot{X}_t$ as the population reaches carrying capacity of $k=100$ times the initial population size.
 The post-juvenile female population in the next year includes both survivors and new recruits: $\dot{N}_{t+1}=\text{min}(\dot{W}_t+\dot{J}_t,r_{max}\dot{N}_t)$.
@@
 |$F_t$|%|Fire disturbance.| 
-|$\hat{C}_t$|animals|Calves observed in calf:cow survey.|
+|$\hat{J}_t$|animals|Calves observed in calf:cow survey.|
 |$\hat{W}_t$|animals|Cows observed in calf:cow survey.|
@@
 |$h_S=0$||Maximum value for $\dot{S}_t$.|
+
+## Model of bias in recruitment estimates from calf:cow surveys
+
+TO DO: get photo of calf:cow survey group. Label classes of animals to illustrate meaning of T, U, V, W, O, P, J, K.
+
+We assume each group of animals in a calf:cow survey contains one or more collared adult females ($T$), and may also include: uncollared adult females misidentified as young bulls ($U$); correctly identified uncollared adult females ($V$); correctly identified young bulls ($O$); young bulls misidentified as uncollared adult females ($P$); observed calves ($J$); and unobserved calves ($K$).
+The apparent number of adult females in the group is $T+V+P=Tw$, where $w$ is a multiplier that defines the apparent number of adult females as a function of the number of collared animals.
+The ratio of young bulls to uncollared adult females in the group is:
+$$q = \frac{P+O}{U+V}.$$
+Assuming an equal probability $u$ of misidentifying young bulls as adult females and vice versa, we get $V=(U+V)(1-u)$ and $P=(O+P)u$. Given a probability $z$ of missing calves, we get $J=(J+K)(1-z)$.
+
+Our objective is to model the sex and bias-corrected recruitment rate $X=\frac{J+K}{2(T+U+V)}$ as a function of the observed calf:cow ratio $R=J/(T+V+P)$, the cow multiplier $w$, the ratio of young bulls to adult females $q$, and the misidentification probabilities $u$ and $z$.
+We start by solving for $T+U+V$ as a function of $q,w,u$ and $T$.
+Recognize that $P=Tw-T-V$, $U+V=V/(1-u)$, and $P+O=P/u$ to write $q$ as
+$$q=\frac{Tw-T-V}{uV/(1-u)}.$$
+Rearrange to get 
+$$V=\frac{T(w-1)(1-u)}{qu+1-u}.$$
+Recognize that $U=Vu/(1-u)$ to write $T+U+V$ as a function of $q,w,u$ and $T$: 
+$$T+U+V=T\frac{qu+w-u}{qu+1-u}.$$
+Recognize that the number of observed calves $J$ is the product of the apparent recruitment rate and the apparent number of adult females $J=RTw$, and that therefore $J+K=RTw/(1-z)$ to rewrite the bias corrected recruitment rate $X=\frac{J+K}{2(T+U+V)}$ as a function of $w,u,z$ and $R$:
+$$X=R\frac{w(1+qu-u)}{2(w+qu-u)(1-z)}.$$
+For simplicity, we write $X$ as a function of a bias correction term $c$: 
+$$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}; X=cR/2$$
+
+If we also choose to adjust for delayed age at first reproduction (), the adjusted recruitment rate becomes $$X=\frac{cR/2}{1+cR/2}.$$
+
+```{r biasPlot, fig.cap="Variation in the sex and bias corrected recruitment rate $X$ with the cow multiplier $w$, the ratio of young bulls to adult females $q$, and the adult misidentification probability $u$, and the probability of missing a calf $z$.
+If only collared females are counted ($w=1$) and there is a risk of missing calves $z>0$ then the apparent recruitment rate $R$ will be lower than $X$ (first column).
+When uncollared animals are counted ($w>1$), the risk of misidentifying adults $u$ inflates the apparent recruitment rate $R$ as on average more cows are misidentified as bulls than vice versa, but this effect is smaller than the effect of missing calves $z$.
+Increasing the proportion of bulls in the group $q$ reduces the effect of adult misidentification $u$.", cache=TRUE}
+library(ggplot2)
+#Notes
+#q: assume more females than males in calf:cow group, so q<1
+
+p=expand.grid(delay=c(T,F),q=c(0,0.3,0.6),w=c(1,3,6),
+              u=c(0,0.1,0.2),z=seq(0,0.2,length.out=3),R=seq(0.15,0.6,length.out=10))
+
+getX<-function(R,delay,q,w,u,z){
+  X = 0.5*R*w*(q*u+1-u)/((q*u+w-u)*(1-z))
+  X=X*(1-delay)+delay*X/(1+X)
+  return(X)
+}
+
+p$X = getX(p$R,p$delay,p$q,p$w,p$u,p$z)
+
+p$grp = paste(p$u,p$z)
+p$u=as.factor(p$u)
+p$z=as.factor(p$z)
+
+base = ggplot(subset(p,!delay),aes(x=R,y=2*X/R,col=z,linetype=u,group=grp))+
+  geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()+ylab("2X/R=c, without delayed reproduction")
+plot(base)
+
+base2 = ggplot(subset(p,delay),aes(x=R,y=2*X/R,col=z,linetype=u,group=grp))+
+  geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()+ylab("2X/R, with delayed reproduction")
+plot(base2)
+
+```
 

### 2023-07-12 analysis/scripts/BiasModelSensitivity.R

@@ -1,0 +1,27 @@
+library(ggplot2)
+#Notes
+#q: assume more females than males in calf:cow group, so q<1
+
+p=expand.grid(delay=c(T,F),q=c(0,0.3,0.6),w=c(1,3,6),
+              u=c(0,0.1,0.2),z=seq(0,0.2,length.out=3),R=seq(0.15,0.6,length.out=10))
+
+getX<-function(R,delay,q,w,u,z){
+  X = 0.5*R*w*(q*u+1-u)/((q*u+w-u)*(1-z))
+  X=X*(1-delay)+delay*X/(1+X)
+  return(X)
+}
+
+p$X = getX(p$R,p$delay,p$q,p$w,p$u,p$z)
+
+p$grp = paste(p$u,p$z)
+names(p)
+p$u=as.factor(p$u)
+p$z=as.factor(p$z)
+
+base = ggplot(subset(p,!delay),aes(x=R,y=2*X/R,col=z,linetype=u,group=grp))+
+  geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()
+plot(base)
+
+base2 = ggplot(subset(p,delay),aes(x=R,y=2*X/R,col=z,linetype=u,group=grp))+
+  geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()
+plot(base2)

### 2023-07-12 analysis/scripts/nationalModelTrajectories.R

@@ -14,8 +14,8 @@ baseDir <- "."
 johnsonCompare <- read.csv(paste0(baseDir,"/data/Johnson et al. figures5_6.csv"))
 
 # get outputs from demographics
-covTableSim <- data.frame(Anthro = seq(0, 10, 20, 30, 40, 50, 
-                                     60, 70, 80, 90), 
+covTableSim <- data.frame(Anthro = c(0, 10, 20, 30, 40, 50,
+                                     60, 70, 80, 90),
                           Fire = 0,
                           fire_excl_anthro = 0)
 covTableSim$polygon <- paste0("Missisa_", covTableSim$Anthro + 1)
@@ -51,6 +51,28 @@ rateSummaries <- demographicRates(
   ignorePrecision = F, returnSample = F, useQuantiles = F
 )
 
+#for each sample, select bias correction term from distributions
+getBiasCorrection<-function(q,w,u,z){
+  c = w*(q*u+1-u)/((q*u+w-u)*(1-z))
+  return(c)
+}
+
+str(rateSamples)
+
+cr = data.frame(replicate=unique(rateSamples$replicate))
+nr= nrow(cr)
+cr$c = getBiasCorrection(q=runif(nr,0,0.6),w=runif(nr,3,9),u=runif(nr,0,0.2),z=runif(nr,0,0.2))
+rateSamples$c=NULL;rateSamples=merge(rateSamples,cr)
+
+str(rateSamplesLarge)
+cr = data.frame(replicate=unique(rateSamplesLarge$replicate))
+nr= nrow(cr)
+cr$c = getBiasCorrection(q=runif(nr,0,0.6),w=runif(nr,3,9),u=runif(nr,0,0.2),z=runif(nr,0,0.2))
+rateSamplesLarge$c=NULL;rateSamplesLarge=merge(rateSamplesLarge,cr)
+
+str(rateSamplesLarge)
+rateSummaries$c = mean(rateSamplesLarge$c)
+
 rateSamples$S_PIlow <- 1
 rateSamples$S_PIhigh <- 1
 rateSamples$rep <- as.factor(rateSamples$replicate)
@@ -66,7 +88,6 @@ rateSamplesLarge$R_PIlow <- 1
 rateSamplesLarge$R_PIhigh <- 1
 rateSamplesLarge$fullGrp <- paste(rateSamplesLarge$rep) # ,rateSamplesLarge$Fire)
 
-str(johnsonCompare)
 johnsonCompare$Anthro=johnsonCompare$anthro
[…]

### 2023-07-18 analysis/paper/paper.Rmd
In progress - methods writeup and composition bias

@@
 
-## Integration of local demographic data and national disturbance-demographic relationships in Bayesian population model
+## Integration of local demographic data and national disturbance-demographic relationships in a Bayesian population model
 
@@
 We model recruitment probability as a function of anthropogenic disturbance and fire with a log link [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a A_t+\beta^R_f F_t+\epsilon^R_t; \epsilon^R_t \sim \text{Normal}(0,\sigma^2_{R}).$$
-Estimated recruitment probability is adjusted for sex ratio to get expected recruits per female: $X_t=R_t/2$.
-We include an option to also adjust for delayed age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012], but to maintain consistency with Johnson et al.
-[-@johnson_science_2020] we do not apply this adjustment by default: $$X_t=\frac{R_t/2}{1+R_t/2}.$$
+Estimated recruitment probability is adjusted for sex ratio and delayed age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012] to get expected recruits per female:
+$$X_t=\frac{R_t/2}{1+R_t/2}.$$
+To maintain consistency with Johnson et al. [-@johnson_science_2020] also provide an option to only adjust for sex ratio $X_t=R_t/2$.
 
@@
 ### Population dynamics
-To simulate plausible plausible demographic trajectories for example populations we used the slightly modified version of Johnson et al's [-@johnson_science_2020] two-stage demographic model described by Dyson et al.
+To simulate plausible plausible demographic trajectories for example populations we used a slightly modified version of Johnson et al's [-@johnson_science_2020] two-stage demographic model described by Dyson et al.
 [-@dyson_existing_2022].
 The true number of post-juvenile females that survive from year $t$ to the census $\dot{W}_t$ is binomially distributed with true survival probability $\dot{S}_t$: $\dot{W}_{t} \sim \text{Binomial}(\dot{N}_t,\dot{S}_t)$.
-As in the IPM, maximum potential recruitment rate is adjusted for sex ratio, misidentification biases (Appendix X) and (optionally) delayed age at first reproduction $$\dot{X}_t=\frac{c\dot{R}_t/2}{1+c\dot{R}_t/2};c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}, $$
+Maximum potential recruitment rate is adjusted for sex ratio, misidentification biases (Appendix X) and (optionally) delayed age at first reproduction $$\dot{X}_t=\frac{c\dot{R}_t/2}{1+c\dot{R}_t/2};c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}, $$
 where $w$ is a multiplier that defines the apparent number of adult females in the composition survey as a function of the number of collared animals, $q$ the ratio of young bulls to adult females, $u$ is the probability of misidentifying young bulls as adult females and vice versa, and $z$ is the probability of missing a calf.
@@
 
-### Observed survival from collars
-
-### Observed recruitment from calf:cow aerial surveys
-
-## Assessing the value of information
-
-TO DO: how to assess or quantify value in this context. Power to distinguish increasing from decreasing population? What else?
-
-## Scenarios
-
-Table 1. Example ranges. Ontario demographic information from [@mnrf_state_2014]. If two numbers given for Anthropogenic disturbance they are BEAD 2010 and 2015.
-*note that we've had some input on monitoring numbers from NL, BC, ON (and MB pending), so some of them may also be willing to give us case study examples to use here?
-Note can get details about SK1 from Phil's report.
-
-```{r}
-tbl <- "Range|Anthropogenic Disturbance|$\\lambda$|R|S|Survey Years
-Missisa|0.41|0.86|14.2|80%|2008-2012
-Churchill||0.96|19.5|87%|2011-2012
-"
-df <- read.delim(textConnection(tbl),header=FALSE,sep="|",strip.white=TRUE,stringsAsFactors=FALSE)
-names(df) <- unname(as.list(df[1,]))
-df <- df[-1,] # remove first row
-row.names(df)<-NULL
-knitr::kable(df)
-```
-TO DO: id examples from elsewhere to complement Ontario examples. Would be good to include an example where status is ok. What else, and why? WBI NWT ranges?
-*from Anna: withouot access to actual estimates of R and S for different ranges, some potentially interesting other ranges to compare to could be an Alberta range or two, where there are ++years of R and S estimates but really low lambda (E.g.
-ESAR and WSAR have 27+ yrs consecutive monitoring but declining numbers and high disturbance), some sections of NWT1 that are thought to be self-sustaining and have pretty good data sets (e.g.
-Dehcho North and South, Hay River Lowlands with 13+yrs of S and R) -could also use other MB ranges as examples for fairly sparse data (though not as bad as ON), and QC or LAB ranges as examples of good data (e.g.
-how about Charlevoix for a very well monitored but tiny pop, northern QC for really broad-expanse pop, or Labrador pops have long term monitoring data but uncertain status...)
-
-NOTE: would be great to info on number of collars per year in each example range. Though note MNRF recruitment estimates in some years came from aerial distribution surveys.
-*feedback from MonitoringWG contacts includes details on collar numbers for BC, ON, NL, (MB pending); e.g.
-in Ontario they aim for 20-30 collars in smaller pops, 35+ in bigger pops; Labrador aims to have between 5-25 collars monitored at any one time, depending on pop size; in BC, real numbers of collared animals have ranged been between 3-59 per pop/year, varying both across-yrs-within-pop and across-pop-within-yr [with COVID yrs by far the lowest].
-
-relevant refs: [@rettie_summary_2017];?
+## Monitoring scenarios
 
@@
 
+We assume collars are deployed in January, and last for up to 6 years.
+The target number of collared females $n$ varies among scenarios, as does the total duration of the monitoring program (Table \@ref(tab:monitoringScenarios)).
+Collars lost to mortality may be replaced each year ($o=1$), or there may be a 3 year gap between deployments ($o=4$).
+
 *Monitoring WG input from BC, ON, NL, (MB pending): lots of detail provided but generally those numbers suggest we could try values of ~3-10 [e.g.
@@
 
-TO DO: figure out disturbance forecast scenarios.
-Can lift Mississa scenario from caribouMetrics paper.
-Get Churchill scenario from Colin's Lac Seul model with roads - can lift it from the Churchill full model if not easy to sub in Lac Seul.
-*and can maybe ask MonitoringWG contacts who seemed keen whether they'd have case studies to offer too...?
+### Observed survival of collared animals
+
+In our simulated populations, mortality risk does not vary over time or among individuals, so the observed state (dead or alive) of a collared animal $i$ in month $m$ and year $t$ depends on the true survival rate $\dot{S}_t$ (eq X) as follows:
+$$\hat{I}_{i,t,m} \sim \text{Bernoulli}(\dot{S}^{1/12}_t\hat{I}_{i,t,m-1}).$$
+### Observed recruitment from composition surveys
+
+The number of cows in the composition survey $\hat{W}_t$ is given by the number of collared cows $\hat{T}_t$ and the apparent number of adult females per collared female calves observed in the composition survey $w$:
+$$\hat{W}_t = w\hat{T}_t.$$
+The number of observed calves $\hat{J}_t$ also depends on the unadjusted apparent recruitment for the population $\dot{R}_t$:
+$$\hat{J}_t \sim \text{Binomial}(\hat{W}_t,\dot{R}_t).$$
+
+## Disturbance scenarios
+
+*from Anna: withouot access to actual estimates of R and S for different ranges, some potentially interesting other ranges to compare to could be an Alberta range or two, where there are ++years of R and S estimates but really low lambda (E.g.
+ESAR and WSAR have 27+ yrs consecutive monitoring but declining numbers and high disturbance), some sections of NWT1 that are thought to be self-sustaining and have pretty good data sets (e.g.
+Dehcho North and South, Hay River Lowlands with 13+yrs of S and R) -could also use other MB ranges as examples for fairly sparse data (though not as bad as ON), and QC or LAB ranges as examples of good data (e.g.
+how about Charlevoix for a very well monitored but tiny pop, northern QC for really broad-expanse pop, or Labrador pops have long term monitoring data but uncertain status...)
+
+*feedback from MonitoringWG contacts includes details on collar numbers for BC, ON, NL, (MB pending); e.g.
+in Ontario they aim for 20-30 collars in smaller pops, 35+ in bigger pops; Labrador aims to have between 5-25 collars monitored at any one time, depending on pop size; in BC, real numbers of collared animals have ranged been between 3-59 per pop/year, varying both across-yrs-within-pop and across-pop-within-yr [with COVID yrs by far the lowest].
+
+relevant refs: [@rettie_summary_2017];?
 
@@
 
-TO DO: run low initial population size scenario, put in appendix of one by one sensitivity checks.
-
-TO DO: calculate change in anthro disturbance over 5 yrs across ranges to get realistic distribution of numbers of this & justify choice of parameters.
-run alternative scenario to put in one by one sensitivity appendix.
+TO DO: run low initial population size scenario, put in appendix of one by one sensitivity checks?
 
@@
 
-## Decision support tool development
+## Probability of correct status assessment
+
+## User interface and open, flexible, reuseable implementation
 
@@
 |$i$|ID|Individual animal identifier.|
-|$y$|years|Assessment period for population growth rate.|
-|$N_0$|animals|Initial adult population size.|
 |**Observed**|
@@
 |$\hat{W}_t$|animals|Cows observed in calf:cow survey.|
+|$\hat{T}_t$|animals|Number of collared cows.|
 |$\hat{S}_t$|probability|Kaplan-Meier estimate of observed adult female survival.|
@@
 |$\dot{J}_t$|animals|New recruits to adult age class.|
-|**True parameters**|
+|**True demographic parameters**|
+|$N_0$|animals|Initial adult population size.|
 |$\dot{\beta}^R_0$||Recruitment intercept.|
@@
 |$h_S=0$||Maximum value for $\dot{S}_t$.|
+|**Observation model parameters**|
+|$d=1,2,4,8,16, \text{ or } 24$|years|Monitoring duration.|
+|$n=15,30, \text{ or } 60$|animals|Target number of collared females.|
+|$o=1 \text{ or } 4$|years|Years between collar deployments.|
+|$y=3$|years|Assessment period for population growth rate.|
+|$w=3,6,\text{ or } 9$|ratio|The apparent number of adult females per collared animal in composition survey.|
+|$q=\text{uniform}(0,0.6)$|ratio|True ratio of young bulls to adult females in composition survey.|
+|$u=\text{uniform}(0,0.2)$|probability|Probability of misidentifying young bulls as adult females and vice versa in composition survey.|
+|$z=\text{uniform}(0,0.2)$|probability|Probability of missing calves in composition survey.|
+|$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}$||Bias correction term (Appendix X).|
 
@@
 
-We assume each group of animals in a calf:cow survey contains one or more collared adult females ($T$), and may also include: uncollared adult females misidentified as young bulls ($U$); correctly identified uncollared adult females ($V$); correctly identified young bulls ($O$); young bulls misidentified as uncollared adult females ($P$); observed calves ($J$); and unobserved calves ($K$).
+We assume each group of animals in a calf:cow survey contains one or more collared adult females ($T$), and may also include: uncollared adult females misidentified as young bulls or unknown sex ($U$); correctly identified uncollared adult females ($V$); young bulls correctly identified as male or unknown sex ($O$); young bulls misidentified as uncollared adult females ($P$); observed calves ($J$); and unobserved calves ($K$).
 The apparent number of adult females in the group is $T+V+P=Tw$, where $w$ is a multiplier that defines the apparent number of adult females as a function of the number of collared animals.
@@
 
-If we also choose to adjust for delayed age at first reproduction (), the adjusted recruitment rate becomes $$X=\frac{cR/2}{1+cR/2}.$$
+If we also adjust for delayed age at first reproduction (DeCesare et al. 2012; Eacker et al. 2019), the adjusted recruitment rate becomes $$X=\frac{cR/2}{1+cR/2}.$$
 
@@
 When uncollared animals are counted ($w>1$), the risk of misidentifying adults $u$ inflates the apparent recruitment rate $R$ as on average more cows are misidentified as bulls than vice versa, but this effect is smaller than the effect of missing calves $z$.
-Increasing the proportion of bulls in the group $q$ reduces the effect of adult misidentification $u$.", cache=TRUE}
+Increasing the proportion of bulls in the group $q$ reduces the effect of adult misidentification $u$.", fig.height=6,cache=TRUE}
 library(ggplot2)
@@
   geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()+ylab("2X/R=c, without delayed reproduction")
-plot(base)
 
@@
   geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()+ylab("2X/R, with delayed reproduction")
-plot(base2)
+
+ggpubr::ggarrange(base, base2, labels = "",
+                  ncol = 1, vjust = 1)
+
+```
+
+```{r biasDistribution, fig.cap="In the Bayesian model, uncertainty about the value of the bias correction term $c$ (grey bars) is approximated with a Gaussian distribution (red).
+In these examples, the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2.
+We assume that the apparent number of adult females per collared animal $w$ is known.", cache=TRUE}
+getBiasCorrection<-function(q,w,u,z){
+  c = w*(q*u+1-u)/((q*u+w-u)*(1-z))
+  return(c)
+}
+
+cr=expand.grid(w = c(1,2,3,4,5,6,7,8,9),rep=seq(1,10000))
+nr=nrow(cr)
+cr$c = getBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2))
+
+#get mean and sd as fn of w
+cs <- cr %>%
+  group_by(w) %>%
+  summarise(mean = mean(c), sd = sd(c))
+
+cr = merge(cr,cs)
+cr$ca= rnorm(nrow(cr),cr$mean,cr$sd)
+
+base = ggplot(cr,aes(x=c))+geom_histogram(aes(y=..density..))+
+  facet_wrap(~w,nrow=3,labeller=label_both)+theme_bw()+
+  geom_density(aes(x=ca),alpha=.2, fill="#FF6666",col = "#FF6666")
+plot(base)
 

### 2023-07-18 analysis/scripts/formatScenarios.R
adding

@@ -5,16 +5,17 @@ rawScns<-read.csv(here::here("tabs/s5.csv"))
 nrow(rawScns)
 rawScns = subset(rawScns,collarCount!=1)
 
-monitoringPars = list(obsYears=list(description = "monitoring duration",unit="years"),
-                      collarCount=list(description="target # of collars", unit="cows"),
-                      cowMult=list(description="calf:cow survey multiplier",unit = "cows per collared cow"),
-                      collarInterval = list(description="time between collar deployments", unit="years"),
-                      assessmentYrs = list(description="assessment duration",unit="years"))
+monitoringPars = list(obsYears=list(symbol= "d", description = "Monitoring duration.",unit="years"),
+                      collarCount=list(symbol="n",description="Target # of collars.", unit="cows"),
+                      collarInterval = list(symbol = "o",description="Years between collar deployments.", unit="years"),
+                      assessmentYrs = list(symbol="y",description="Assessment period for population growth rate.",unit="years"),
+                      cowMult=list(symbol="w",description="Cows per collared animal in composition survey.",unit = "ratio")
+                      )
 
 start=T
 for(par in names(monitoringPars)){
   #par="st"
-  crow = data.frame(name=par,description=monitoringPars[[par]]$description,
+  crow = data.frame(name=paste(monitoringPars[[par]]$symbol,par,sep=", "),description=monitoringPars[[par]]$description,
                     units=monitoringPars[[par]]$unit,
                     values = paste(unique(rawScns[,par]),collapse=","))
 

### 2023-07-18 analysis/scripts/nationalModelTrajectories.R

@@ -10,6 +10,12 @@ dia_shp <- 23
 err_col <- "grey50"
 
 baseDir <- "."
+
+getBiasCorrection<-function(w,q,u,z){
+  c = w*(q*u+1-u)/((q*u+w-u)*(1-z))
+  return(c)
+}
+
 #numbers from Johnson et al for validation
 johnsonCompare <- read.csv(paste0(baseDir,"/data/Johnson et al. figures5_6.csv"))
 
@@ -52,12 +58,8 @@ rateSummaries <- demographicRates(
 )
 
 #for each sample, select bias correction term from distributions
-getBiasCorrection<-function(q,w,u,z){
-  c = w*(q*u+1-u)/((q*u+w-u)*(1-z))
-  return(c)
-}
 
-str(rateSamples)
+str(popGrowthPars)
 
 cr = data.frame(replicate=unique(rateSamples$replicate))
 nr= nrow(cr)

### 2023-07-18 analysis/scripts/sensitivityMinimalTest.R

@@ -4,21 +4,24 @@ n_reps <- "all"
 
 # Run batches from Rscript that uses parallel backend and new caribouMetrics functions
 cpageId <- commandArgs(trailingOnly = TRUE)
-# cpageId <- 5
+
+cDir = getwd()
 
 library(caribouMetrics)
+
+# cpageId <- 5;n_reps <- 1;cDir = "C:/Users/HughesJo/Documents/gitprojects/Caribou-Demographic-Projection-Paper"
 
 setName = "s5"
 
 #######################
-dir.create(paste0("figs/",setName),recursive=T)
-dir.create(paste0("tabs/",setName),recursive=T)
-dir.create(paste0("results/",setName),recursive=T)
+dir.create(paste0(cDir,"/figs/",setName),recursive=T)
+dir.create(paste0(cDir,"/tabs/",setName),recursive=T)
+dir.create(paste0(cDir,"/results/",setName),recursive=T)
 
 
 simBig<-getSimsNational() #If called with default parameters, use saved object to speed things up.
 
-allScns = read.csv(paste0("tabs/",setName,".csv"))
+allScns = read.csv(paste0(cDir,"/tabs/",setName,".csv"))
 unique(allScns$pageId)
 ####################
 eParsIn = list()
@@ -38,9 +41,10 @@ message("batch ", cpageId, " started")
 if(n_reps=="all"){
   scResults = caribouMetrics:::runScnSet(scns,eParsIn,simBig,getKSDists=F,printProgress=F)
 }else{
-  scResults = caribouMetrics:::runScnSet(subset(scns,rep<=n_reps),eParsIn,simBig,getKSDists=F,printProgress=F)
+  unique(scns$rep)
+  scResults = caribouMetrics:::runScnSet(subset(scns,rep<=n_reps)[1,],eParsIn,simBig,getKSDists=F,printProgress=F)
 }
 
-saveRDS(scResults,paste0("results/",setName,"/rTest",cpageId,".Rds"))
+saveRDS(scResults,paste0("results/",setName,"/rTest",cpageId,n_reps,cpageId,".Rds"))
 
 message("batch ", cpageId, " complete")

### 2023-07-19 analysis/paper/paper.Rmd
Bias adjustement and methods writeup continued.

@@
 We model recruitment probability as a function of anthropogenic disturbance and fire with a log link [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a A_t+\beta^R_f F_t+\epsilon^R_t; \epsilon^R_t \sim \text{Normal}(0,\sigma^2_{R}).$$
-Estimated recruitment probability is adjusted for sex ratio and delayed age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012] to get expected recruits per female:
-$$X_t=\frac{R_t/2}{1+R_t/2}.$$
-To maintain consistency with Johnson et al. [-@johnson_science_2020] also provide an option to only adjust for sex ratio $X_t=R_t/2$.
+Estimated recruitment probability is adjusted for sex ratio, misidentification biases (Appendix X), and (optionally) age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012] to get expected recruits per female:
+$$X_t=\frac{cR_t/2}{1+cR_t/2}.$$
+To maintain consistency with Johnson et al. [-@johnson_science_2020] we also provide an option to only adjust for sex ratio and misidentification biases $X_t=cR_t/2$.
+
+The composition survey bias correction term $c$ depends on the apparent number of adult females per collared animal $w$, the ratio of young bulls to adult females $q$, the probability of misidentifying young bulls as adult females and vice versa $u$, and the probability of missing a calf $z$ (Appendix X):
+$$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}$$
+In the Bayesian model, prior uncertainty about the value of the bias correction term $c$ is approximated with a Gaussian distribution (Appendix X).
+Given the apparent number of adult females per collared animal $w$ in each monitoring scenario (Table \@ref(tab:monitoringScenarios)), we calculate the mean and standard deviation of 10,000 samples of $c$ assuming the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2.
 
@@
 All regression coefficients are assumed to be Gaussian distributed, with priors derived from the expected values and 95% confidence intervals estimated by Johnson et al (2020) (Table \@ref(tab:priors)).
-We calibrated the prior distributions so that the 95% prediction interval from the IPM matches the range between the 2.5% and 97.5% quantiles of 1000 simulated recruitment (Appendix A).
+We calibrated the prior distributions so that the 95% prediction interval from the Bayesian model matches the range between the 2.5% and 97.5% quantiles of 1000 simulated recruitment (Appendix A).
 
@@
 # Appendix
-##Definitions of variables and parameters in the Bayesian integrated population model
+##Definitions of variables and parameters in the Bayesian population model
 
-Table A1. Variables and parameters in the Bayesian integrated population model and the national demographic model.
+Table A1. Variables and parameters in the Bayesian population model and the national demographic model.
 
@@
 
-We assume each group of animals in a calf:cow survey contains one or more collared adult females ($T$), and may also include: uncollared adult females misidentified as young bulls or unknown sex ($U$); correctly identified uncollared adult females ($V$); young bulls correctly identified as male or unknown sex ($O$); young bulls misidentified as uncollared adult females ($P$); observed calves ($J$); and unobserved calves ($K$).
+We assume each group of animals in a calf:cow composition survey contains one or more collared adult females ($T$), and may also include: uncollared adult females misidentified as young bulls or unknown sex ($U$); correctly identified uncollared adult females ($V$); young bulls correctly identified as male or unknown sex ($O$); young bulls misidentified as uncollared adult females ($P$); observed calves ($J$); and unobserved calves ($K$).
 The apparent number of adult females in the group is $T+V+P=Tw$, where $w$ is a multiplier that defines the apparent number of adult females as a function of the number of collared animals.
@@
 getX<-function(R,delay,q,w,u,z){
-  X = 0.5*R*w*(q*u+1-u)/((q*u+w-u)*(1-z))
+  c = compositionBiasCorrection(w,q,u,z)
+  X = 0.5*R*c
   X=X*(1-delay)+delay*X/(1+X)
@@
 We assume that the apparent number of adult females per collared animal $w$ is known.", cache=TRUE}
-getBiasCorrection<-function(q,w,u,z){
-  c = w*(q*u+1-u)/((q*u+w-u)*(1-z))
-  return(c)
-}
 
@@
 nr=nrow(cr)
-cr$c = getBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2))
+cr$c = compositionBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2))
 

### 2023-07-19 analysis/scripts/sensitivity1.R
Setting up new result set 6, with bias adjustment and delayed reproduction.

@@ -16,7 +16,7 @@
 #monitoringScns = expand.grid(P=c(1,4,8,16),st=c(30,60),cmult=c(3),ri=c(1,2,4),assessmentYrs=1)
 #stateScns = expand.grid(iA=c(0),rep=seq(1:30))
 
-setName = "s5"
+setName = "s6"
 monitoringScns = expand.grid(obsYears=c(1,2,4,8,16,24),collarCount=c(1,15,30,60),cowMult=c(3,6,9),collarInterval=c(1,4),assessmentYrs=c(3))
 #TO DO - in next iteration, remove multiple years of one collar
 monitoringScns = subset(monitoringScns, !((obsYears>1)&(collarCount==1))&!((collarCount==1)&(cowMult>3))&!((collarCount==1)&(collarInterval>1)))
@@ -34,6 +34,7 @@ unique(scns$iAnthro)
 scns$repBatch = ceiling(scns$rep/50)
 table(scns$repBatch)
 scns$N0 = 1000
+scns$adjustR = TRUE
 #scns=scns[10,]
 #scResults = readRDS("temp.Rds")
 head(scns)

### 2023-07-20 analysis/scripts/sensitivityMinimal.R
update to set 6

@@ -6,7 +6,7 @@ cpageId <- commandArgs(trailingOnly = TRUE)
 #setwd("C:/Users/HughesJo/Documents/gitprojects/Caribou-Demographic-Projection-Paper")
 library(caribouMetrics)
 
-setName = "s5"
+setName = "s6"
 
 #######################
 dir.create(paste0("figs/",setName),recursive=T)

### 2023-07-20 analysis/scripts/sensitivityMinimalTest.R

@@ -1,6 +1,6 @@
 #!/usr/bin/env Rscript
 #nohup Rscript --vanilla "analysis/scripts/sensitivityMinimalTest.R" 1 &
-n_reps <- "all"
+n_reps <- 1
 
 # Run batches from Rscript that uses parallel backend and new caribouMetrics functions
 cpageId <- commandArgs(trailingOnly = TRUE)
@@ -11,7 +11,7 @@ library(caribouMetrics)
 
 # cpageId <- 5;n_reps <- 1;cDir = "C:/Users/HughesJo/Documents/gitprojects/Caribou-Demographic-Projection-Paper"
 
-setName = "s5"
+setName = "s6"
 
 #######################
 dir.create(paste0(cDir,"/figs/",setName),recursive=T)

### 2023-07-20 make.R

@@ -24,8 +24,7 @@ message("install caribouMetrics from GitHub")
 
 # do deps separately so they can be installed from binaries
 # only do download once
-cm_download <- remotes::remote_download(remotes::github_remote("LandSciTech/caribouMetrics",
-                                                               ref = "add-bayes-demog"))
+cm_download <- remotes::remote_download(remotes::github_remote("LandSciTech/caribouMetrics"))
 
 try(remotes::install_deps(cm_download, upgrade = "never"))
 # this should do ones that were not available from binary

### 2023-07-20 cloud/make_batch_scripts.R
use stored credentials

@@ -1,10 +1,22 @@
-# write a version of run_caribou.sh for each batch
+# Uses secrets stored in the local credential store with the keyring package
+if(0){
+  # prompts you to enter the secret
+  keyring::key_set("Azure_SASURL")
+  keyring::key_set("Azure_subnetId")
+  # retrieves the secret
+  keyring::key_get("Azure_SASURL")
+}
+
+
+
+# write a version of run_caribou.sh and task json for each batch
 
 make_files <- function(batch){
   batch <- as.character(batch)
   # need this to make it have unix line endings
   outfile <-  file(paste0("cloud/run_caribou", batch, ".sh"), "wb")
   readLines("cloud/run_caribou.sh") |> stringr::str_replace_all("<batch>", batch)|>
+    stringr::str_replace_all("<PAT>", gh::gh_token()) |>
     writeLines(con = outfile)
   close(outfile)
 }
@@ -17,7 +29,16 @@ make_task <- function(batch){
   # need this to make it have unix line endings
   outfile <-  file(paste0("cloud/caribouDemo", batch, ".json"), "wb")
   readLines("cloud/caribouDemo.json") |> stringr::str_replace_all("<batch>", batch)|>
+    stringr::str_replace_all("<SASURL>", keyring::key_get("Azure_SASURL"))|>
     writeLines(con = outfile)
   close(outfile)
 }
 purrr::walk(1:24, make_task)
+
+# Add secrets to pool
+# need this to make it have unix line endings
+outfile <-  file(paste0("cloud/caribou_add_pool1.json"), "wb")
+readLines("cloud/caribou_add_pool.json") |>
+  stringr::str_replace_all("<subnetId>", keyring::key_get("Azure_subnetId"))|>
+  writeLines(con = outfile)
+close(outfile)

### 2023-07-24 analysis/paper/paper.Rmd
in progress - bias term.

@@
 $$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}$$
-In the Bayesian model, prior uncertainty about the value of the bias correction term $c$ is approximated with a Gaussian distribution (Appendix X).
-Given the apparent number of adult females per collared animal $w$ in each monitoring scenario (Table \@ref(tab:monitoringScenarios)), we calculate the mean and standard deviation of 10,000 samples of $c$ assuming the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2.
+In the Bayesian model, prior uncertainty about the value of the bias correction term $c$ is approximated with a Log-normal distribution (Appendix X).
+Given the apparent number of adult females per collared animal $w$ in each monitoring scenario (Table \@ref(tab:monitoringScenarios)), we calculate the mean and standard deviation of 10,000 samples of $\log{c}$ assuming the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2.
 
@@
 
-```{r biasDistribution, fig.cap="In the Bayesian model, uncertainty about the value of the bias correction term $c$ (grey bars) is approximated with a Gaussian distribution (red).
+```{r biasDistribution, fig.cap="In the Bayesian model, uncertainty about the value of the bias correction term $c$ (grey bars) is approximated with a Log-normal distribution (red).
 In these examples, the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2.
@@
   group_by(w) %>%
-  summarise(mean = mean(c), sd = sd(c))
+  summarise(mean = mean(log(c)), sd = sd(log(c)))
 
 cr = merge(cr,cs)
-cr$ca= rnorm(nrow(cr),cr$mean,cr$sd)
+cr$ca= exp(rnorm(nrow(cr),cr$mean,cr$sd))
 

### 2023-07-24 analysis/scripts/BiasModelSensitivity.R

@@ -1,4 +1,6 @@
 library(ggplot2)
+library(tidyverse)
+library(caribouMetrics)
 #Notes
 #q: assume more females than males in calf:cow group, so q<1
 
@@ -25,3 +27,41 @@ plot(base)
 base2 = ggplot(subset(p,delay),aes(x=R,y=2*X/R,col=z,linetype=u,group=grp))+
   geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()
 plot(base2)
+
+#############
+#approximate distribution for Bayesian model - lognormal
+cr=expand.grid(w = c(2,3,4,5,6,7,8,9),rep=seq(1,10000))
+nr=nrow(cr)
+cr$c = compositionBiasCorrection(q=runif(nr,0,0),w=cr$w,u=runif(nr,0,0),z=runif(nr,0.5,0.5))
+
+#get mean and sd as fn of w
+cs <- cr %>%
+  group_by(w) %>%
+  summarise(m = mean(c), v = var(c))
+cs$v[cs$v==0]=0.000001
+
+#now find lognormal parameters from mean and variance
+#https://www.johndcook.com/blog/2022/02/24/find-log-normal-parameters/
+cs$sig2 = log(1+cs$v/cs$m^2)
+cs$mu = log(cs$m)-cs$sig2/2
+
+cr = merge(cr,cs)
+cr$ca= exp(rnorm(nrow(cr),mean=cr$mu,sd=cr$sig2^0.5))
+
+base = ggplot(cr,aes(x=c))+geom_histogram(aes(y=..density..))+
+  facet_wrap(~w,nrow=3,labeller=label_both)+theme_bw()+
+  geom_density(aes(x=ca),alpha=.2, fill="#FF6666",col = "#FF6666")
+plot(base)
+
+
+mns <- cr %>%
+  group_by(w) %>%
+  summarise(c_mean = mean(c), c_sd = sd(c),ca_mean=mean(ca),ca_sd=sd(ca))
+
+base2 = ggplot(mns,aes(x=c_mean,y=ca_mean))+geom_point()+geom_abline(slope=1,intercept=0)
+plot(base2)
+#consider high and low bounds. When q and u = 1, c= 1/(1-z)
+z = seq(0,0.99,length.out=100)
+c = 1/(1-z)
+plot(log(c)~z)
+

### 2023-07-24 analysis/scripts/sensitivityMinimalTest.R

@@ -36,7 +36,13 @@ eParsIn$collarNumYears=6
 
 scns = subset(allScns, pageId==cpageId)
 
+str(scns)
+
+scns$zMin= 0.5; scns$zMax = 0.5
+scns$uMin = 0; scns$uMax = 0
+scns$qMin = 0; scns$qMax = 0
 message("batch ", cpageId, " started")
+
 
 if(n_reps=="all"){
   scResults = caribouMetrics:::runScnSet(scns,eParsIn,simBig,getKSDists=F,printProgress=F)

### 2023-07-26 analysis/paper/paper.Rmd
In progress - writeup of methods and results.

@@
 
+## Anthropogenic disturbance and monitoring scenarios
+
+```{r monitoringScenarios, echo=FALSE,message=FALSE}
+
+source(here::here("analysis/scripts/formatScenarios.R"))
+
+knitr::kable(tbl,caption="Monitoring scenarios parameters and values.")
+```
+
+```{r figDisturbance, fig.cap="Anthropogenic disturbance and monitoring scenarios.", cache=TRUE}
+knitr::include_graphics(here::here("figs/forPaper/distScnscmult6ay3aSf1.png"))
+
+```
+
+For this investigation we assume collars are deployed in January, and last for up to 6 years.
+The target number of collared females $n$ varies among scenarios, as does the total duration of the monitoring program (Table \@ref(tab:monitoringScenarios)).
+Collars lost to mortality may be replaced each year ($o=1$), or there may be a 3 year gap between deployments ($o=4$).
+
+*Monitoring WG input from BC, ON, NL, (MB pending): lots of detail provided but generally those numbers suggest we could try values of ~3-10 [e.g.
+3,6,9?] for the cow multiplier, 5-50 for collars [e.g.
+5,10,20,40,60?], [not sure about renewal interval, but see SaraM email for some insights].
+We know (e.g.
+from the MonitoringWG perspectives report/update to Rettie summary) that number of years of monitoring can be as low as 2 or as many as 20+
+
+*from Anna: withouot access to actual estimates of R and S for different ranges, some potentially interesting other ranges to compare to could be an Alberta range or two, where there are ++years of R and S estimates but really low lambda (E.g.
+ESAR and WSAR have 27+ yrs consecutive monitoring but declining numbers and high disturbance), some sections of NWT1 that are thought to be self-sustaining and have pretty good data sets (e.g.
+Dehcho North and South, Hay River Lowlands with 13+yrs of S and R) -could also use other MB ranges as examples for fairly sparse data (though not as bad as ON), and QC or LAB ranges as examples of good data (e.g.
+how about Charlevoix for a very well monitored but tiny pop, northern QC for really broad-expanse pop, or Labrador pops have long term monitoring data but uncertain status...)
+
+*feedback from MonitoringWG contacts includes details on collar numbers for BC, ON, NL, (MB pending); e.g.
+in Ontario they aim for 20-30 collars in smaller pops, 35+ in bigger pops; Labrador aims to have between 5-25 collars monitored at any one time, depending on pop size; in BC, real numbers of collared animals have ranged been between 3-59 per pop/year, varying both across-yrs-within-pop and across-pop-within-yr [with COVID yrs by far the lowest].
+
+relevant refs: [@rettie_summary_2017];?
+
+TO DO: Fire scenarios? Add high fire scenario in appendix of one by one sensitivity checks.
+
+TO DO: run low initial population size scenario, put in appendix of one by one sensitivity checks?
+
+## Simulation of local population dynamics and monitoring
+
+### Population dynamics informed by national disturbance-demographic relationships
+
+```{r exampleTrajectories, fig.cap="Examples of simulated population trajectories from a model informed only by national disturbance-demographic relationships with (top row) and without (bottom row) delayed reproduction.
+White dots show means of 500 example trajectories.
+For reference, black lines and and shaded bands show means and 95\\% predictive intervals from Johnson et al's (2020) regression models without adjustment for misidentification bias and delayed reproduction.", cache=TRUE}
+
+knitr::include_graphics(here::here("analysis/paper/figs/DemographicRatesDelay.png"))
+
+```
+
+To simulate plausible plausible demographic trajectories for example populations we used a modified version of Johnson et al's [-@johnson_science_2020] two-stage demographic model described by Dyson et al.
+[-@dyson_existing_2022].
+The true number of post-juvenile females that survive from year $t$ to the census $\dot{W}_t$ is binomially distributed with true survival probability $\dot{S}_t$: $\dot{W}_{t} \sim \text{Binomial}(\dot{N}_t,\dot{S}_t)$.
+Maximum potential recruitment rate is adjusted for sex ratio, misidentification biases (Appendix X) and (optionally) delayed age at first reproduction $$\dot{X}_t=\frac{c\dot{R}_t/2}{1+c\dot{R}_t/2};c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}, $$
+where $w$ is a multiplier that defines the apparent number of adult females in the composition survey as a function of the number of collared animals, $q$ the ratio of young bulls to adult females, $u$ is the probability of misidentifying young bulls as adult females and vice versa, and $z$ is the probability of missing a calf.
+
+Realized recruitment rate varies with population density [@lacy_vortex_2017], and the number of juveniles recruiting to the post-juvenile class at the census is a binomially distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate: $$\dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{N_0k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}).$$ Given parameters in Table X, recruitment rate is lowest $(0.5\dot{X}_t)$ when $\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$ at intermediate population sizes, and declines to $0.6\dot{X}_t$ as the population reaches carrying capacity of $k=100$ times the initial population size.
+The post-juvenile female population in the next year includes both survivors and new recruits: $\dot{N}_{t+1}=\text{min}(\dot{W}_t+\dot{J}_t,r_{max}\dot{N}_t)$.
+
+Interannual variation in survival and recruitment is modelled using truncated beta distributions [rtrunc function; @novomestky_package_2016]: $\dot{R}_t \sim \text{TruncatedBeta}(\bar{R}_t,\nu_R,l_R,h_R); \dot{S}_t \sim \text{TruncatedBeta}(\bar{S}_t,\nu_S,l_S,h_S)$.
+Coefficients of variation among years $(\nu_R,\nu_S)$ and maximum/minimum values $l_R,h_R,l_S,h_S$ for recruitment and survival are given in Table X.
+Expected recruitment ($\bar{R}_t$) and survival ($\bar{S}_t$) vary with disturbance according the beta regression models estimated by Johnson et al.
+[@johnson_science_2020]: $$ \bar{R}_t \sim Beta(\mu^R_t,\phi^R);log(\mu^R_t)=\dot{\beta^R_0}+\dot{\beta^R_a}A_t+\dot{\beta}^R_fF_t,$$
+$$\bar{S}_t \sim (46\times Beta(\mu^S_t,\phi^S)-0.5)/45;log(\mu^S_t)=\dot{\beta^S_0}+\dot{\beta^S_a}A_t.$$
+$\phi^R \sim \text{Normal}(19.862,2.229)$ and $\phi^S \sim \text{Normal}(63.733,8.311)$ are precisions of the Beta distributed errors [@ferrari_beta_2004].
+At the beginning of a simulation for an example population, regression coefficient values are sampled from Gaussian distributions (see Table \@ref(tab:priors)) and the population is assigned to quantiles of the Beta error distributions for survival and recruitment.
+The population remains in these quantiles as disturbance changes over time, so there is substantial persistent variation in recruitment and survival among example among populations (Figure \@ref(fig:exampleTrajectories)).
+
+### Observed survival of collared animals
+
+In our simulated populations, mortality risk does not vary over time or among individuals, so the observed state (dead or alive) of a collared animal $i$ in month $m$ and year $t$ depends on the true survival rate $\dot{S}_t$ (eq X) as follows:
+$$\hat{I}_{i,t,m} \sim \text{Bernoulli}(\dot{S}^{1/12}_t\hat{I}_{i,t,m-1}).$$
+
+### Observed recruitment from composition surveys
+
+The number of cows in the composition survey $\hat{W}_t$ is given by the number of collared cows $\hat{T}_t$ and the apparent number of adult females per collared female calves observed in the composition survey $w$:
+$$\hat{W}_t = w\hat{T}_t.$$
+The number of observed calves $\hat{J}_t$ also depends on the unadjusted apparent recruitment for the population $\dot{R}_t$:
+$$\hat{J}_t \sim \text{Binomial}(\hat{W}_t,\dot{R}_t).$$
+
 ## Integration of local demographic data and national disturbance-demographic relationships in a Bayesian population model
@@
 All regression coefficients are assumed to be Gaussian distributed, with priors derived from the expected values and 95% confidence intervals estimated by Johnson et al (2020) (Table \@ref(tab:priors)).
-We calibrated the prior distributions so that the 95% prediction interval from the Bayesian model matches the range between the 2.5% and 97.5% quantiles of 1000 simulated recruitment (Appendix A).
+We calibrated the prior distributions so that the 95% prior prediction intervals for survival and recruitment from the Bayesian model match the range between the 2.5% and 97.5% quantiles of 1000 simulated survival and recruitment trajectories (Appendix A, Figure \@ref(fig:examplePredictions)).
 
@@
 #TO DO: indicate which of these are from Johnson, and which are calibrated.
+```
+
+```{r examplePredictions, fig.cap="Prior (top row) and posterior (bottom row) predictions from the Bayesian model for an example population trajectory.
+In this example scenario anthropogenic disturbance increases by 2\\% per year from 0 to 100\\%.
+The population is monitored for 15 years with 30 collars per year, and 6 cows per collared cow in the composition surveys.
+The prior means and 95\\% predictive intervals (top row, orange lines and shading) are similar to the means and ranges between the 2.5\\% and 97.5\\% quantiles of 3000 simulated survival and recruitment trajectories from the national model (top row, blue lines and shading).
+Local population data (open triangles) reduces uncertainty (bottom row), but when disturbance is very high local data is not required to predict viability (population growth < 1).", cache=TRUE}
+
+knitr::include_graphics(here::here("analysis/paper/figs/bayesianModelExamples.png"))
+
 ```
@@
 
-## A method for simulating example local demographic data
-
-### Population dynamics
-To simulate plausible plausible demographic trajectories for example populations we used a slightly modified version of Johnson et al's [-@johnson_science_2020] two-stage demographic model described by Dyson et al.
-[-@dyson_existing_2022].
-The true number of post-juvenile females that survive from year $t$ to the census $\dot{W}_t$ is binomially distributed with true survival probability $\dot{S}_t$: $\dot{W}_{t} \sim \text{Binomial}(\dot{N}_t,\dot{S}_t)$.
-Maximum potential recruitment rate is adjusted for sex ratio, misidentification biases (Appendix X) and (optionally) delayed age at first reproduction $$\dot{X}_t=\frac{c\dot{R}_t/2}{1+c\dot{R}_t/2};c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}, $$
-where $w$ is a multiplier that defines the apparent number of adult females in the composition survey as a function of the number of collared animals, $q$ the ratio of young bulls to adult females, $u$ is the probability of misidentifying young bulls as adult females and vice versa, and $z$ is the probability of missing a calf.
-
-Realized recruitment rate varies with population density [@lacy_vortex_2017], and the number of juveniles recruiting to the post-juvenile class at the census is a binomially distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate: $$\dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{N_0k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}).$$ Given parameters in Table X, recruitment rate is lowest $(0.5\dot{X}_t)$ when $\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$ at intermediate population sizes, and declines to $0.6\dot{X}_t$ as the population reaches carrying capacity of $k=100$ times the initial population size.
-The post-juvenile female population in the next year includes both survivors and new recruits: $\dot{N}_{t+1}=\text{min}(\dot{W}_t+\dot{J}_t,r_{max}\dot{N}_t)$.
-
-Interannual variation in survival and recruitment is modelled using truncated beta distributions [rtrunc function; @novomestky_package_2016]: $\dot{R}_t \sim \text{TruncatedBeta}(\bar{R}_t,\nu_R,l_R,h_R); \dot{S}_t \sim \text{TruncatedBeta}(\bar{S}_t,\nu_S,l_S,h_S)$.
-Coefficients of variation among years $(\nu_R,\nu_S)$ and maximum/minimum values $l_R,h_R,l_S,h_S$ for recruitment and survival are given in Table X.
-Expected recruitment ($\bar{R}_t$) and survival ($\bar{S}_t$) vary with disturbance according the beta regression models estimated by Johnson et al.
-[@johnson_science_2020]: $$ \bar{R}_t \sim Beta(\mu^R_t,\phi^R);log(\mu^R_t)=\dot{\beta^R_0}+\dot{\beta^R_a}A_t+\dot{\beta}^R_fF_t,$$
-$$\bar{S}_t \sim (46\times Beta(\mu^S_t,\phi^S)-0.5)/45;log(\mu^S_t)=\dot{\beta^S_0}+\dot{\beta^S_a}A_t.$$
-$\phi^R \sim \text{Normal}(19.862,2.229)$ and $\phi^S \sim \text{Normal}(63.733,8.311)$ are precisions of the Beta distributed errors [@ferrari_beta_2004].
-At the beginning of a simulation for an example population, regression coefficient values are sampled from Gaussian distributions (see Table \@ref(tab:priors)) and the population is assigned to quantiles of the Beta error distributions for survival and recruitment.
-The population remains in these quantiles as disturbance changes over time, so there is substantial persistent variation in recruitment and survival among example among populations (example figure).
-
-## Monitoring scenarios
-
-```{r monitoringScenarios, echo=FALSE,message=FALSE}
-
-source(here::here("analysis/scripts/formatScenarios.R"))
-
-knitr::kable(tbl,caption="Monitoring scenarios parameters and values.")
-```
-
-We assume collars are deployed in January, and last for up to 6 years.
-The target number of collared females $n$ varies among scenarios, as does the total duration of the monitoring program (Table \@ref(tab:monitoringScenarios)).
-Collars lost to mortality may be replaced each year ($o=1$), or there may be a 3 year gap between deployments ($o=4$).
-
-*Monitoring WG input from BC, ON, NL, (MB pending): lots of detail provided but generally those numbers suggest we could try values of ~3-10 [e.g.
-3,6,9?] for the cow multiplier, 5-50 for collars [e.g.
-5,10,20,40,60?], [not sure about renewal interval, but see SaraM email for some insights].
-We know (e.g.
-from the MonitoringWG perspectives report/update to Rettie summary) that number of years of monitoring can be as low as 2 or as many as 20+
-
-### Observed survival of collared animals
-
-In our simulated populations, mortality risk does not vary over time or among individuals, so the observed state (dead or alive) of a collared animal $i$ in month $m$ and year $t$ depends on the true survival rate $\dot{S}_t$ (eq X) as follows:
-$$\hat{I}_{i,t,m} \sim \text{Bernoulli}(\dot{S}^{1/12}_t\hat{I}_{i,t,m-1}).$$
-### Observed recruitment from composition surveys
-
-The number of cows in the composition survey $\hat{W}_t$ is given by the number of collared cows $\hat{T}_t$ and the apparent number of adult females per collared female calves observed in the composition survey $w$:
-$$\hat{W}_t = w\hat{T}_t.$$
-The number of observed calves $\hat{J}_t$ also depends on the unadjusted apparent recruitment for the population $\dot{R}_t$:
-$$\hat{J}_t \sim \text{Binomial}(\hat{W}_t,\dot{R}_t).$$
-
-## Disturbance scenarios
-
-*from Anna: withouot access to actual estimates of R and S for different ranges, some potentially interesting other ranges to compare to could be an Alberta range or two, where there are ++years of R and S estimates but really low lambda (E.g.
-ESAR and WSAR have 27+ yrs consecutive monitoring but declining numbers and high disturbance), some sections of NWT1 that are thought to be self-sustaining and have pretty good data sets (e.g.
-Dehcho North and South, Hay River Lowlands with 13+yrs of S and R) -could also use other MB ranges as examples for fairly sparse data (though not as bad as ON), and QC or LAB ranges as examples of good data (e.g.
-how about Charlevoix for a very well monitored but tiny pop, northern QC for really broad-expanse pop, or Labrador pops have long term monitoring data but uncertain status...)
-
-*feedback from MonitoringWG contacts includes details on collar numbers for BC, ON, NL, (MB pending); e.g.
-in Ontario they aim for 20-30 collars in smaller pops, 35+ in bigger pops; Labrador aims to have between 5-25 collars monitored at any one time, depending on pop size; in BC, real numbers of collared animals have ranged been between 3-59 per pop/year, varying both across-yrs-within-pop and across-pop-within-yr [with COVID yrs by far the lowest].
-
-relevant refs: [@rettie_summary_2017];?
-
-TO DO: Fire scenarios? Add high fire scenario in appendix of one by one sensitivity checks.
-
-TO DO: run low initial population size scenario, put in appendix of one by one sensitivity checks?
-
-```{r fig1, fig.cap="Anthropogenic disturbance and monitoring scenarios.", cache=TRUE}
-knitr::include_graphics(here::here("figs/forPaper/distScnscmult3ay3aSf1.png"))
-
-```
+TO DO: add example trajectory figure
 
@@
 
-TO DO: consider showing results in terms of average number of collars deployed per year rather than target # and renewal interval.
-
-```{r scn1, cache=TRUE,echo=FALSE, message=FALSE,results='hide'}
-if(runScenarios){
-  numObsYrs=c(2);startsByYr = 30 #25
-  scns=expand.grid(P=numObsYrs,sQ=c(0.5),st=startsByYr)
-  scResults = runScenario(scns)
-
-  png(here::here("figs/Figure1_Example.png"),
-    height = 4, width = fig_widths["mid"], units = "in",res=600)
-  print(plotRes(scResults$rr.summary.all, "Female population size",obs=scResults$obs.all,
-              lowBound=0,highBound=2000,facetVars=c("P","sQ")))
-  dev.off()
-}
-
-```
-
-```{r fig2, fig.cap="Example projections, assuming a target of 30 collars, renewed every 4 years, with 3 cows per collared cow included in aerial surveys, and a 1% increase in anthropogenic disturbance per year (as in Figure 1).
-The number of monitoring years varies from P=0 (top row) to P=24 (bottom row), and the anthropogenic disturbance footprint in 2023 varies from 0% (left column) to 60% (right column).
-A status assessment error occurs when the projected population growth rate is above one and the true rate is below one or vice versa.
-When disturbance is very high (e.g.
-Anthro2023 = 60) populations that differ substantially from one another (compare 'true' lines for high and low quantile populations) share the same status, and a projection informed only by national demographic disturbance relationships (P=0) is accurate for both.
-When disturbance is intermediate, populations that differ substantially from one another are more likely to differ in their status.
-In these cases, integrating local monitoring data improves projection accuracy (i.e.
-smaller distances between projected and true lines) and reduces status assessment errors.", cache=TRUE}
-knitr::include_graphics(here::here("figs/forPaper/examples.png"))
-```
-
-```{r fig3, fig.cap="Power analysis example result.", cache=TRUE}
-knitr::include_graphics(here::here("figs/forPaper/powercmult3ay3aSf1.png"))
-knitr::include_graphics(here::here("figs/forPaper/powerEffortcmult3ay3aSf1.png"))
-
-#TO DO: figure out how to put no monitoring results on these graphs.
+```{r figStatusErrors, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure 1 for scenarios).
+Probabilities are calculated for 200 example population trajectories (see example trajectories in Figure 2).
+For each example population, the status assessment is correct if the the Bayesian posterior mean and the true population growth rate are both less than or greater than one.
+When disturbance is very high the probability of correct status assessment is high even without local monitoring.
+Where disturbance is moderate, increasing the number of years of monitoring and the number of collars reduces uncertainty, but there are diminishing returns on additional monitoring as sample size increases.", cache=TRUE}
+knitr::include_graphics(here::here("figs/forPaper/powercmult6ay3aSf1.png"))
 ```
@@
 cr$c = compositionBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2))
-
-#get mean and sd as fn of w
-cs <- cr %>%
-  group_by(w) %>%
-  summarise(mean = mean(log(c)), sd = sd(log(c)))
+cs = compositionBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2),approx=T)
 
 cr = merge(cr,cs)
-cr$ca= exp(rnorm(nrow(cr),cr$mean,cr$sd))
+cr$ca= exp(rnorm(nrow(cr),mean=cr$mu,sd=cr$sig2^0.5))
 

### 2023-07-26 analysis/scripts/BiasModelSensitivity.R

@@ -32,18 +32,9 @@ plot(base2)
 #approximate distribution for Bayesian model - lognormal
 cr=expand.grid(w = c(2,3,4,5,6,7,8,9),rep=seq(1,10000))
 nr=nrow(cr)
-cr$c = compositionBiasCorrection(q=runif(nr,0,0),w=cr$w,u=runif(nr,0,0),z=runif(nr,0.5,0.5))
+cr$c = compositionBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2))
 
-#get mean and sd as fn of w
-cs <- cr %>%
-  group_by(w) %>%
-  summarise(m = mean(c), v = var(c))
-cs$v[cs$v==0]=0.000001
-
-#now find lognormal parameters from mean and variance
-#https://www.johndcook.com/blog/2022/02/24/find-log-normal-parameters/
-cs$sig2 = log(1+cs$v/cs$m^2)
-cs$mu = log(cs$m)-cs$sig2/2
+cs = compositionBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2),approx=T)
 
 cr = merge(cr,cs)
 cr$ca= exp(rnorm(nrow(cr),mean=cr$mu,sd=cr$sig2^0.5))
@@ -52,7 +43,6 @@ base = ggplot(cr,aes(x=c))+geom_histogram(aes(y=..density..))+
   facet_wrap(~w,nrow=3,labeller=label_both)+theme_bw()+
   geom_density(aes(x=ca),alpha=.2, fill="#FF6666",col = "#FF6666")
 plot(base)
-
 
 mns <- cr %>%
   group_by(w) %>%

### 2023-07-26 analysis/scripts/bayesianModelExamples.R

@@ -1,0 +1,69 @@
+# Prior and posterior predictions from Bayesian model
+library(caribouMetrics)
+library(tidyverse)
+library(ggpubr)
+theme_set(theme_bw())
+library(RColorBrewer)
+
+baseDir <- "."
+
+simBig<-getSimsNational(replicates=3000,forceUpdate=T) #If called with default parameters, use saved object to speed things up.
+
+monitoringScns = data.frame(obsYears=c(1,15),collarCount=c(1,30),cowMult=c(6),collarInterval=c(1),assessmentYrs=c(3))
+stateScns = data.frame(obsAnthroSlope=c(2),projAnthroSlope=c(2))
+stateScns = merge(stateScns,data.frame(rep=seq(1:1)))
+
+stateScns$sQuantile=0.8
+stateScns$rQuantile = 0.8
+scns=merge(monitoringScns,stateScns)
+
+scns$iAnthro = 0
+scns$tA = scns$iAnthro+(scns$obsYears)*scns$obsAnthroSlope
+scns$projYears = 50-scns$obsYears
+scns$N0 = 1000
+scns$adjustR = TRUE
+
+####################
+eParsIn = list()
+eParsIn$cowCounts <- data.frame(Year = 1981:2023,
+                                Count = 100,
+                                Class = "cow")
+eParsIn$freqStartsByYear <- data.frame(Year = 1981:2023,
+                                       numStarts = 30)
+eParsIn$collarOnTime=1
+eParsIn$collarOffTime=12
+eParsIn$collarNumYears=1
+
+labFontSize = 10; breakInterval=5
+
+priorResult = caribouMetrics:::runScnSet(scns[1,],eParsIn,simBig,getKSDists=F,printProgress=F)
+priorResult$obs.all=NULL
+recPrior =  plotRes(priorResult, "Recruitment", lowBound=0, highBound = 0.75,
+                    legendPosition="left",breakInterval=breakInterval,labFontSize=labFontSize)
+plot(recPrior)
+survPrior =  plotRes(priorResult, "Adult female survival", lowBound=0.6,
+                     legendPosition="none",breakInterval=breakInterval,labFontSize=labFontSize)
+plot(survPrior)
+lambdaPrior =  plotRes(priorResult, "Population growth rate", lowBound=0.6,
+                       legendPosition="none",breakInterval=breakInterval,labFontSize=labFontSize)
+plot(lambdaPrior)
+
[…]

### 2023-07-26 analysis/scripts/nationalModelTrajectories.R

@@ -11,11 +11,6 @@ err_col <- "grey50"
 
 baseDir <- "."
 
-getBiasCorrection<-function(w,q,u,z){
-  c = w*(q*u+1-u)/((q*u+w-u)*(1-z))
-  return(c)
-}
-
 #numbers from Johnson et al for validation
 johnsonCompare <- read.csv(paste0(baseDir,"/data/Johnson et al. figures5_6.csv"))
 
@@ -63,13 +58,13 @@ str(popGrowthPars)
 
 cr = data.frame(replicate=unique(rateSamples$replicate))
 nr= nrow(cr)
-cr$c = getBiasCorrection(q=runif(nr,0,0.6),w=runif(nr,3,9),u=runif(nr,0,0.2),z=runif(nr,0,0.2))
+cr$c = compositionBiasCorrection(q=runif(nr,0,0.6),w=runif(nr,3,9),u=runif(nr,0,0.2),z=runif(nr,0,0.2))
 rateSamples$c=NULL;rateSamples=merge(rateSamples,cr)
 
 str(rateSamplesLarge)
 cr = data.frame(replicate=unique(rateSamplesLarge$replicate))
 nr= nrow(cr)
-cr$c = getBiasCorrection(q=runif(nr,0,0.6),w=runif(nr,3,9),u=runif(nr,0,0.2),z=runif(nr,0,0.2))
+cr$c = compositionBiasCorrection(q=runif(nr,0,0.6),w=runif(nr,3,9),u=runif(nr,0,0.2),z=runif(nr,0,0.2))
 rateSamplesLarge$c=NULL;rateSamplesLarge=merge(rateSamplesLarge,cr)
 
 str(rateSamplesLarge)
@@ -142,7 +137,7 @@ plot_lambda <- ggplot(oo,
             alpha = 1) + scale_color_manual(values=pal)+
   scale_x_continuous(limits = c(-1, 90), breaks = c(0, 20, 40, 60, 80)) +
   xlab("Anthropogenic disturbance (%)") +
-  ylab(expression("Population Growth Rate " * lambda)) +
+  ylab(expression("Population Growth Rate no delay" * lambda)) +
   theme(legend.position = "none", plot.margin = margin(l = 0.6, unit = "cm"))
 
 str(pars1)
@@ -159,7 +154,7 @@ plot_recruitment3 <- ggplot(data = rateSummaries,
[…]

### 2023-07-26 analysis/scripts/viewSensitivityResults.R

@@ -13,10 +13,9 @@ scn_defaults <- eval(formals(getScenarioDefaults))
 
 ########################
 #sensitivity
-setName = "s5"
+setName = "s6"
 dir.create(paste0("figs/",setName),recursive=T)
-scns = read.csv(here::here("tabs/s4.csv"))
-scns = subset(scns,cmult!=0)
+scns = read.csv(here::here("tabs/s6.csv"))
 pagesa=unique(scns$pageId)
 
 pages = pagesa

### 2023-08-21 analysis/paper/paper.Rmd
prior calibration

@@
 runScenarios = F 
-doPriorCalibration=T #Note the prior calibration is time consuming. 
+doPriorCalibration=F #Note the prior calibration is time consuming. 
 ```
@@
 
-  #TO DO: move CaribouDemo_v1.15Reduce.R to paper directory. And produce pngs, rather than pdfs.
   source(here::here("analysis/scripts/calibratePriors.R"))
@@
 }
+  #library(ggplot2);fig_widths=data.frame(two=7)
+  KS$SigmaMean = KS$sInterannualVar
+  KS$SigmaSD = KS$sInterannualVarSE
+  KS=subset(KS,scn=="All")
   png(here::here("figs/PriorKSDists.png"),
-    height = 4, width = fig_widths["two"], units = "in",res=600)
-    base = ggplot(KS,aes(x=lse,y=KSDistance,color=sse))+
-    geom_point()+facet_grid(scn~ssv,labeller="label_both",scales="free_y")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("lse: multiplier on SD of survival intercept")
+    height = 3, width = fig_widths["two"], units = "in",res=100)
+    base = ggplot(KS,aes(x=sIntSEMod,y=KSDistance,color=SigmaMean))+
+    geom_point()+facet_wrap(~SigmaSD,labeller="label_both")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("sIntSEMod: multiplier on SD of survival intercept")
     print(base)
@@
 
+  KSRec$SigmaMean = KSRec$rInterannualVar
+  KSRec$SigmaSD = KSRec$rInterannualVarSE
+  KSRec=subset(KSRec,scn=="All")
   png(here::here("figs/PriorKSDistsRec.png"),
-    height = 4, width = fig_widths["two"], units = "in",res=600)
-    base = ggplot(KSRec,aes(x=lre,y=KSDistance,color=sre))+
-    geom_point()+facet_grid(scn~srv,labeller="label_both",scales="free_y")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("lre: multiplier on SD of recruitment intercept")
+    height = 3, width = fig_widths["two"], units = "in",res=600)
+    base = ggplot(KSRec,aes(x=rIntSEMod,y=KSDistance,color=SigmaMean))+
+    geom_point()+facet_wrap(~SigmaSD,labeller="label_both")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("rIntSEMod: multiplier on SD of recruitment intercept")
     print(base)
@@
 
-```{r PriorKSDists, fig.cap="Prior parameter values should be set so that, given minimal monitoring data ('All' scenario: 1 year of 1 collar per year), the posterior survival distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022].
-Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 45 collars per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local survival, rather than priors from the national model.
-When the multiplier on the prior standard deviation of the survival intercept (lse: $\\beta^S_0$ in Table @ref(tab:priors)) is 5, the prior is informative enough to meet the former condition, and vague enough to meet the latter.
-Given lse=5, low values of the mean standard deviations of the prior distribution of the random effect of year on survival (sse=0.034784, ssv=0.01, $\\sigma^2_S$ in Table @ref(tab:priors)) yield priors that meet both conditions (Figs @ref(fig:priorAllFits) and @ref(fig:priorLowFits)).
+Prior parameter values should be set so that, given no monitoring data, the prior survival distribution is similar to the distribution of simulated outcomes from the national model in the case of no anthropogenic disturbance.
+This condition is met when the multiplier on the prior standard deviation of the survival intercept (sIntSEMod: $\\beta^S_0$ in Table @ref(tab:priors)) is 5 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\\sigma^2_S$ in Table @ref(tab:priors)) (Figs @ref(fig:priorKSDists) and @ref(fig:priorAllFits)).
+Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 60 collars per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local survival, rather than priors from the national model.
+Setting the mean and standard deviations of the prior distribution of the random effect of year on survival to moderately low values (SigmaMean=0.034784, SigmaSD=0.03, $\\sigma^2_S$ in Table @ref(tab:priors)) allows both conditions to be met (Figs @ref(fig:priorKSDists), @ref(fig:priorAllFits) and @ref(fig:priorLowFits)).
+
+```{r PriorKSDists, fig.cap="Similarity of the prior survival distribution to the distribution of simulated outcomes from the national model measured by Kolmogorov-Smirnov distance [ks.test function in @R_R_2022].
+The distributions are similar when the multiplier on the prior standard deviation of the survival intercept (sIntSEMod: $\\beta^S_0$ in Table @ref(tab:priors)) is 5 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\\sigma^2_S$ in Table @ref(tab:priors)).
 Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
@@
 
-#Note: file wouldn't knit for Anna without access to all the newest png files below, so just commenting out the problem ones for now
 ```
 
-```{r PriorAllFits, fig.cap="Differences between posterior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
-In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0.
+```{r PriorAllFits, fig.cap="Differences between prior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring.
+In these examples, the standard deviation of the random effect prior (SigmaSD: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and anthropogenic disturbance is 0.
 Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
-# COMMENTED OUT FOR NOW CAUSE I DIDN"T HAVE ACCESS
-knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.03.png"))
+
+knitr::include_graphics(here::here("figs/SurvInterceptAllsInterannualVarSE0.03.png"))
 ```
 
-```{r PriorLowFits, fig.cap="Posterior predictive survival distributions from the Bayesian model ('local') assuming intensive monitoring (15 years of 45 collars per year) of populations from the 2.5% quantile of the national model.
-In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0.
-Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
-# COMMENTED OUT FOR NOW CAUSE I DIDN"T HAVE ACCESS
-knitr::include_graphics(here::here("figs/SurvsQStarts45lowssv0.03.png"))
+```{r PriorLowFits, fig.cap="Posterior predictive survival distributions from the Bayesian model ('local') assuming intensive monitoring (15 years of 60 collars per year) of populations from the 2.5 \\% quantile of the national model.
+In these examples, the standard deviation of the random effect prior (SigmaSD: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0.
+Bands show 95 \\% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.
+When the multiplier on the prior standard deviation of the survival intercept (sIntSEMod: $\\beta^S_0$ in Table @ref(tab:priors)) is too low (<5) the insufficiently vague priors substantially influence the posterior distributions, yielding poor local parameter estimates.", cache=TRUE}
+
+knitr::include_graphics(here::here("figs/SurvsurvInterceptLowsInterannualVarSE0.03.png"))
 ```
 
-```{r PriorAllFitsAnthro90, fig.cap="Differences between posterior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
-In these examples, anthropogenic disturbance is 90%.
-When the multiplier on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) is 3, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model.
-Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts1Anthro90.png"))
+```{r PriorAllFitsAnthro90, fig.cap="Differences between prior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring.
+In these examples, anthropogenic disturbance is 90\\%.
+When the multiplier on the prior standard deviation of the survival slope (sSlopeSEMod: $\\beta^S_a$ in Table @ref(tab:priors)) is 3, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model.
+Bands show 95\\% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvSlopeAll.png"))
 ```
 
-```{r PriorHighFitsAnthro90, fig.cap="Posterior predictive survival distributions from the Bayesian model, assuming intensive monitoring (15 years of 45 collars per year) of populations from the 97.5% quantile of the national model.
-In these examples, anthropogenic disturbance is 90%, and the effect of anthropogenic disturbance varies from none (sS=0) through expected (sS=1) to double the expected slope (sS=2).
-All considered multipliers on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.
-Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts45Anthro90high.png"))
+```{r PriorHighFitsAnthro90, fig.cap="Posterior predictive survival distributions from the Bayesian model, assuming intensive monitoring (15 years of 60 collars per year) of populations from the 97.5 \\% quantile of the national model.
+In these examples, anthropogenic disturbance is 90\\%, and the effect of anthropogenic disturbance varies from none (sSlopeMod=0) through expected (sSlopeMod=1) to double the expected slope (sSlopeMod=2).
+All considered multipliers on the prior standard deviation of the survival slope (sSlopeSEMod: $\\beta^S_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.
+Bands show 95 \\% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
+knitr::include_graphics(here::here("figs/SurvSlopeHigh.png"))
 ```
 
-```{r PriorKSDistsRec, fig.cap="Prior parameter values should be set so that, given no monitoring data ('All' scenario), the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022].
-Setting the multiplier on the prior standard deviation of the recruitment intercept (lre: $\\beta^R_0$ in Table @ref(tab:priors)) to 3, and values of the mean and standard deviations of the prior distribution of the random effect of year on recruitment to sre=0.23 and srv=0.22, respectively ($\\sigma^2_R$ in Table @ref(tab:priors)) allows us to meet this condition (Fig @ref(fig:priorAllFitsRec)).
-Priors should also be vague enough so that, given intensive monitoring of populations where recruitment is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 135 cows per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local recruitment, rather than priors from the national model.
-When the multiplier on the prior standard deviation of the recruitment intercept (lre: $\\beta^R_0$ in Table @ref(tab:priors)) is 1, the prior is not sufficiently vague to meet the latter condition (Fig @ref(fig:PriorLowFitsRec)).
+Similarly, prior parameter values should be set so that, given no monitoring data, the prior recruitment distribution is similar to the distribution of simulated outcomes from the national model in the case of no anthropogenic disturbance.
+This condition is met when the multiplier on the prior standard deviation of the recruitment intercept (rIntSEMod: $\\beta^R_0$ in Table @ref(tab:priors)) is 4 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\\sigma^2_R$ in Table @ref(tab:priors)) (Figs @ref(fig:priorKSDistsRec) and @ref(fig:priorAllFitsRec)).
+Priors should also be vague enough so that, given intensive monitoring of populations where recruitment is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 9*60 cows per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local recruitment, rather than priors from the national model.
+Setting the mean and standard deviations of the prior distribution of the random effect of year on recruitment to moderately low values (SigmaMean=0.23, SigmaSD=0.22, $\\sigma^2_R$ in Table @ref(tab:priors)) allows both conditions to be met (Figs @ref(fig:priorKSDistsRec), @ref(fig:priorAllFitsRec) and @ref(fig:priorLowFitsRec)).
+
+```{r PriorKSDistsRec, fig.cap="Similarity of the prior recruitment distribution to the distribution of simulated outcomes from the national model measured by Kolmogorov-Smirnov distance [ks.test function in @R_R_2022].
+The distributions are similar when the multiplier on the prior standard deviation of the survival intercept (rIntSEMod: $\\beta^R_0$ in Table @ref(tab:priors)) is 4 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\\sigma^2_R$ in Table @ref(tab:priors)).
 Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
@@
 
-```{r PriorAllFitsRec, fig.cap="Differences between posterior predictive recruitment distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no monitoring.
-In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0.
+```{r PriorAllFitsRec, fig.cap="Differences between prior predictive recruitment distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring.
+In these examples, the standard deviation of the random effect prior (SigmaSD: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and anthropogenic disturbance is 0.
 Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
-knitr::include_graphics(here::here("figs/RecrQStarts1srv0.22.png"))
+knitr::include_graphics(here::here("figs/RecInterceptAllrInterannualVarSE0.22.png"))
 ```
 
-```{r PriorLowFitsRec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model ('local') assuming intensive monitoring (135 cows per year for 15 years) of populations from the 2.5% quantile of the national model.
-In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0.
+```{r PriorHighFitsRec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model ('local') assuming intensive monitoring (9*60 cows per year for 15 years) of populations from the 97.5% quantile of the national model.
+In these examples, the standard deviation the random effect prior (SigmaSD: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0.
 Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/RecrQStarts45lowsrv0.22.png"))
+knitr::include_graphics(here::here("figs/RecInterceptHighrInterannualVarSE0.22.png"))
 ```
 
-```{r PriorAllFitsAnthro90Rec, fig.cap="Differences between posterior recruitment predictive distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no monitoring.
-In these examples, anthropogenic disturbance is 90%.
-When the multiplier on the prior standard deviation of the slope of the relationship between recruitment and anthropogenic disturbance (bre: $\\beta^R_a$ in Table @ref(tab:priors)) is 4, the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model.
-Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
-knitr::include_graphics(here::here("figs/RecrQStarts1Anthro90.png"))
+```{r PriorAllFitsAnthro90Rec, fig.cap="Differences between prior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring.
+In these examples, anthropogenic disturbance is 90\\%.
+When the multiplier on the prior standard deviation of the recruitment slope (rSlopeSEMod: $\\beta^R_a$ in Table @ref(tab:priors)) is 4, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model.
+Bands show 95\\% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
+knitr::include_graphics(here::here("figs/RecSlopeAll.png"))
 ```
 
-```{r PriorHighFitsAnthro90Rec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model, assuming intensive monitoring (15 years of 135 collars per year) of populations from the 97.5% quantile of the national model.
-In these examples, anthropogenic disturbance is 90%, and the effect of anthropogenic disturbance varies from none (rS=0) through expected (rS=1) to double the expected slope (rS=2).
-All multipliers on the prior standard deviation of the survival slope (bre: $\\beta^R_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.
-Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/RecrQStarts45Anthro90high.png"))
+```{r PriorHighFitsAnthro90Rec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model, assuming intensive monitoring (15 years of 60 collars per year) of populations from the 97.5 \\% quantile of the national model.
+In these examples, anthropogenic disturbance is 90\\%, and the effect of anthropogenic disturbance varies from none (rSlopeMod=0) through expected (rSlopeMod=1) to double the expected slope (rSlopeMod=2).
+All considered multipliers on the prior standard deviation of the recruitment slope >1 (rSlopeSEMod: $\\beta^R_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.
+Bands show 95 \\% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
+knitr::include_graphics(here::here("figs/RecSlopeHigh.png"))
 ```

### 2023-08-21 R/CaribouDemographicFunctions.R

@@ -43,15 +43,15 @@ runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMei
   return(scResults)
 }
 
-makeInterceptPlots <- function(scResults, addBit = "", facetVars = c("P", "sQ"),
+makeInterceptPlots <- function(scResults, addBit = "", facetVars = c("sIntSEMod","sInterannualVar"),
                                loopVars = NULL,
                                whichPlots = c("Adult female survival",
                                               "Population growth rate",
                                               "Recruitment",
                                               "Female population size"),
                                survLow = 0.6, type = "png", useNational = T) {
-  # facetVars=c("lre","sre");loopVars="srv";scResults=scResultsHigh
-
+  # facetVars=c("sIntSEMod","sInterannualVar");loopVars="sInterannualVarSE";scResults=scResults
+  #useNational=T;type="png"; survLow=0.6
   if (!is.null(loopVars)) {
     loopSet <- unique(subset(scResults$rr.summary.all, select = loopVars))
     loopSet$dummy <- 1
@@ -74,10 +74,14 @@ makeInterceptPlots <- function(scResults, addBit = "", facetVars = c("P", "sQ"),
 
     addBitO <- paste0(addBit, aa)
 
-    if (useNational) {
-      simRange <- merge(scResults$sim.all, crow)
-    } else {
-      simRange <- NULL
+    scResCur = scResults
+    scResCur$sim.all = merge(scResCur$sim.all,crow)
+    scResCur$rr.summary.all = merge(scResCur$rr.summary.all, crow)
+    if(is.element("obs.all",names(scResCur))){
+      scResCur$obs.all = merge(scResCur$obs.all, crow)
+    }
+    if (!(useNational)) {
+      scResCur$sim.all = NULL
     }
 
     if (is.element("Adult female survival", whichPlots)) {
@@ -88,9 +92,8 @@ makeInterceptPlots <- function(scResults, addBit = "", facetVars = c("P", "sQ"),
       } else {
         pdf(paste0("figs/Surv", addBitO, ".pdf"), width = 10, height = 7)
       }
-      print(caribouMetrics:::plotRes(merge(scResults$rr.summary.all, crow), "Adult female survival",
-                    obs = merge(scResults$obs.all, crow),
-                    lowBound = survLow, simRange = simRange, facetVars = facetVars
+      print(caribouMetrics:::plotRes(scResCur, "Adult female survival",
[…]

### 2023-08-21 analysis/scripts/calibratePriors.R

@@ -1,36 +1,37 @@
 #devtools::load_all(here::here())
 
-scnsApplyAll = data.frame(qMin=0,qMax=0,uMin=0,uMax=0,zMin=0,zMax=0,adjustR=F,
-                          obsYears=1,collarCount=1,cowMult=1,projYears=1)
-#TO DO: resume here. Make cowMult=0 work.
-
+scnsApplyAll = data.frame(qMin=0,qMax=0,uMin=0,uMax=0,zMin=0,zMax=0,adjustR=F)
+scnsNoData = data.frame(obsYears=1,collarCount=0,cowMult=0,projYears=1)
+scnsLow = data.frame(obsYears=15,collarCount=60,cowMult=9,projYears=1,rQuantile=0.025,sQuantile=0.025)
+scnsHigh = scnsLow;scnsHigh$rQuantile =0.975;scnsHigh$sQuantile=0.975
 ########################
 #Survival
 #vary sIntSEMod, sInterannualVar, and sInterannualVarSE in factorial array. Outcome of interest is KS distance
 sIntSEMod=seq(1,10,by=2);sInterannualVar=0.08696*seq(0.2,1,by=0.2);sInterannualVarSE=seq(0.01,0.09,by=0.02)
 
 #Looking for low KS distance from full range of input sims when given almost no info
-scns=expand.grid(obsAnthroSlope = 0, projAnthroSlope = 0,sIntSEMod=sIntSEMod,sInterannualVar=sInterannualVar,sInterannualVarSE=sInterannualVarSE)
-scns=merge(scns,scnsApplyAll)
+scnsSInt=expand.grid(obsAnthroSlope = 0, projAnthroSlope = 0,sIntSEMod=sIntSEMod,
+                     sInterannualVar=sInterannualVar,sInterannualVarSE=sInterannualVarSE)
+scnsSInt$SigmaMean = scnsSInt$sInterannualVar
+scns=merge(scnsSInt,scnsApplyAll);scns=merge(scns,scnsNoData)
 scResults = runScenario(scns,survAnalysisMethod = "Exponential")
-KSAll = subset(scResults$ksDists,(Year==2019)&(Parameter=="Adult female survival"))
-addBit = paste0("sQStarts",startsByYr)
[…]

### 2023-08-21 analysis/scripts/sensitivityMinimalTest.R

@@ -38,7 +38,7 @@ scns = subset(allScns, pageId==cpageId)
 
 str(scns)
 
-scns$zMin= 0.5; scns$zMax = 0.5
+scns$zMin= 0; scns$zMax = 0
 scns$uMin = 0; scns$uMax = 0
 scns$qMin = 0; scns$qMax = 0
 message("batch ", cpageId, " started")

### 2023-08-23 analysis/paper/paper.Rmd
Separating supplementary material from main text.

--- 2023-07-26 analysis/paper/paper.Rmd
@@
 
+<!--
 Motivating example - forecasting impacts of disturbance - WBI and RoF example. In order to reduce uncertainty, need to integrate local information on current state of the population into forecasts.
@@
 More generally, need for tools to help assess the value of additional monitoring and the best way to invest additional monitoring resources, given the info that is already available.
-
-Topics of interest, questions, goals:
-
-* How get a demographic forecast informed by local demographic data and national disturbance-demographic relationships?
-* How to assess the value of collecting additional demographic data? How to prioritize whether to e.g. monitor many areas for a short time or fewer areas over the long term? 
-* How much of a bias/misinterpretation can be (has been?) incurred if we don't recognize local differences?
-* How does that value depend on current status of the range, available existing demographic data, and projected future disturbance.
-When and why is more information required, and when do we know enough?
-* Provide decision support tools to enable others to get forecasts informed by their local data and national models, and assess the value of monitoring scenarios.
+-->
 
@@
 
+TO DO: Anna draft this section?
 Marking of caribou cows with telemetry GPS/VHF collars is the traditional approach to estimating demographic parameters (refs/examples, e.g.
@@
 
-knitr::kable(tbl,caption="Monitoring scenarios parameters and values.")
+knitr::kable(tbl,caption="Monitoring scenario parameters.")
 ```
@@
 Collars lost to mortality may be replaced each year ($o=1$), or there may be a 3 year gap between deployments ($o=4$).
+We focus on exploring a set of scenarios in which anthropogenic disturbance, measured as the percentage of area within 500 m of disturbance [@johnson_science_2020;@ECCC_scientific_2011], increases over time from a variety of starting points (Table \@ref(tab:monitoringScenarios)).
+Scenarios in which anthropogenic disturbance does not increase over time are included in Supplement D.
 
+<!--
 *Monitoring WG input from BC, ON, NL, (MB pending): lots of detail provided but generally those numbers suggest we could try values of ~3-10 [e.g.
@@
 TO DO: run low initial population size scenario, put in appendix of one by one sensitivity checks? 
+-->
 
@@
 The true number of post-juvenile females that survive from year $t$ to the census $\dot{W}_t$ is binomially distributed with true survival probability $\dot{S}_t$: $\dot{W}_{t} \sim \text{Binomial}(\dot{N}_t,\dot{S}_t)$.
-Maximum potential recruitment rate is adjusted for sex ratio, misidentification biases (Appendix X) and (optionally) delayed age at first reproduction $$\dot{X}_t=\frac{c\dot{R}_t/2}{1+c\dot{R}_t/2};c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}, $$
-where $w$ is a multiplier that defines the apparent number of adult females in the composition survey as a function of the number of collared animals, $q$ the ratio of young bulls to adult females, $u$ is the probability of misidentifying young bulls as adult females and vice versa, and $z$ is the probability of missing a calf.
+Maximum potential recruitment rate is adjusted for sex ratio, misidentification biases (Supplement B) and (optionally) delayed age at first reproduction [@eacker_web-based_2019;@decesare_estimating_2012]: $$\dot{X}_t=\frac{c\dot{R}_t/2}{1+c\dot{R}_t/2};c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}, $$
+where $w$ is a multiplier that defines the apparent number of adult females in the composition survey as a function of the number of collared animals, $q$ is the ratio of young bulls to adult females, $u$ is the probability of misidentifying young bulls as adult females and vice versa, and $z$ is the probability of missing a calf.
 
-Realized recruitment rate varies with population density [@lacy_vortex_2017], and the number of juveniles recruiting to the post-juvenile class at the census is a binomially distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate: $$\dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{N_0k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}).$$ Given parameters in Table X, recruitment rate is lowest $(0.5\dot{X}_t)$ when $\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$ at intermediate population sizes, and declines to $0.6\dot{X}_t$ as the population reaches carrying capacity of $k=100$ times the initial population size.
+Realized recruitment rate varies with population density [@lacy_vortex_2017], and the number of juveniles recruiting to the post-juvenile class at the census is a binomially distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate: $$\dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{N_0k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}).$$ Given parameters in Table A1, recruitment rate is lowest $(0.5\dot{X}_t)$ when $\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$ at intermediate population sizes, and declines to $0.6\dot{X}_t$ as the population reaches carrying capacity of $k=100$ times the initial population size.
 The post-juvenile female population in the next year includes both survivors and new recruits: $\dot{N}_{t+1}=\text{min}(\dot{W}_t+\dot{J}_t,r_{max}\dot{N}_t)$.
@@
 Interannual variation in survival and recruitment is modelled using truncated beta distributions [rtrunc function; @novomestky_package_2016]: $\dot{R}_t \sim \text{TruncatedBeta}(\bar{R}_t,\nu_R,l_R,h_R); \dot{S}_t \sim \text{TruncatedBeta}(\bar{S}_t,\nu_S,l_S,h_S)$.
-Coefficients of variation among years $(\nu_R,\nu_S)$ and maximum/minimum values $l_R,h_R,l_S,h_S$ for recruitment and survival are given in Table X.
+Coefficients of variation among years $(\nu_R,\nu_S)$ and maximum/minimum values $l_R,h_R,l_S,h_S$ for recruitment and survival are given in Table A1.
 Expected recruitment ($\bar{R}_t$) and survival ($\bar{S}_t$) vary with disturbance according the beta regression models estimated by Johnson et al.
@@
 
-In our simulated populations, mortality risk does not vary over time or among individuals, so the observed state (dead or alive) of a collared animal $i$ in month $m$ and year $t$ depends on the true survival rate $\dot{S}_t$ (eq X) as follows:
+In our simulated populations, mortality risk does not vary over time or among individuals, so the observed state (dead or alive) of a collared animal $i$ in month $m$ and year $t$ depends on the true survival rate $\dot{S}_t$ as follows:
 $$\hat{I}_{i,t,m} \sim \text{Bernoulli}(\dot{S}^{1/12}_t\hat{I}_{i,t,m-1}).$$
@@
 
-We model recruitment probability as a function of anthropogenic disturbance and fire with a log link [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a A_t+\beta^R_f F_t+\epsilon^R_t; \epsilon^R_t \sim \text{Normal}(0,\sigma^2_{R}).$$
-Estimated recruitment probability is adjusted for sex ratio, misidentification biases (Appendix X), and (optionally) age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012] to get expected recruits per female:
-$$X_t=\frac{cR_t/2}{1+cR_t/2}.$$
-To maintain consistency with Johnson et al. [-@johnson_science_2020] we also provide an option to only adjust for sex ratio and misidentification biases $X_t=cR_t/2$.
+We model recruitment probability as a function of anthropogenic disturbance and fire with a log link [@johnson_science_2020;@stewart_climate-informed_2023;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a A_t+\beta^R_f F_t+\epsilon^R_t; \epsilon^R_t \sim \text{Normal}(0,\sigma^2_{R}).$$
+Estimated recruitment probability is adjusted for sex ratio, misidentification biases (Supplement B), and age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012] to get expected recruits per female:
+$$X_t=\frac{cR_t/2}{1+cR_t/2}; c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}.$$
+The composition survey bias correction term $c$ depends on the apparent number of adult females per collared animal $w$, the ratio of young bulls to adult females $q$, the probability of misidentifying young bulls as adult females and vice versa $u$, and the probability of missing a calf $z$ (Supplement B).
+To maintain consistency with Johnson et al.
+[-@johnson_science_2020] we also provide an option to only adjust for sex ratio and misidentification biases $X_t=cR_t/2$.
 
-The composition survey bias correction term $c$ depends on the apparent number of adult females per collared animal $w$, the ratio of young bulls to adult females $q$, the probability of misidentifying young bulls as adult females and vice versa $u$, and the probability of missing a calf $z$ (Appendix X):
-$$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}$$
-In the Bayesian model, prior uncertainty about the value of the bias correction term $c$ is approximated with a Log-normal distribution (Appendix X).
+In the Bayesian model, prior uncertainty about the value of the bias correction term $c$ is approximated with a Log-normal distribution (Supplement B).
 Given the apparent number of adult females per collared animal $w$ in each monitoring scenario (Table \@ref(tab:monitoringScenarios)), we calculate the mean and standard deviation of 10,000 samples of $\log{c}$ assuming the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2.
 
-Also following [@eacker_web-based_2019], Kaplan-Meier estimates of observed adult female survival probability $\hat{S}_t$ and associated precisions $\tau_t$ are estimated from known-fate radio collar data using the survival R package [@therneau_modeling_2000], with precisions for years with no observed mortality estimated using a separate Bayesian model.
+Also following Eacker et al.
+[-@eacker_web-based_2019], Kaplan-Meier estimates of observed adult female survival probability $\hat{S}_t$ and associated precisions $\tau_t$ are estimated from known-fate radio collar data using the survival R package [@therneau_modeling_2000], with precisions for years with no observed mortality estimated using a separate Bayesian model.
 Observed estimated survival is a Gaussian distributed function of the true survival probability: $$\hat{S_t} \sim \text{Normal}(S_t,1/\tau_t).$$
-We model true survival probability as an adjusted function of buffered anthropogenic disturbance $A_t$ with a log link [@johnson_science_2020;@stewart_climate-informed_nodate;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a A_t+\epsilon^S_t; \epsilon^S_t \sim \text{Normal}(0,\sigma^2_{S}).$$
-To assess sensitivity of results to the survival estimation method, we also implemented an alternative parametric exponential survival model, wherein the observed state (alive or dead, $\hat{I}_{i,t,m}$) of an individual animal $i$ in year $t$ and month $m$ is Bernoulli distributed (ref IPM book): $$\hat{I}_{i,t,m} \sim \text{Bernoulli}(S^{1/12}_t\hat{I}_{i,t,m-1}).$$
+As in the national demographic model, we model survival probability as an adjusted function of buffered anthropogenic disturbance $A_t$ with a log link [@johnson_science_2020;@stewart_climate-informed_2023;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a A_t+\epsilon^S_t; \epsilon^S_t \sim \text{Normal}(0,\sigma^2_{S}).$$
+To assess sensitivity of results to the survival estimation method and handle cases in which there is little or no survival data, we also implemented an alternative parametric exponential survival model, wherein the observed state (alive or dead, $\hat{I}_{i,t,m}$) of an individual animal $i$ in year $t$ and month $m$ is Bernoulli distributed [@schaub_integrated_2021]: $$\hat{I}_{i,t,m} \sim \text{Bernoulli}(S^{1/12}_t\hat{I}_{i,t,m-1}).$$
 
 All regression coefficients are assumed to be Gaussian distributed, with priors derived from the expected values and 95% confidence intervals estimated by Johnson et al (2020) (Table \@ref(tab:priors)).
-We calibrated the prior distributions so that the 95% prior prediction intervals for survival and recruitment from the Bayesian model match the range between the 2.5% and 97.5% quantiles of 1000 simulated survival and recruitment trajectories (Appendix A, Figure \@ref(fig:examplePredictions)).
-
+We calibrated the prior distributions so that the 95% prior prediction intervals for survival and recruitment from the Bayesian model match the range between the 2.5% and 97.5% quantiles of 1000 simulated survival and recruitment trajectories (Supplement C, Figure \@ref(fig:examplePredictions)).
 ```{r priors, echo=FALSE,message=FALSE}
@@
 There is probably some more elegant way to deal with this, but it won't make any difference as long as the collar deployment month is set to 1.
-* add references to documentation of popGrowthJohnson and other caribouMetrics functions.
 * consider & comment on alternative choices made in this example https://poissonconsulting.shinyapps.io/bboushiny/. Follow Eacker example in all particulars, change some aspects, or do comparisons?
 * Ensure notation consistency amongst manuscript, UI, documentaiton and code where appropriate.
-
-TO DO: add example trajectory figure
 
@@
 
-\newpage
-
-# Appendix
-##Definitions of variables and parameters in the Bayesian population model
-
-Table A1. Variables and parameters in the Bayesian population model and the national demographic model.
-
-|**Symbol**|**Units**|**Description**|
-|:-----|-----|:-----------|
-|**Indices & constants**|
-|$t$|year|Year.|
-|$m$|month|Month.|
-|$i$|ID|Individual animal identifier.|
-|**Observed**|
-|$A_t$|%|Buffered anthropogenic disturbance.|
-|$F_t$|%|Fire disturbance.| 
-|$\hat{J}_t$|animals|Calves observed in calf:cow survey.|
-|$\hat{W}_t$|animals|Cows observed in calf:cow survey.|
-|$\hat{T}_t$|animals|Number of collared cows.|
-|$\hat{S}_t$|probability|Kaplan-Meier estimate of observed adult female survival.|
-|$\tau_t$||Precision of Kaplan-Meier survival estimate.|
-|$\hat{I}_{i,t,m}$|0 or 1|Indicator of whether animal $i$ is dead or alive.|
-|**Estimated state**|
-|$R_t$|rate|Recruitment (calves per cow).|
-|$S_t$|probability|Adult female survival.|
-|$N_t$|animals|Adult female population size.|
-|$\bar{\lambda}_{y,t}$|rate|Average population growth over $y$ years.|
-|$X_t$|rate|Recruitment adjusted for sex ratio and (optionally) delayed age of first reproduction.|
-|$W_t$|animals|Surviving adult females before recruitment.|
-|$J_t$|animals|New recruits to adult age class.|
-|$\lambda_t$|rate|Annual population growth.|
-|**Estimated Parameters**|
-|$\beta^R_0$||Recruitment intercept.|
-|$\beta^R_a$||Anthropogenic disturbance slope for recruitment.|
-|$\beta^R_f$||Fire slope for recruitment.|
-|$\sigma^2_{R}$||Variance in random effect of year on recruitment.|
-|$\beta^S_0$||Survival intercept.|
-|$\beta^S_a$||Anthropogenic disturbance slope for survival.|
-|$\sigma^2_{S}$||Variance in random effect of year on survival.|
-|**True state**|
-|$\dot{R}_t$|rate|Recruitment (calves per cow).|
-|$\dot{S}_t$|probability|Adult female survival.|
-|$\dot{N}_t$|animals|Adult female population size.|
-|$\dot{X}_t$|rate|Recruitment adjusted for sex ratio and (optionally) delayed age of first reproduction.|
-|$\dot{W}_t$|animals|Surviving adult females before recruitment.|
-|$\dot{J}_t$|animals|New recruits to adult age class.|
-|**True demographic parameters**|
-|$N_0$|animals|Initial adult population size.|
-|$\dot{\beta}^R_0$||Recruitment intercept.|
-|$\dot{\beta}^R_a$||Anthropogenic disturbance slope for recruitment.|
-|$\dot{\beta}^R_f$||Fire slope for recruitment.|
-|$\dot{\beta}^S_0$||Survival intercept.|
-|$\dot{\beta}^S_a$||Anthropogenic disturbance slope for survival.|
-|$p_0=1$||Maximum recruitment multiplier.|
-|$p_k=0.6$||Recruitment multiplier at carrying capacity.|
-|$k=100$||Carrying capacity multiplier.|
-|$b=4$||Density dependence shape parameter.|
-|$a=1$||Allee effect parameter.|
-|$r_{max}=1.3$||Maximum population growth rate.|
-|$\nu_R=0.46$||Coefficient of variation in $\dot{R}_t$ among years.|
-|$\nu_S=0.08696$||Coefficient of variation in $\dot{S}_t$ among years.|
-|$l_R=0$||Minimum value for $\dot{R}_t$.|
-|$l_S=0$||Minimum value for $\dot{S}_t$.|
-|$h_R=0$||Maximum value for $\dot{R}_t$.|
-|$h_S=0$||Maximum value for $\dot{S}_t$.|
-|**Observation model parameters**|
-|$d=1,2,4,8,16, \text{ or } 24$|years|Monitoring duration.|
-|$n=15,30, \text{ or } 60$|animals|Target number of collared females.|
-|$o=1 \text{ or } 4$|years|Years between collar deployments.|
-|$y=3$|years|Assessment period for population growth rate.|
-|$w=3,6,\text{ or } 9$|ratio|The apparent number of adult females per collared animal in composition survey.|
-|$q=\text{uniform}(0,0.6)$|ratio|True ratio of young bulls to adult females in composition survey.|
-|$u=\text{uniform}(0,0.2)$|probability|Probability of misidentifying young bulls as adult females and vice versa in composition survey.|
-|$z=\text{uniform}(0,0.2)$|probability|Probability of missing calves in composition survey.|
-|$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}$||Bias correction term (Appendix X).|
-
-## Model of bias in recruitment estimates from calf:cow surveys
-
-TO DO: get photo of calf:cow survey group. Label classes of animals to illustrate meaning of T, U, V, W, O, P, J, K.
-
-We assume each group of animals in a calf:cow composition survey contains one or more collared adult females ($T$), and may also include: uncollared adult females misidentified as young bulls or unknown sex ($U$); correctly identified uncollared adult females ($V$); young bulls correctly identified as male or unknown sex ($O$); young bulls misidentified as uncollared adult females ($P$); observed calves ($J$); and unobserved calves ($K$).
-The apparent number of adult females in the group is $T+V+P=Tw$, where $w$ is a multiplier that defines the apparent number of adult females as a function of the number of collared animals.
-The ratio of young bulls to uncollared adult females in the group is:
-$$q = \frac{P+O}{U+V}.$$
-Assuming an equal probability $u$ of misidentifying young bulls as adult females and vice versa, we get $V=(U+V)(1-u)$ and $P=(O+P)u$. Given a probability $z$ of missing calves, we get $J=(J+K)(1-z)$.
-
-Our objective is to model the sex and bias-corrected recruitment rate $X=\frac{J+K}{2(T+U+V)}$ as a function of the observed calf:cow ratio $R=J/(T+V+P)$, the cow multiplier $w$, the ratio of young bulls to adult females $q$, and the misidentification probabilities $u$ and $z$.
-We start by solving for $T+U+V$ as a function of $q,w,u$ and $T$.
-Recognize that $P=Tw-T-V$, $U+V=V/(1-u)$, and $P+O=P/u$ to write $q$ as
-$$q=\frac{Tw-T-V}{uV/(1-u)}.$$
-Rearrange to get 
-$$V=\frac{T(w-1)(1-u)}{qu+1-u}.$$
-Recognize that $U=Vu/(1-u)$ to write $T+U+V$ as a function of $q,w,u$ and $T$: 
-$$T+U+V=T\frac{qu+w-u}{qu+1-u}.$$
-Recognize that the number of observed calves $J$ is the product of the apparent recruitment rate and the apparent number of adult females $J=RTw$, and that therefore $J+K=RTw/(1-z)$ to rewrite the bias corrected recruitment rate $X=\frac{J+K}{2(T+U+V)}$ as a function of $w,u,z$ and $R$:
-$$X=R\frac{w(1+qu-u)}{2(w+qu-u)(1-z)}.$$
-For simplicity, we write $X$ as a function of a bias correction term $c$: 
-$$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}; X=cR/2$$
-
-If we also adjust for delayed age at first reproduction (DeCesare et al. 2012; Eacker et al. 2019), the adjusted recruitment rate becomes $$X=\frac{cR/2}{1+cR/2}.$$
-
-```{r biasPlot, fig.cap="Variation in the sex and bias corrected recruitment rate $X$ with the cow multiplier $w$, the ratio of young bulls to adult females $q$, and the adult misidentification probability $u$, and the probability of missing a calf $z$.
-If only collared females are counted ($w=1$) and there is a risk of missing calves $z>0$ then the apparent recruitment rate $R$ will be lower than $X$ (first column).
-When uncollared animals are counted ($w>1$), the risk of misidentifying adults $u$ inflates the apparent recruitment rate $R$ as on average more cows are misidentified as bulls than vice versa, but this effect is smaller than the effect of missing calves $z$.
-Increasing the proportion of bulls in the group $q$ reduces the effect of adult misidentification $u$.", fig.height=6,cache=TRUE}
-library(ggplot2)
-#Notes
-#q: assume more females than males in calf:cow group, so q<1
-
-p=expand.grid(delay=c(T,F),q=c(0,0.3,0.6),w=c(1,3,6),
-              u=c(0,0.1,0.2),z=seq(0,0.2,length.out=3),R=seq(0.15,0.6,length.out=10))
-
-getX<-function(R,delay,q,w,u,z){
-  c = compositionBiasCorrection(w,q,u,z)
-  X = 0.5*R*c
-  X=X*(1-delay)+delay*X/(1+X)
-  return(X)
-}
-
-p$X = getX(p$R,p$delay,p$q,p$w,p$u,p$z)
-
-p$grp = paste(p$u,p$z)
-p$u=as.factor(p$u)
-p$z=as.factor(p$z)
-
-base = ggplot(subset(p,!delay),aes(x=R,y=2*X/R,col=z,linetype=u,group=grp))+
-  geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()+ylab("2X/R=c, without delayed reproduction")
-
-base2 = ggplot(subset(p,delay),aes(x=R,y=2*X/R,col=z,linetype=u,group=grp))+
-  geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()+ylab("2X/R, with delayed reproduction")
-
-ggpubr::ggarrange(base, base2, labels = "",
-                  ncol = 1, vjust = 1)
-
-```
-
-```{r biasDistribution, fig.cap="In the Bayesian model, uncertainty about the value of the bias correction term $c$ (grey bars) is approximated with a Log-normal distribution (red).
-In these examples, the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2.
-We assume that the apparent number of adult females per collared animal $w$ is known.", cache=TRUE}
-
-cr=expand.grid(w = c(1,2,3,4,5,6,7,8,9),rep=seq(1,10000))
-nr=nrow(cr)
-cr$c = compositionBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2))
-cs = compositionBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2),approx=T)
-
-cr = merge(cr,cs)
-cr$ca= exp(rnorm(nrow(cr),mean=cr$mu,sd=cr$sig2^0.5))
-
-base = ggplot(cr,aes(x=c))+geom_histogram(aes(y=..density..))+
-  facet_wrap(~w,nrow=3,labeller=label_both)+theme_bw()+
-  geom_density(aes(x=ca),alpha=.2, fill="#FF6666",col = "#FF6666")
-plot(base)
-
-```
-
-## Calibration of Priors
-
-```{r priorCalibration, cache=TRUE,echo=FALSE, message=FALSE,results='hide'}
-if(doPriorCalibration){
-
-  #TO DO: move CaribouDemo_v1.15Reduce.R to paper directory. And produce pngs, rather than pdfs.
-  source(here::here("analysis/scripts/calibratePriors.R"))
-}else{
-  KS<-read.csv(here::here("tabs/SurvKSAll.csv"))
-  KSRec<-read.csv(here::here("tabs/RecKSAll.csv"))
-
-}
-  png(here::here("figs/PriorKSDists.png"),
-    height = 4, width = fig_widths["two"], units = "in",res=600)
-    base = ggplot(KS,aes(x=lse,y=KSDistance,color=sse))+
-    geom_point()+facet_grid(scn~ssv,labeller="label_both",scales="free_y")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("lse: multiplier on SD of survival intercept")
-    print(base)
-  dev.off()
-
-  png(here::here("figs/PriorKSDistsRec.png"),
-    height = 4, width = fig_widths["two"], units = "in",res=600)
-    base = ggplot(KSRec,aes(x=lre,y=KSDistance,color=sre))+
-    geom_point()+facet_grid(scn~srv,labeller="label_both",scales="free_y")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("lre: multiplier on SD of recruitment intercept")
-    print(base)
-  dev.off()
-```
-
-```{r PriorKSDists, fig.cap="Prior parameter values should be set so that, given minimal monitoring data ('All' scenario: 1 year of 1 collar per year), the posterior survival distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022].
-Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 45 collars per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local survival, rather than priors from the national model.
-When the multiplier on the prior standard deviation of the survival intercept (lse: $\\beta^S_0$ in Table @ref(tab:priors)) is 5, the prior is informative enough to meet the former condition, and vague enough to meet the latter.
-Given lse=5, low values of the mean standard deviations of the prior distribution of the random effect of year on survival (sse=0.034784, ssv=0.01, $\\sigma^2_S$ in Table @ref(tab:priors)) yield priors that meet both conditions (Figs @ref(fig:priorAllFits) and @ref(fig:priorLowFits)).
-Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
-knitr::include_graphics(here::here("figs/PriorKSDists.png"))
-
-#TO DO: figure out what to do about citations and references in figure captions. Or just fill them in manually at a later stage.
-
-#Note: file wouldn't knit for Anna without access to all the newest png files below, so just commenting out the problem ones for now
-```
-
-```{r PriorAllFits, fig.cap="Differences between posterior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
-In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0.
-Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
-# COMMENTED OUT FOR NOW CAUSE I DIDN"T HAVE ACCESS
-knitr::include_graphics(here::here("figs/SurvsQStarts1ssv0.03.png"))
-```
-
-```{r PriorLowFits, fig.cap="Posterior predictive survival distributions from the Bayesian model ('local') assuming intensive monitoring (15 years of 45 collars per year) of populations from the 2.5% quantile of the national model.
-In these examples, the standard deviation the random effect prior (ssv: $\\sigma^2_S$ in Table @ref(tab:priors)) is 0.03, and  anthropogenic disturbance is 0.
-Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
-# COMMENTED OUT FOR NOW CAUSE I DIDN"T HAVE ACCESS
-knitr::include_graphics(here::here("figs/SurvsQStarts45lowssv0.03.png"))
-```
-
-```{r PriorAllFitsAnthro90, fig.cap="Differences between posterior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming minimal monitoring (1 years of 1 collar per year) of populations from the 50% quantile of the national model.
-In these examples, anthropogenic disturbance is 90%.
-When the multiplier on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) is 3, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model.
-Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts1Anthro90.png"))
-```
-
-```{r PriorHighFitsAnthro90, fig.cap="Posterior predictive survival distributions from the Bayesian model, assuming intensive monitoring (15 years of 45 collars per year) of populations from the 97.5% quantile of the national model.
-In these examples, anthropogenic disturbance is 90%, and the effect of anthropogenic disturbance varies from none (sS=0) through expected (sS=1) to double the expected slope (sS=2).
-All considered multipliers on the prior standard deviation of the survival slope (bse: $\\beta^S_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.
-Bands show 95% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/SurvsQStarts45Anthro90high.png"))
-```
-
-```{r PriorKSDistsRec, fig.cap="Prior parameter values should be set so that, given no monitoring data ('All' scenario), the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model, measured here by Komogorov-Smirnov distance [ks.test function in @R_R_2022].
-Setting the multiplier on the prior standard deviation of the recruitment intercept (lre: $\\beta^R_0$ in Table @ref(tab:priors)) to 3, and values of the mean and standard deviations of the prior distribution of the random effect of year on recruitment to sre=0.23 and srv=0.22, respectively ($\\sigma^2_R$ in Table @ref(tab:priors)) allows us to meet this condition (Fig @ref(fig:priorAllFitsRec)).
-Priors should also be vague enough so that, given intensive monitoring of populations where recruitment is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 135 cows per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local recruitment, rather than priors from the national model.
-When the multiplier on the prior standard deviation of the recruitment intercept (lre: $\\beta^R_0$ in Table @ref(tab:priors)) is 1, the prior is not sufficiently vague to meet the latter condition (Fig @ref(fig:PriorLowFitsRec)).
-Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
-knitr::include_graphics(here::here("figs/PriorKSDistsRec.png"))
-```
-
-```{r PriorAllFitsRec, fig.cap="Differences between posterior predictive recruitment distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no monitoring.
-In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0.
-Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
-knitr::include_graphics(here::here("figs/RecrQStarts1srv0.22.png"))
-```
-
-```{r PriorLowFitsRec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model ('local') assuming intensive monitoring (135 cows per year for 15 years) of populations from the 2.5% quantile of the national model.
-In these examples, the standard deviation the random effect prior (srv: $\\sigma^2_R$ in Table @ref(tab:priors)) is 0.22, and  anthropogenic disturbance is 0.
-Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/RecrQStarts45lowsrv0.22.png"))
-```
-
-```{r PriorAllFitsAnthro90Rec, fig.cap="Differences between posterior recruitment predictive distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no monitoring.
-In these examples, anthropogenic disturbance is 90%.
-When the multiplier on the prior standard deviation of the slope of the relationship between recruitment and anthropogenic disturbance (bre: $\\beta^R_a$ in Table @ref(tab:priors)) is 4, the posterior recruitment distribution is similar to the distribution of simulated outcomes from the national model.
-Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
-knitr::include_graphics(here::here("figs/RecrQStarts1Anthro90.png"))
-```
-
-```{r PriorHighFitsAnthro90Rec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model, assuming intensive monitoring (15 years of 135 collars per year) of populations from the 97.5% quantile of the national model.
-In these examples, anthropogenic disturbance is 90%, and the effect of anthropogenic disturbance varies from none (rS=0) through expected (rS=1) to double the expected slope (rS=2).
-All multipliers on the prior standard deviation of the survival slope (bre: $\\beta^R_a$ in Table @ref(tab:priors)) are sufficiently vague to reflect local variation among populations.
-Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/RecrQStarts45Anthro90high.png"))
-```
-
-\newpage
-
 ### Colophon

### 2023-08-23 analysis/scripts/bayesianModelExamples.R

@@ -7,9 +7,9 @@ library(RColorBrewer)
 
 baseDir <- "."
 
-simBig<-getSimsNational(replicates=3000,forceUpdate=T) #If called with default parameters, use saved object to speed things up.
+simBig<-getSimsNational(replicates=3000,forceUpdate=T)
 
-monitoringScns = data.frame(obsYears=c(1,15),collarCount=c(1,30),cowMult=c(6),collarInterval=c(1),assessmentYrs=c(3))
+monitoringScns = data.frame(obsYears=c(1,15),collarCount=c(0,30),cowMult=c(6),collarInterval=c(1),assessmentYrs=c(3))
 stateScns = data.frame(obsAnthroSlope=c(2),projAnthroSlope=c(2))
 stateScns = merge(stateScns,data.frame(rep=seq(1:1)))
 

### 2023-08-28 analysis/paper/paper.Rmd
ongoing writing

@@
 Demographic projections derived from local range-specific data are also uncertain because few places have enough long-term monitoring data to characterize local demographic-disturbance relationships.
-We present a simple Bayesian integrated population model that integrates prior information from national analysis of demographic-disturbance relationships with available local demographic data to reduce uncertainty in population viability projections.
+We present a simple Bayesian population model that integrates prior information from national analysis of demographic-disturbance relationships with available local demographic data to reduce uncertainty in population viability projections.
 We combine this model with methods for simulating local population dynamics and monitoring to show how monitoring requirements depend on landscape condition.
@@
 Collars lost to mortality may be replaced each year ($o=1$), or there may be a 3 year gap between deployments ($o=4$).
-We focus on exploring a set of scenarios in which anthropogenic disturbance, measured as the percentage of area within 500 m of disturbance [@johnson_science_2020;@ECCC_scientific_2011], increases over time from a variety of starting points (Table \@ref(tab:monitoringScenarios)).
+We focus on exploring a set of scenarios in which anthropogenic disturbance, measured as the percentage of area within 500 m of disturbance [@johnson_science_2020;@eccc_scientific_2011], increases over time from a variety of starting points (Table \@ref(tab:monitoringScenarios)).
 Scenarios in which anthropogenic disturbance does not increase over time are included in Supplement D.
@@
 
+Re paragraph on costs of monitoring.
+The reality is that costs vary with remoteness and the difficulty of finding animals to collar (which also depends on how many collars are deployed).
+Since costs are so context-specific, there isn't a good way to incorporate them into this general analysis.
+It would be interesting to work through a case study (i.e.
+Newfoundland) informed by context-specific costs, and consider how incorporation of those costs may change the relative value of alternative monitoring scenarios.
+
+Biases in composition survey estimates also vary with vegetation cover (i.e.
+the difficulty of seeing animals) and the care taken to correctly identify the sex of unknown animals.
+Misidentification rates could be estimated from discrepancies among results from multiple independent observers analyzing photos of animal groups.
+
+The tool we have developed is a proof of concept, designed to demonstrate possibilities and help users develop a general understanding of the relative value of alternative monitoring strategies for boreal caribou.
+In current form the tool will have limited value for managers who have some local data and wish to know the value of collecting more, because it is designed to either integrate real local data or simulated local data, but not both.
+Development of a tool designed to help decision makers assess the relative value of strategies for additional monitoring given their existing local data and national-demographic disturbance relationships would be a useful followup to this work.
+
+A tool that integrates real local data with national demographic disturbance relationships can also be used to provide consistent and transparent assessments of the strength of evidence for local deviations from the expected mean of national demographic-disturbance relationships.
+Help distinguish real variation among populations from apparent variation that can be attributed to lack of local data.
+
 # Acknowledgements

### 2023-08-30 analysis/paper/paper.Rmd
chipping away at writing.

@@
 Collars lost to mortality may be replaced each year ($o=1$), or there may be a 3 year gap between deployments ($o=4$).
-We focus on exploring a set of scenarios in which anthropogenic disturbance, measured as the percentage of area within 500 m of disturbance [@johnson_science_2020;@eccc_scientific_2011], increases over time from a variety of starting points (Table \@ref(tab:monitoringScenarios)).
+We focus on exploring a set of scenarios in which anthropogenic disturbance, measured as the percentage of area within 500 m of disturbance [@johnson_science_2020;@eccc_scientific_2011], increases over time from a variety of starting points (Table \@ref(tab:monitoringScenarios), Figure \@ref(fig:figDisturbance)).
 Scenarios in which anthropogenic disturbance does not increase over time are included in Supplement D.
@@
 
+In order to assess whether a monitoring program is likely sufficient to inform decisions, it is important to select outcome measures that are relevant to decision making [@gregory_structured_2012].
+For boreal caribou and other endangered species, it is important to know whether a population is declining or not.
+If uncertainty about population status is low, additional investment in reducing uncertainty about the population growth rate may not be necessary or warranted.
+Thus, in this analysis our outcome of interest is how often the Bayesian posterior mean correctly predicts the true population status (increasing or decreasing).
+To calculate this outcome for each disturbance and monitoring scenario (Figure \@ref(fig:figDisturbance)) we fit the Bayesian model to 200 example population trajectories (e.g.
+Figure \@ref(fig:exampleTrajectories)).
+The probability of correct status assessment for that scenario is the percentage of these 200 examples in which the Bayesian posterior mean population growth rate and the true population growth rate are both above or below 1.
+
 ## User interface and open, flexible, reuseable implementation
@@
 
-```{r figStatusErrors, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure 1 for scenarios).
-Probabilities are calculated for 200 example population trajectories (see example trajectories in Figure 2).
+```{r figStatusErrors, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure \\@ref(fig:figDisturbance) for scenarios).
+Probabilities are calculated for 200 example population trajectories (see example trajectories in Figure \\@ref(fig:exampleTrajectories)).
 For each example population, the status assessment is correct if the the Bayesian posterior mean and the true population growth rate are both less than or greater than one.
@@
 Since this kind of aspatial model will mostly be using small population sizes in order to see any effects, we'll jsut have to acknowledge that a caveat of this is that we're not fully able to explore all the population size effects with this model, but generally Josie has found little impact of initial population size to the results.
+
+Include some commentary on addressing limitations of aspatial demographic models with spatial models.
+Not a panacea.
+It is possible (cite examples) to develop spatially explicit models of movement and adult mortality risk - though sample size is a challenge.
+Recruitment, which is more variable among ranges and with anthropogenic disturbance, is difficult to characterize at finer spatial scales because it depends on the conditions experienced by mothers over years as they move across the range, and on spatial variation in calf mortality risk that is not measured directly.
 

### 2023-09-05 analysis/paper/supplement.Rmd

---
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      number_sections: yes
bibliography: CaribouDemographicModel.bib
csl: "../templates/journal-of-forest-ecology-and-management.csl" # Insert path for the bib-style

editor_options: 
  chunk_output_type: inline
---

# (APPENDIX) Supplementary Material for National demographic-disturbance relationships and local monitoring both reduce uncertainty in projections of boreal caribou population viability: tools to support monitoring decisions {-}

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figs/",
  dpi = 300
)
```

```{r setup2}
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(purrr)
library(ggpubr)
library(MetBrewer)
library(here)
library(tmap)
library(sf)
library(caribouMetrics)

devtools::load_all(here::here())

#set paths
data_path_raw <- "analysis/data/raw_data/"
data_path_drvd <- "analysis/data/derived_data"

pal_nm <- "Redon"

fig_widths <- c(min = 1.18, single = 3.543, mid = 5.51, two = 7.48)

runScenarios = F 
doPriorCalibration=F #Note the prior calibration is time consuming.

resultScenario = "S6"
```

# Definitions of variables and parameters in the Bayesian population model

Table A1. Variables and parameters in the Bayesian population model and the national demographic model.

|**Symbol**|**Units**|**Description**|
|:-----|-----|:-----------|
|**Indices & constants**|
|$t$|year|Year.|
|$m$|month|Month.|
|$i$|ID|Individual animal identifier.|
|**Observed**|
|$A_t$|%|Buffered anthropogenic disturbance.|
|$F_t$|%|Fire disturbance.| 
|$\hat{J}_t$|animals|Calves observed in calf:cow survey.|
|$\hat{W}_t$|animals|Cows observed in calf:cow survey.|
|$\hat{T}_t$|animals|Number of collared cows.|
|$\hat{S}_t$|probability|Kaplan-Meier estimate of observed adult female survival.|
|$\tau_t$||Precision of Kaplan-Meier survival estimate.|
|$\hat{I}_{i,t,m}$|0 or 1|Indicator of whether animal $i$ is dead or alive.|
|**Estimated state**|
|$R_t$|rate|Recruitment (calves per cow).|
|$S_t$|probability|Adult female survival.|
|$N_t$|animals|Adult female population size.|
|$\bar{\lambda}_{y,t}$|rate|Average population growth over $y$ years.|
|$X_t$|rate|Recruitment adjusted for sex ratio and (optionally) delayed age of first reproduction.|
|$W_t$|animals|Surviving adult females before recruitment.|
|$J_t$|animals|New recruits to adult age class.|
|$\lambda_t$|rate|Annual population growth.|
|**Estimated Parameters**|
|$\beta^R_0$||Recruitment intercept.|
|$\beta^R_a$||Anthropogenic disturbance slope for recruitment.|
|$\beta^R_f$||Fire slope for recruitment.|
|$\sigma^2_{R}$||Variance in random effect of year on recruitment.|
|$\beta^S_0$||Survival intercept.|
|$\beta^S_a$||Anthropogenic disturbance slope for survival.|
|$\sigma^2_{S}$||Variance in random effect of year on survival.|
|**True state**|
|$\dot{R}_t$|rate|Recruitment (calves per cow).|
|$\dot{S}_t$|probability|Adult female survival.|
|$\dot{N}_t$|animals|Adult female population size.|
|$\dot{X}_t$|rate|Recruitment adjusted for sex ratio and (optionally) delayed age of first reproduction.|
|$\dot{W}_t$|animals|Surviving adult females before recruitment.|
|$\dot{J}_t$|animals|New recruits to adult age class.|
|**True demographic parameters**|
|$N_0$|animals|Initial adult population size.|
|$\dot{\beta}^R_0$||Recruitment intercept.|
|$\dot{\beta}^R_a$||Anthropogenic disturbance slope for recruitment.|
|$\dot{\beta}^R_f$||Fire slope for recruitment.|
|$\dot{\beta}^S_0$||Survival intercept.|
|$\dot{\beta}^S_a$||Anthropogenic disturbance slope for survival.|
|$p_0=1$||Maximum recruitment multiplier.|
|$p_k=0.6$||Recruitment multiplier at carrying capacity.|
|$k=100$||Carrying capacity multiplier.|
|$b=4$||Density dependence shape parameter.|
|$a=1$||Allee effect parameter.|
|$r_{max}=1.3$||Maximum population growth rate.|
|$\nu_R=0.46$||Coefficient of variation in $\dot{R}_t$ among years.|
|$\nu_S=0.08696$||Coefficient of variation in $\dot{S}_t$ among years.|
|$l_R=0$||Minimum value for $\dot{R}_t$.|
|$l_S=0$||Minimum value for $\dot{S}_t$.|
|$h_R=0$||Maximum value for $\dot{R}_t$.|
|$h_S=0$||Maximum value for $\dot{S}_t$.|
|**Observation model parameters**|
|$d=1,2,4,8,16, \text{ or } 24$|years|Monitoring duration.|
|$n=15,30, \text{ or } 60$|animals|Target number of collared females.|
|$o=1 \text{ or } 4$|years|Years between collar deployments.|
|$y=3$|years|Assessment period for population growth rate.|
|$w=3,6,\text{ or } 9$|ratio|The apparent number of adult females per collared animal in composition survey.|
|$q=\text{uniform}(0,0.6)$|ratio|True ratio of young bulls to adult females in composition survey.|
|$u=\text{uniform}(0,0.2)$|probability|Probability of misidentifying young bulls as adult females and vice versa in composition survey.|
|$z=\text{uniform}(0,0.2)$|probability|Probability of missing calves in composition survey.|
|$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}$||Bias correction term (Appendix X).|

# Model of bias in recruitment estimates from calf:cow surveys

TO DO: get photo of calf:cow survey group. Label classes of animals to illustrate meaning of T, U, V, W, O, P, J, K.

We assume each group of animals in a calf:cow composition survey contains one or more collared adult females ($T$), and may also include: uncollared adult females misidentified as young bulls or unknown sex ($U$); correctly identified uncollared adult females ($V$); young bulls correctly identified as male or unknown sex ($O$); young bulls misidentified as uncollared adult females ($P$); observed calves ($J$); and unobserved calves ($K$). The apparent number of adult females in the group is $T+V+P=Tw$, where $w$ is a multiplier that defines the apparent number of adult females as a function of the number of collared animals. The ratio of young bulls to uncollared adult females in the group is:
$$q = \frac{P+O}{U+V}.$$
Assuming an equal probability $u$ of misidentifying young bulls as adult females and vice versa, we get $V=(U+V)(1-u)$ and $P=(O+P)u$. Given a probability $z$ of missing calves, we get $J=(J+K)(1-z)$.

Our objective is to model the sex and bias-corrected recruitment rate $X=\frac{J+K}{2(T+U+V)}$ as a function of the observed calf:cow ratio $R=J/(T+V+P)$, the cow multiplier $w$, the ratio of young bulls to adult females $q$, and the misidentification probabilities $u$ and $z$. We start by solving for $T+U+V$ as a function of $q,w,u$ and $T$. Recognize that $P=Tw-T-V$, $U+V=V/(1-u)$, and $P+O=P/u$ to write $q$ as 
$$q=\frac{Tw-T-V}{uV/(1-u)}.$$
Rearrange to get 
$$V=\frac{T(w-1)(1-u)}{qu+1-u}.$$
Recognize that $U=Vu/(1-u)$ to write $T+U+V$ as a function of $q,w,u$ and $T$: 
$$T+U+V=T\frac{qu+w-u}{qu+1-u}.$$
Recognize that the number of observed calves $J$ is the product of the apparent recruitment rate and the apparent number of adult females $J=RTw$, and that therefore $J+K=RTw/(1-z)$ to rewrite the bias corrected recruitment rate $X=\frac{J+K}{2(T+U+V)}$ as a function of $w,u,z$ and $R$:
$$X=R\frac{w(1+qu-u)}{2(w+qu-u)(1-z)}.$$
For simplicity, we write $X$ as a function of a bias correction term $c$: 
$$c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}; X=cR/2$$

If we also adjust for delayed age at first reproduction (DeCesare et al. 2012; Eacker et al. 2019), the adjusted recruitment rate becomes $$X=\frac{cR/2}{1+cR/2}.$$

```{r biasPlot, fig.cap="Variation in the sex and bias corrected recruitment rate $X$ with the cow multiplier $w$, the ratio of young bulls to adult females $q$, and the adult misidentification probability $u$, and the probability of missing a calf $z$. If only collared females are counted ($w=1$) and there is a risk of missing calves $z>0$ then the apparent recruitment rate $R$ will be lower than $X$ (first column). When uncollared animals are counted ($w>1$), the risk of misidentifying adults $u$ inflates the apparent recruitment rate $R$ as on average more cows are misidentified as bulls than vice versa, but this effect is smaller than the effect of missing calves $z$. Increasing the proportion of bulls in the group $q$ reduces the effect of adult misidentification $u$.", fig.height=6,cache=TRUE}
library(ggplot2)
#Notes
#q: assume more females than males in calf:cow group, so q<1

p=expand.grid(delay=c(T,F),q=c(0,0.3,0.6),w=c(1,3,6),
              u=c(0,0.1,0.2),z=seq(0,0.2,length.out=3),R=seq(0.15,0.6,length.out=10))

getX<-function(R,delay,q,w,u,z){
  c = compositionBiasCorrection(w,q,u,z)
  X = 0.5*R*c
  X=X*(1-delay)+delay*X/(1+X)
  return(X)
}

p$X = getX(p$R,p$delay,p$q,p$w,p$u,p$z)

p$grp = paste(p$u,p$z)
p$u=as.factor(p$u)
p$z=as.factor(p$z)

base = ggplot(subset(p,!delay),aes(x=R,y=2*X/R,col=z,linetype=u,group=grp))+
  geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()+ylab("2X/R=c, without delayed reproduction")

base2 = ggplot(subset(p,delay),aes(x=R,y=2*X/R,col=z,linetype=u,group=grp))+
  geom_hline(yintercept=1)+geom_line()+facet_grid(q~w,labeller=label_both)+theme_bw()+ylab("2X/R, with delayed reproduction")

ggpubr::ggarrange(base, base2, labels = "",
                  ncol = 1, vjust = 1)

```

```{r biasDistribution,fig.cap="In the Bayesian model, uncertainty about the value of the bias correction term $c$ (grey bars) is approximated with a Log-normal distribution (red). In these examples, the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2. We assume that the apparent number of adult females per collared animal $w$ is known.", cache=TRUE}

cr=expand.grid(w = c(1,2,3,4,5,6,7,8,9),rep=seq(1,10000))
nr=nrow(cr)
cr$c = compositionBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2))
cs = compositionBiasCorrection(q=runif(nr,0,0.6),w=cr$w,u=runif(nr,0,0.2),z=runif(nr,0,0.2),approx=T)

cr = merge(cr,cs)
cr$ca= exp(rnorm(nrow(cr),mean=cr$mu,sd=cr$sig2^0.5))

base = ggplot(cr,aes(x=c))+geom_histogram(aes(y=..density..))+
  facet_wrap(~w,nrow=3,labeller=label_both)+theme_bw()+
  geom_density(aes(x=ca),alpha=.2, fill="#FF6666",col = "#FF6666")
plot(base)

```

# Calibration of Priors

```{r priorCalibration, cache=TRUE,echo=FALSE, message=FALSE,results='hide'}
if(doPriorCalibration){

  source(here::here("analysis/scripts/calibratePriors.R"))
}else{
  KS<-read.csv(here::here("tabs/SurvKSAll.csv"))
  KSRec<-read.csv(here::here("tabs/RecKSAll.csv"))

}
  #library(ggplot2);fig_widths=data.frame(two=7)
  KS$SigmaMean = KS$sInterannualVar
  KS$SigmaSD = KS$sInterannualVarSE
  KS=subset(KS,scn=="All")
  png(here::here("figs/PriorKSDists.png"),
    height = 3, width = fig_widths["two"], units = "in",res=600)
    base = ggplot(KS,aes(x=sIntSEMod,y=KSDistance,color=SigmaMean))+
    geom_point()+facet_wrap(~SigmaSD,labeller="label_both")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("sIntSEMod: multiplier on SD of survival intercept")
    print(base)
  dev.off()

  KSRec$SigmaMean = KSRec$rInterannualVar
  KSRec$SigmaSD = KSRec$rInterannualVarSE
  KSRec=subset(KSRec,scn=="All")
  png(here::here("figs/PriorKSDistsRec.png"),
    height = 3, width = fig_widths["two"], units = "in",res=600)
    base = ggplot(KSRec,aes(x=rIntSEMod,y=KSDistance,color=SigmaMean))+
    geom_point()+facet_wrap(~SigmaSD,labeller="label_both")+theme_bw()+ylab("Kolmogorov-Smirnov Distance")+xlab("rIntSEMod: multiplier on SD of recruitment intercept")
    print(base)
  dev.off()
```

```{r PriorKSDists, fig.cap="Similarity of the prior survival distribution to the distribution of simulated outcomes from the national model measured by Kolmogorov-Smirnov distance [ks.test function in @r_core_team_r_2022]. The distributions are similar when the multiplier on the prior standard deviation of the survival intercept (sIntSEMod: $\\beta^S_0$ in Table 2) is 5 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\\sigma^2_S$ in Table 2). Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
knitr::include_graphics(here::here("figs/PriorKSDists.png"))

```

```{r PriorAllFits, fig.cap="Differences between prior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring. In these examples, the standard deviation of the random effect prior (SigmaSD: $\\sigma^2_S$ in Table 2) is 0.03, and anthropogenic disturbance is 0. Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}

knitr::include_graphics(here::here("figs/SurvInterceptAllsInterannualVarSE0.03.png"))
```

```{r PriorLowFits, fig.cap="Posterior predictive survival distributions from the Bayesian model ('local') assuming intensive monitoring (15 years of 60 collars per year) of populations from the 2.5 \\% quantile of the national model. In these examples, the standard deviation of the random effect prior (SigmaSD: $\\sigma^2_S$ in Table 2) is 0.03, and  anthropogenic disturbance is 0. Bands show 95 \\% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates. When the multiplier on the prior standard deviation of the survival intercept (sIntSEMod: $\\beta^S_0$ in Table 2) is too low (<5) the insufficiently vague priors substantially influence the posterior distributions, yielding poor local parameter estimates.", cache=TRUE}

knitr::include_graphics(here::here("figs/SurvsurvInterceptLowsInterannualVarSE0.03.png"))
```

```{r PriorAllFitsAnthro90, fig.cap="Differences between prior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring. In these examples, anthropogenic disturbance is 90\\%. When the multiplier on the prior standard deviation of the survival slope (sSlopeSEMod: $\\beta^S_a$ in Table 2) is 3, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model. Bands show 95\\% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
knitr::include_graphics(here::here("figs/SurvSlopeAll.png"))
```

```{r PriorHighFitsAnthro90, fig.cap="Posterior predictive survival distributions from the Bayesian model, assuming intensive monitoring (15 years of 60 collars per year) of populations from the 97.5 \\% quantile of the national model. In these examples, anthropogenic disturbance is 90\\%, and the effect of anthropogenic disturbance varies from none (sSlopeMod=0) through expected (sSlopeMod=1) to double the expected slope (sSlopeMod=2). All considered multipliers on the prior standard deviation of the survival slope (sSlopeSEMod: $\\beta^S_a$ in Table 2) are sufficiently vague to reflect local variation among populations. Bands show 95 \\% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
knitr::include_graphics(here::here("figs/SurvSlopeHigh.png"))
```

Prior parameter values should be set so that, given no monitoring data, the prior survival distribution is similar to the distribution of simulated outcomes from the national model in the case of no anthropogenic disturbance. This condition is met when the multiplier on the prior standard deviation of the survival intercept (sIntSEMod: $\beta^S_0$ in Table 2) is 5 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\sigma^2_S$ in Table 2) (Figs \@ref(fig:PriorKSDists) and \@ref(fig:PriorAllFits)). Priors should also be vague enough so that, given intensive monitoring of populations where survival is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 60 collars per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local survival, rather than priors from the national model. Setting the mean and standard deviations of the prior distribution of the random effect of year on survival to moderately low values (SigmaMean=0.034784, SigmaSD=0.03, $\sigma^2_S$ in Table 2) allows both conditions to be met (Figs \@ref(fig:PriorKSDists), \@ref(fig:PriorAllFits) and \@ref(fig:PriorLowFits)).

```{r PriorKSDistsRec, fig.cap="Similarity of the prior recruitment distribution to the distribution of simulated outcomes from the national model measured by Kolmogorov-Smirnov distance [ks.test function in @r_core_team_r_2022]. The distributions are similar when the multiplier on the prior standard deviation of the survival intercept (rIntSEMod: $\\beta^R_0$ in Table 2) is 4 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\\sigma^2_R$ in Table 2). Anthropogenic disturbance is set to 0 in these scenarios." , cache=TRUE}
knitr::include_graphics(here::here("figs/PriorKSDistsRec.png"))
```

```{r PriorAllFitsRec, fig.cap="Differences between prior predictive recruitment distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring. In these examples, the standard deviation of the random effect prior (SigmaSD: $\\sigma^2_R$ in Table 2) is 0.22, and anthropogenic disturbance is 0. Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
knitr::include_graphics(here::here("figs/RecInterceptAllrInterannualVarSE0.22.png"))
```

```{r PriorHighFitsRec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model ('local') assuming intensive monitoring (9*60 cows per year for 15 years) of populations from the 97.5% quantile of the national model.  In these examples, the standard deviation the random effect prior (SigmaSD: $\\sigma^2_R$ in Table 2) is 0.22, and  anthropogenic disturbance is 0. Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
knitr::include_graphics(here::here("figs/RecInterceptHighrInterannualVarSE0.22.png"))
```

```{r PriorAllFitsAnthro90Rec, fig.cap="Differences between prior predictive survival distributions from the Bayesian model ('local') and simulated outcomes from the national model ('national'), assuming no local monitoring. In these examples, anthropogenic disturbance is 90\\%. When the multiplier on the prior standard deviation of the recruitment slope (rSlopeSEMod: $\\beta^R_a$ in Table 2) is 4, the posterior survival distribution is similar to the distribution of simulated outcomes from the national model. Bands show 95\\% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.", cache=TRUE}
knitr::include_graphics(here::here("figs/RecSlopeAll.png"))
```

```{r PriorHighFitsAnthro90Rec, fig.cap="Posterior predictive recruitment distributions from the Bayesian model, assuming intensive monitoring (15 years of 60 collars per year) of populations from the 97.5 \\% quantile of the national model. In these examples, anthropogenic disturbance is 90\\%, and the effect of anthropogenic disturbance varies from none (rSlopeMod=0) through expected (rSlopeMod=1) to double the expected slope (rSlopeMod=2). All considered multipliers on the prior standard deviation of the recruitment slope >1 (rSlopeSEMod: $\\beta^R_a$ in Table 2) are sufficiently vague to reflect local variation among populations. Bands show 95 \\% posterior predictive intervals, black dots show true survival rates from the simulated population, and open triangles show observed survival rates.", cache=TRUE}
knitr::include_graphics(here::here("figs/RecSlopeHigh.png"))
```

Similarly, prior parameter values should be set so that, given no monitoring data, the prior recruitment distribution is similar to the distribution of simulated outcomes from the national model in the case of no anthropogenic disturbance. This condition is met when the multiplier on the prior standard deviation of the recruitment intercept (rIntSEMod: $\beta^R_0$ in Table 2) is 4 and the mean and standard deviations of the prior distribution of the random effect of year on survival are low  (SigmaMean & SigmaSD, $\sigma^2_R$ in Table 2) (Figs \@ref(fig:PriorKSDistsRec) and \@ref(fig:PriorAllFitsRec)). Priors should also be vague enough so that, given intensive monitoring of populations where recruitment is higher or lower than the national average ('High' and 'Low' scenarios: 15 years of 9*60 cows per year from 97.5% and 2.5% quantiles of the national model), posteriors reflect local recruitment, rather than priors from the national model. Setting the mean and standard deviations of the prior distribution of the random effect of year on recruitment to moderately low values (SigmaMean=0.23, SigmaSD=0.22, $\sigma^2_R$ in Table 2) allows both conditions to be met (Figs \@ref(fig:PriorKSDistsRec), \@ref(fig:PriorAllFitsRec) and \@ref(fig:PriorHighFitsRec)).

\newpage

# Alternative disturbance and monitoring scenarios

```{r figDisturbanceAnthroStatic, fig.cap="Anthropogenic disturbance and monitoring scenarios.", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/distScnscmult6ay3aSf0.png")))

```

```{r powercmult3ay3aSf0, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure \\@ref(fig:figDisturbanceAnthroStatic) for for static anthropogenic disturbance scenarios). In these examples there are 3 cows per collared cow in recruitment surveys. See Figure 4 for additional details. ", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/powercmult3ay3aSf0.png")))
```

```{r powercmult6ay3aSf0, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure \\@ref(fig:figDisturbanceAnthroStatic) for static anthropogenic disturbance scenarios). In these examples there are 6 cows per collared cow in recruitment surveys. See Figure 4 for additional details. ", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/powercmult6ay3aSf0.png")))
```

```{r powercmult9ay3aSf0, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure \\@ref(fig:figDisturbanceAnthroStatic) for static anthropogenic disturbance scenarios). In these examples there are 9 cows per collared cow in recruitment surveys. See Figure 4 for additional details. ", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/powercmult9ay3aSf0.png")))
```

```{r powercmult3ay3aSf1, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure 1 for scenarios in which anthropogenic disturbance increases over time). In these examples there are 3 cows per collared cow in recruitment surveys. See Figure 4 for additional details. ", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/powercmult3ay3aSf1.png")))
```

```{r powercmult9ay3aSf1, fig.cap="The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), and monitoring effort (see Figure 1 for scenarios in which anthropogenic disturbance increases over time). In these examples there are 9 cows per collared cow in recruitment surveys. See Figure 4 for additional details. ", cache=TRUE}
knitr::include_graphics(here::here(paste0("figs/",resultScenario,"/powercmult9ay3aSf1.png")))
```

# Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
#if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
#if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2023-09-05 analysis/paper/paper.Rmd
minor text changes and added UI paragraph to methods

@@
 This tool can help users assess monitoring needs and strategies.
-
 keywords: |
@@
   chunk_output_type: inline
+  markdown: 
+    wrap: 72
 ---
@@
 <!-- With the following code you can access and display values from the yml header above. -->
+
 <!-- Link to data here: https://drive.google.com/file/d/179lw6MPK4l3xGmded9kofzJkB6VlJVb9/view?usp=sharing -->
@@
 <!-- The actual document text starts here: -->
+
 # Introduction
 
-NOTE: find working version of intro and discussion [here:] (https://docs.google.com/document/d/1RD7bJ6yhtBJULjJFRJfHzObQscteyObM/edit?usp=sharing&ouid=111541730342945675365&rtpof=true&sd=true).
+NOTE: find working version of intro and discussion [here:]
+(<https://docs.google.com/document/d/1RD7bJ6yhtBJULjJFRJfHzObQscteyObM/edit?usp=sharing&ouid=111541730342945675365&rtpof=true&sd=true>).
 
+```{=html}
 <!--
@@
 -->
-
+```
 # Methods
@@
 
-TO DO: Anna draft this section?
-Marking of caribou cows with telemetry GPS/VHF collars is the traditional approach to estimating demographic parameters (refs/examples, e.g.
-key boreal caribou projections and status documents reliant on these ests, also methods papers like Courtois et al.
-2003, RettieMessier 1998).
+TO DO: Anna draft this section? Marking of caribou cows with telemetry
+GPS/VHF collars is the traditional approach to estimating demographic
+parameters (refs/examples, e.g. key boreal caribou projections and
+status documents reliant on these ests, also methods papers like
+Courtois et al. 2003, RettieMessier 1998).
 
-Adult survival is estimated by tracking the fate of marked individuals (Kaplan-Meier ests) [brief description of field/logistical details - can get details from NBCKC Telemetry writeup (https://www.cclmportal.ca/sites/default/files/2021-10/Telemetry%20draft%20October%2027%202021.pdf), e.g.
-GPS allow for remote detection vs VHF-only require physical relocation of individual; collar sends out a
-mortality signal when no longer active after a predetermined period of time; males only occasionally collared; age of collared individuals is usually unknown]
+Adult survival is estimated by tracking the fate of marked individuals
+(Kaplan-Meier ests) [brief description of field/logistical details - can
+get details from NBCKC Telemetry writeup
+(<https://www.cclmportal.ca/sites/default/files/2021-10/Telemetry%20draft%20October%2027%202021.pdf>),
+e.g. GPS allow for remote detection vs VHF-only require physical
+relocation of individual; collar sends out a mortality signal when no
+longer active after a predetermined period of time; males only
+occasionally collared; age of collared individuals is usually unknown]
 
-Collared females are also used in traditional estimates of recruitment (refs/examples): collared individuals are located by aircraft and then the entire group of caribou is surveyed to estimate the ratio of cows:calves in the group...
-Because this estimate is based on more than just the collared sample of cows, the sample size for recruitment ests is typically larger than for survival (introduce the multiplier parameter).
+Collared females are also used in traditional estimates of recruitment
+(refs/examples): collared individuals are located by aircraft and then
+the entire group of caribou is surveyed to estimate the ratio of
+cows:calves in the group... Because this estimate is based on more than
+just the collared sample of cows, the sample size for recruitment ests
+is typically larger than for survival (introduce the multiplier
+parameter).
 
@@
 
-For this investigation we assume collars are deployed in January, and last for up to 6 years.
-The target number of collared females $n$ varies among scenarios, as does the total duration of the monitoring program (Table \@ref(tab:monitoringScenarios)).
-Collars lost to mortality may be replaced each year ($o=1$), or there may be a 3 year gap between deployments ($o=4$).
-We focus on exploring a set of scenarios in which anthropogenic disturbance, measured as the percentage of area within 500 m of disturbance [@johnson_science_2020;@eccc_scientific_2011], increases over time from a variety of starting points (Table \@ref(tab:monitoringScenarios), Figure \@ref(fig:figDisturbance)).
-Scenarios in which anthropogenic disturbance does not increase over time are included in Supplement D.
+For this investigation we assume collars are deployed in January, and
+last for up to 6 years. The target number of collared females $n$ varies
+among scenarios, as does the total duration of the monitoring program
+(Table \@ref(tab:monitoringScenarios)). Collars lost to mortality may be
+replaced each year ($o=1$), or there may be a 3 year gap between
+deployments ($o=4$). We focus on exploring a set of scenarios in which
+anthropogenic disturbance, measured as the percentage of area within 500
+m of disturbance [@johnson_science_2020; @eccc_scientific_2011],
+increases over time from a variety of starting points (Table
+\@ref(tab:monitoringScenarios), Figure \@ref(fig:figDisturbance)).
+Scenarios in which anthropogenic disturbance does not increase over time
+are included in Supplement D.
 
+```{=html}
 <!--
@@
 -->
-
+```
 ## Simulation of local population dynamics and monitoring
@@
 
-To simulate plausible plausible demographic trajectories for example populations we used a modified version of Johnson et al's [-@johnson_science_2020] two-stage demographic model described by Dyson et al.
-[-@dyson_existing_2022].
-The true number of post-juvenile females that survive from year $t$ to the census $\dot{W}_t$ is binomially distributed with true survival probability $\dot{S}_t$: $\dot{W}_{t} \sim \text{Binomial}(\dot{N}_t,\dot{S}_t)$.
-Maximum potential recruitment rate is adjusted for sex ratio, misidentification biases (Supplement B) and (optionally) delayed age at first reproduction [@eacker_web-based_2019;@decesare_estimating_2012]: $$\dot{X}_t=\frac{c\dot{R}_t/2}{1+c\dot{R}_t/2};c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}, $$
-where $w$ is a multiplier that defines the apparent number of adult females in the composition survey as a function of the number of collared animals, $q$ is the ratio of young bulls to adult females, $u$ is the probability of misidentifying young bulls as adult females and vice versa, and $z$ is the probability of missing a calf.
+To simulate plausible demographic trajectories for example populations
+we used a modified version of Johnson et al's [-@johnson_science_2020]
+two-stage demographic model described by Dyson et al.
+[-@dyson_existing_2022]. The true number of post-juvenile females that
+survive from year $t$ to the census, $\dot{W}_t$, is binomially
+distributed with true survival probability $\dot{S}_t$:
+$\dot{W}_{t} \sim \text{Binomial}(\dot{N}_t,\dot{S}_t)$. Maximum
+potential recruitment rate is adjusted for sex ratio, misidentification
+biases (Supplement B) and (optionally) delayed age at first reproduction
+[@eacker_web-based_2019; @decesare_estimating_2012]:
+$$\dot{X}_t=\frac{c\dot{R}_t/2}{1+c\dot{R}_t/2};c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}, $$
+where $w$ is a multiplier that defines the apparent number of adult
+females in the composition survey as a function of the number of
+collared animals, $q$ is the ratio of young bulls to adult females, $u$
+is the probability of misidentifying young bulls as adult females and
+vice versa, and $z$ is the probability of missing a calf.
 
-Realized recruitment rate varies with population density [@lacy_vortex_2017], and the number of juveniles recruiting to the post-juvenile class at the census is a binomially distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate: $$\dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{N_0k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}).$$ Given parameters in Table A1, recruitment rate is lowest $(0.5\dot{X}_t)$ when $\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$ at intermediate population sizes, and declines to $0.6\dot{X}_t$ as the population reaches carrying capacity of $k=100$ times the initial population size.
-The post-juvenile female population in the next year includes both survivors and new recruits: $\dot{N}_{t+1}=\text{min}(\dot{W}_t+\dot{J}_t,r_{max}\dot{N}_t)$.
+Realized recruitment rate varies with population density
+[@lacy_vortex_2017], and the number of juveniles recruiting to the
+post-juvenile class at the census is a binomially distributed function
+of the number of surviving post-juvenile females and the adjusted
+recruitment rate:
+$$\dot{J}_{t} \sim \text{Binomial}(\dot{W}_t,\dot{X}_t[p_0-(p_0-p_k)(\frac{\dot{W}_t}{N_0k})^b]\frac{\dot{W}_t}{\dot{W}_t+a}).$$
+Given parameters in Table A1, recruitment rate is lowest
+$(0.5\dot{X}_t)$ when $\dot{N}_t=1$, approaches a maximum of $\dot{X}_t$
+at intermediate population sizes, and declines to $0.6\dot{X}_t$ as the
+population reaches carrying capacity of $k=100$ times the initial
+population size. The post-juvenile female population in the next year
+includes both survivors and new recruits:
+$\dot{N}_{t+1}=\text{min}(\dot{W}_t+\dot{J}_t,r_{max}\dot{N}_t)$.
 
-Interannual variation in survival and recruitment is modelled using truncated beta distributions [rtrunc function; @novomestky_package_2016]: $\dot{R}_t \sim \text{TruncatedBeta}(\bar{R}_t,\nu_R,l_R,h_R); \dot{S}_t \sim \text{TruncatedBeta}(\bar{S}_t,\nu_S,l_S,h_S)$.
-Coefficients of variation among years $(\nu_R,\nu_S)$ and maximum/minimum values $l_R,h_R,l_S,h_S$ for recruitment and survival are given in Table A1.
-Expected recruitment ($\bar{R}_t$) and survival ($\bar{S}_t$) vary with disturbance according the beta regression models estimated by Johnson et al.
-[@johnson_science_2020]: $$ \bar{R}_t \sim Beta(\mu^R_t,\phi^R);log(\mu^R_t)=\dot{\beta^R_0}+\dot{\beta^R_a}A_t+\dot{\beta}^R_fF_t,$$
+Interannual variation in survival and recruitment is modelled using
+truncated beta distributions [rtrunc function;
+@novomestky_package_2016]:
+$\dot{R}_t \sim \text{TruncatedBeta}(\bar{R}_t,\nu_R,l_R,h_R); \dot{S}_t \sim \text{TruncatedBeta}(\bar{S}_t,\nu_S,l_S,h_S)$.
+Coefficients of variation among years $(\nu_R,\nu_S)$ and
+maximum/minimum values $l_R,h_R,l_S,h_S$ for recruitment and survival
+are given in Table A1. Expected recruitment ($\bar{R}_t$) and survival
+($\bar{S}_t$) vary with disturbance according the beta regression models
+estimated by Johnson et al. [@johnson_science_2020]:
+$$ \bar{R}_t \sim Beta(\mu^R_t,\phi^R);log(\mu^R_t)=\dot{\beta^R_0}+\dot{\beta^R_a}A_t+\dot{\beta}^R_fF_t,$$
 $$\bar{S}_t \sim (46\times Beta(\mu^S_t,\phi^S)-0.5)/45;log(\mu^S_t)=\dot{\beta^S_0}+\dot{\beta^S_a}A_t.$$
-$\phi^R \sim \text{Normal}(19.862,2.229)$ and $\phi^S \sim \text{Normal}(63.733,8.311)$ are precisions of the Beta distributed errors [@ferrari_beta_2004].
-At the beginning of a simulation for an example population, regression coefficient values are sampled from Gaussian distributions (see Table \@ref(tab:priors)) and the population is assigned to quantiles of the Beta error distributions for survival and recruitment.
-The population remains in these quantiles as disturbance changes over time, so there is substantial persistent variation in recruitment and survival among example among populations (Figure \@ref(fig:exampleTrajectories)).
+$\phi^R \sim \text{Normal}(19.862,2.229)$ and
+$\phi^S \sim \text{Normal}(63.733,8.311)$ are precisions of the Beta
+distributed errors [@ferrari_beta_2004]. At the beginning of a
+simulation for an example population, regression coefficient values are
+sampled from Gaussian distributions (see Table \@ref(tab:priors)) and
+the population is assigned to quantiles of the Beta error distributions
+for survival and recruitment. The population remains in these quantiles
+as disturbance changes over time, so there is substantial persistent
+variation in recruitment and survival among example among populations
+(Figure \@ref(fig:exampleTrajectories)).
 
@@
 
-In our simulated populations, mortality risk does not vary over time or among individuals, so the observed state (dead or alive) of a collared animal $i$ in month $m$ and year $t$ depends on the true survival rate $\dot{S}_t$ as follows:
+In our simulated populations, mortality risk does not vary over time or
+among individuals, so the observed state (dead or alive) of a collared
+animal $i$ in month $m$ and year $t$ depends on the true survival rate
+$\dot{S}_t$ as follows:
 $$\hat{I}_{i,t,m} \sim \text{Bernoulli}(\dot{S}^{1/12}_t\hat{I}_{i,t,m-1}).$$
@@
 
-The number of cows in the composition survey $\hat{W}_t$ is given by the number of collared cows $\hat{T}_t$ and the apparent number of adult females per collared female calves observed in the composition survey $w$:
-$$\hat{W}_t = w\hat{T}_t.$$
-The number of observed calves $\hat{J}_t$ also depends on the unadjusted apparent recruitment for the population $\dot{R}_t$:
-$$\hat{J}_t \sim \text{Binomial}(\hat{W}_t,\dot{R}_t).$$
+The number of cows in the composition survey $\hat{W}_t$ is given by the
+number of collared cows $\hat{T}_t$ and the apparent number of adult
+females per collared female observed in the composition survey $w$:
+$$\hat{W}_t = w\hat{T}_t.$$ The number of observed calves $\hat{J}_t$
+also depends on the unadjusted apparent recruitment for the population
+$\dot{R}_t$: $$\hat{J}_t \sim \text{Binomial}(\hat{W}_t,\dot{R}_t).$$
 
@@
 
-Following Eacker et al.
-[-@eacker_web-based_2019], the observed number of calves $\hat{J}_t$ is a binomially distributed function of the observed number of adult female caribou $\hat{W}_t$ and the estimated recruitment rate $R_t$: $$\hat{J}_t \sim \text{Binomial}(\hat{W}_t,R_t).$$
+Following Eacker et al. [-@eacker_web-based_2019], the observed number
+of calves $\hat{J}_t$ is a binomially distributed function of the
+observed number of adult female caribou $\hat{W}_t$ and the estimated
+recruitment rate $R_t$:
+$$\hat{J}_t \sim \text{Binomial}(\hat{W}_t,R_t).$$
 
-We model recruitment probability as a function of anthropogenic disturbance and fire with a log link [@johnson_science_2020;@stewart_climate-informed_2023;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$\log(R_t)=\beta^R_0+\beta^R_a A_t+\beta^R_f F_t+\epsilon^R_t; \epsilon^R_t \sim \text{Normal}(0,\sigma^2_{R}).$$
-Estimated recruitment probability is adjusted for sex ratio, misidentification biases (Supplement B), and age of first reproduction [@eacker_web-based_2019;@decesare_estimating_2012] to get expected recruits per female:
-$$X_t=\frac{cR_t/2}{1+cR_t/2}; c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}.$$
-The composition survey bias correction term $c$ depends on the apparent number of adult females per collared animal $w$, the ratio of young bulls to adult females $q$, the probability of misidentifying young bulls as adult females and vice versa $u$, and the probability of missing a calf $z$ (Supplement B).
-To maintain consistency with Johnson et al.
-[-@johnson_science_2020] we also provide an option to only adjust for sex ratio and misidentification biases $X_t=cR_t/2$.
+We model recruitment probability as a function of anthropogenic
+disturbance and fire with a log link [@johnson_science_2020;
+@stewart_climate-informed_2023; @dyson_existing_2022] and Gaussian
+distributed random year effect [@eacker_web-based_2019]:
+$$\log(R_t)=\beta^R_0+\beta^R_a A_t+\beta^R_f F_t+\epsilon^R_t; \epsilon^R_t \sim \text{Normal}(0,\sigma^2_{R}).$$
+Estimated recruitment probability is adjusted for sex ratio,
+misidentification biases (Supplement B), and age of first reproduction
+[@eacker_web-based_2019; @decesare_estimating_2012] to get expected
+recruits per female:
+$$X_t=\frac{cR_t/2}{1+cR_t/2}; c=\frac{w(1+qu-u)}{(w+qu-u)(1-z)}.$$ The
+composition survey bias correction term $c$ depends on the apparent
+number of adult females per collared animal $w$, the ratio of young
+bulls to adult females $q$, the probability of misidentifying young
+bulls as adult females and vice versa $u$, and the probability of
+missing a calf $z$ (Supplement B). To maintain consistency with Johnson
+et al. [-@johnson_science_2020] we also provide an option to only adjust
+for sex ratio and misidentification biases $X_t=cR_t/2$.
 
-In the Bayesian model, prior uncertainty about the value of the bias correction term $c$ is approximated with a Log-normal distribution (Supplement B).
-Given the apparent number of adult females per collared animal $w$ in each monitoring scenario (Table \@ref(tab:monitoringScenarios)), we calculate the mean and standard deviation of 10,000 samples of $\log{c}$ assuming the ratio of young bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult misidentification probability $u$ varies uniformly between 0 and 0.2, and the probability of missing a calf $z$ varies uniformly between 0 and 0.2.
+In the Bayesian model, prior uncertainty about the value of the bias
+correction term $c$ is approximated with a Log-normal distribution
+(Supplement B). Given the apparent number of adult females per collared
+animal $w$ in each monitoring scenario (Table
+\@ref(tab:monitoringScenarios)), we calculate the mean and standard
+deviation of 10,000 samples of $\log{c}$ assuming the ratio of young
+bulls to adult females $q$ varies uniformly between 0 and 0.6, the adult
+misidentification probability $u$ varies uniformly between 0 and 0.2,
+and the probability of missing a calf $z$ varies uniformly between 0 and
+0.2.
 
-Also following Eacker et al.
-[-@eacker_web-based_2019], Kaplan-Meier estimates of observed adult female survival probability $\hat{S}_t$ and associated precisions $\tau_t$ are estimated from known-fate radio collar data using the survival R package [@therneau_modeling_2000], with precisions for years with no observed mortality estimated using a separate Bayesian model.
-Observed estimated survival is a Gaussian distributed function of the true survival probability: $$\hat{S_t} \sim \text{Normal}(S_t,1/\tau_t).$$
-As in the national demographic model, we model survival probability as an adjusted function of buffered anthropogenic disturbance $A_t$ with a log link [@johnson_science_2020;@stewart_climate-informed_2023;@dyson_existing_2022] and Gaussian distributed random year effect [@eacker_web-based_2019]: $$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a A_t+\epsilon^S_t; \epsilon^S_t \sim \text{Normal}(0,\sigma^2_{S}).$$
-To assess sensitivity of results to the survival estimation method and handle cases in which there is little or no survival data, we also implemented an alternative parametric exponential survival model, wherein the observed state (alive or dead, $\hat{I}_{i,t,m}$) of an individual animal $i$ in year $t$ and month $m$ is Bernoulli distributed [@schaub_integrated_2021]: $$\hat{I}_{i,t,m} \sim \text{Bernoulli}(S^{1/12}_t\hat{I}_{i,t,m-1}).$$
+Also following Eacker et al. [-@eacker_web-based_2019], Kaplan-Meier
+estimates of observed adult female survival probability $\hat{S}_t$ and
+associated precisions $\tau_t$ are estimated from known-fate radio
+collar data using the survival R package [@therneau_modeling_2000], with
+precisions for years with no observed mortality estimated using a
+separate Bayesian model. Observed estimated survival is a Gaussian
+distributed function of the true survival probability:
+$$\hat{S_t} \sim \text{Normal}(S_t,1/\tau_t).$$ As in the national
+demographic model, we model survival probability as an adjusted function
+of buffered anthropogenic disturbance $A_t$ with a log link
+[@johnson_science_2020; @stewart_climate-informed_2023;
+@dyson_existing_2022] and Gaussian distributed random year effect
+[@eacker_web-based_2019]:
+$$S_t=\text{min}(46 \tilde{S_t}-0.5)/45,1); \log(\tilde{S_t})=\beta^S_0+\beta^S_a A_t+\epsilon^S_t; \epsilon^S_t \sim \text{Normal}(0,\sigma^2_{S}).$$
+To assess sensitivity of results to the survival estimation method and
+handle cases in which there is little or no survival data, we also
+implemented an alternative parametric exponential survival model,
+wherein the observed state (alive or dead, $\hat{I}_{i,t,m}$) of an
+individual animal $i$ in year $t$ and month $m$ is Bernoulli distributed
+[@schaub_integrated_2021]:
+$$\hat{I}_{i,t,m} \sim \text{Bernoulli}(S^{1/12}_t\hat{I}_{i,t,m-1}).$$
 
-All regression coefficients are assumed to be Gaussian distributed, with priors derived from the expected values and 95% confidence intervals estimated by Johnson et al (2020) (Table \@ref(tab:priors)).
-We calibrated the prior distributions so that the 95% prior prediction intervals for survival and recruitment from the Bayesian model match the range between the 2.5% and 97.5% quantiles of 1000 simulated survival and recruitment trajectories (Supplement C, Figure \@ref(fig:examplePredictions)).
+All regression coefficients are assumed to be Gaussian distributed, with
+priors derived from the expected values and 95% confidence intervals
+estimated by Johnson et al (2020) (Table \@ref(tab:priors)). We
+calibrated the prior distributions so that the 95% prior prediction
+intervals for survival and recruitment from the Bayesian model match the
+range between the 2.5% and 97.5% quantiles of 1000 simulated survival
+and recruitment trajectories (Supplement C, Figure
+\@ref(fig:examplePredictions)).
+
 ```{r priors, echo=FALSE,message=FALSE}
@@
 
-We assume a simple two-class demographic model, with a census in late winter of surviving females and new recruits to the post-juvenile (aged 1+) class.
-The number of post-juvenile females that survive from year $t$ to the census $W_t$ is binomially distributed with survival probability $S_t$: $W_{t} \sim \text{Binomial}(N_t,S_t)$.
-The number of juveniles recruiting to the post-juvenile class at the census is a Poisson distributed function of the number of surviving post-juvenile females and the adjusted recruitment rate $X_t$: $J_{t} \sim \text{Poisson}(X_tW_t)$.
-The post-juvenile female population in the next year includes both survivors and new recruits: $N_{t+1}=W_t+J_t$.
-Annual population growth rate $\lambda_t$ is set to 0 when $N_t$ is 0, and is otherwise $\lambda_t=N_{t+1}/N_t$.
-For more stable estimates of population status, we report average population growth over a period of $y$ years, calculated as $\bar{\lambda}_{y,t}=\sum_{x=t-y+1}^{x=t}\lambda_x$.
-The initial post-juvenile population size $N_0$ is 1000 animals.
-For reference, all variables and parameters are defined in Table A1.
+We assume a simple two-class demographic model, with a census in late
+winter of surviving females and new recruits to the post-juvenile (aged
+1+) class. The number of post-juvenile females that survive from year
+$t$ to the census $W_t$ is binomially distributed with survival
+probability $S_t$: $W_{t} \sim \text{Binomial}(N_t,S_t)$. The number of
+juveniles recruiting to the post-juvenile class at the census is a
+Poisson distributed function of the number of surviving post-juvenile
+females and the adjusted recruitment rate $X_t$:
+$J_{t} \sim \text{Poisson}(X_tW_t)$. The post-juvenile female population
+in the next year includes both survivors and new recruits:
+$N_{t+1}=W_t+J_t$. Annual population growth rate $\lambda_t$ is set to 0
+when $N_t$ is 0, and is otherwise $\lambda_t=N_{t+1}/N_t$. For more
+stable estimates of population status, we report average population
+growth over a period of $y$ years, calculated as
+$\bar{\lambda}_{y,t}=\sum_{x=t-y+1}^{x=t}\lambda_x$. The initial
+post-juvenile population size $N_0$ is 1000 animals. For reference, all
+variables and parameters are defined in Table A1.
 
-$$J_{t} \sim \text{Poisson}(W_tR_t/2)$$
-<!--
+$$J_{t} \sim \text{Poisson}(W_tR_t/2)$$ <!--
 Annotated equations for poster.
@@
 
-TO DO: 
-* note distribution of random effect parameters are truncated to ensure R and S remain within reasonable bounds (not guaranteed given log link). Not yet documented here.
-* note Eacker survival method gives biased results in years when <12 months of survival data is available.
-* Note I am dropping survival information for animals that are not alive at the beginning of a year to deal with censoring issues.
-There is probably some more elegant way to deal with this, but it won't make any difference as long as the collar deployment month is set to 1.
-* consider & comment on alternative choices made in this example https://poissonconsulting.shinyapps.io/bboushiny/. Follow Eacker example in all particulars, change some aspects, or do comparisons?
-* Ensure notation consistency amongst manuscript, UI, documentaiton and code where appropriate.
+TO DO: \* note distribution of random effect parameters are truncated to
+ensure R and S remain within reasonable bounds (not guaranteed given log
+link). Not yet documented here. \* note Eacker survival method gives
+biased results in years when \<12 months of survival data is available.
+\* Note I am dropping survival information for animals that are not
+alive at the beginning of a year to deal with censoring issues. There is
+probably some more elegant way to deal with this, but it won't make any
+difference as long as the collar deployment month is set to 1. \*
+consider & comment on alternative choices made in this example
+<https://poissonconsulting.shinyapps.io/bboushiny/>. Follow Eacker
+example in all particulars, change some aspects, or do comparisons? \*
+Ensure notation consistency amongst manuscript, UI, documentaiton and
+code where appropriate.
 
@@
 
-In order to assess whether a monitoring program is likely sufficient to inform decisions, it is important to select outcome measures that are relevant to decision making [@gregory_structured_2012].
-For boreal caribou and other endangered species, it is important to know whether a population is declining or not.
-If uncertainty about population status is low, additional investment in reducing uncertainty about the population growth rate may not be necessary or warranted.
-Thus, in this analysis our outcome of interest is how often the Bayesian posterior mean correctly predicts the true population status (increasing or decreasing).
-To calculate this outcome for each disturbance and monitoring scenario (Figure \@ref(fig:figDisturbance)) we fit the Bayesian model to 200 example population trajectories (e.g.
-Figure \@ref(fig:exampleTrajectories)).
-The probability of correct status assessment for that scenario is the percentage of these 200 examples in which the Bayesian posterior mean population growth rate and the true population growth rate are both above or below 1.
+In order to assess whether a monitoring program is likely sufficient to
+inform decisions, it is important to select outcome measures that are
+relevant to decision making [@gregory_structured_2012]. For boreal
+caribou and other endangered species, it is important to know whether a
+population is declining or not. If uncertainty about population status
+is low, additional investment in reducing uncertainty about the
+population growth rate may not be necessary or warranted. Thus, in this
+analysis our outcome of interest is how often the Bayesian posterior
+mean correctly predicts the true population status (increasing or
+decreasing). To calculate this outcome for each disturbance and
+monitoring scenario (Figure \@ref(fig:figDisturbance)) we fit the
+Bayesian model to 200 example population trajectories (e.g. Figure
+\@ref(fig:exampleTrajectories)). The probability of correct status
+assessment for that scenario is the percentage of these 200 examples in
+which the Bayesian posterior mean population growth rate and the true
+population growth rate are both above or below 1.
 
@@
 
-Integration into caribouMetrics, shiny UI, integration into SpaDES & SyncroSim.
+Integration into caribouMetrics, shiny UI, integration into SpaDES &
+SyncroSim.
+
+To make this work more user friendly and accessible to decision makers
+we built a user interface that allows for the exploration of different
+monitoring and anthropogenic disturbance scenarios. We have also made
+the R code and functions needed to carry out this analysis available as
+an R package caribouMetrics. We built the user interface as an R Shiny
+app [@chang2023], building off and expanding an app published by Eacker
+et al. [-@eacker_web-based_2019]. The app allows users to provide
+existing local data or simulate data based on national models as shown
+above, define an anthropogenic disturbance scenario and make
+projections. The results shown are similar to Figure
+\@ref(fig:examplePredictions) and allow for comparisons to the national
+model. Many parameters controlling the Bayesian model, local data
+simulation and antropogenic scenario are exposed in the app with
+sensible defaults set, such that novice users can get reasonable results
+quickly while advanced user can customise the projections to suite their
+needs. The caribouMetrics package is available on GitHub at
+<https://github.com/LandSciTech/caribouMetrics> and a package website
+with detailed documentation is also available
+(<https://landscitech.github.io/caribouMetrics/>).
 
@@
 
-Will need a section on the fundamental assumptions of these models/aspatial nature of them, e.g.
-that collared cows  and cows observed in surveys are a random representative sample of the larger population, and that there is full mixing in the population so that estimated survival/recruitment rates apply equally to all animals in the population.
-We can discuss this with some contrasting examples, e.g.
-collaring 30 animals from a very small/restricted population (like Little Smoky, 308,606ha, pop size estimate 78 as of COSEWIC 2014) would likely be much more representative of the population vs.
-collaring 30 animals from a relatively large/expansive population (northern Quebec range Q6, 62,156,186 ha, pop size 9000 as of COSEWIC 2014).
-But because of the way this model is built, we assume that the sample is equally representative in either of those cases, so changing the initial starting N won't affect model performance/predictions much here...
+Will need a section on the fundamental assumptions of these
+models/aspatial nature of them, e.g. that collared cows and cows
+observed in surveys are a random representative sample of the larger
+population, and that there is full mixing in the population so that
+estimated survival/recruitment rates apply equally to all animals in the
+population. We can discuss this with some contrasting examples, e.g.
+collaring 30 animals from a very small/restricted population (like
+Little Smoky, 308,606ha, pop size estimate 78 as of COSEWIC 2014) would
+likely be much more representative of the population vs. collaring 30
+animals from a relatively large/expansive population (northern Quebec
+range Q6, 62,156,186 ha, pop size 9000 as of COSEWIC 2014). But because
+of the way this model is built, we assume that the sample is equally
+representative in either of those cases, so changing the initial
+starting N won't affect model performance/predictions much here...
 
-On a related topic to population size; Note that the stuff we got from ArtR regarding simulations for sample size does acknowledge that pops of different sizes have different monitoring requirements, and Sara's Labrador numbers also show targets relative to pop size...
-Since this kind of aspatial model will mostly be using small population sizes in order to see any effects, we'll jsut have to acknowledge that a caveat of this is that we're not fully able to explore all the population size effects with this model, but generally Josie has found little impact of initial population size to the results.
+On a related topic to population size; Note that the stuff we got from
+ArtR regarding simulations for sample size does acknowledge that pops of
+different sizes have different monitoring requirements, and Sara's
+Labrador numbers also show targets relative to pop size... Since this
+kind of aspatial model will mostly be using small population sizes in
+order to see any effects, we'll jsut have to acknowledge that a caveat
+of this is that we're not fully able to explore all the population size
+effects with this model, but generally Josie has found little impact of
+initial population size to the results.
 
-Include some commentary on addressing limitations of aspatial demographic models with spatial models.
-Not a panacea.
-It is possible (cite examples) to develop spatially explicit models of movement and adult mortality risk - though sample size is a challenge.
-Recruitment, which is more variable among ranges and with anthropogenic disturbance, is difficult to characterize at finer spatial scales because it depends on the conditions experienced by mothers over years as they move across the range, and on spatial variation in calf mortality risk that is not measured directly.
+Include some commentary on addressing limitations of aspatial
+demographic models with spatial models. Not a panacea. It is possible
+(cite examples) to develop spatially explicit models of movement and
+adult mortality risk - though sample size is a challenge. Recruitment,
+which is more variable among ranges and with anthropogenic disturbance,
+is difficult to characterize at finer spatial scales because it depends
+on the conditions experienced by mothers over years as they move across
+the range, and on spatial variation in calf mortality risk that is not
+measured directly.
 
-We should likely also have a pgph about costs of monitoring - it'll be pretty short, cause basically we have no idea what it costs since nobody will tell us :) Through the NBCKC Monitoring Working Group work, we tried over and over to get some basic cost estimates from P/T reps and practitioners, and hit a wall - nobody wanted to tell us a thing.
-Don't know if we'd now find anyone is more open to sharing, but don't think we should get our hopes up.
-But at the very least we'll want to acknowledge that we haven't accounted for relative costs of monitoring...
-[e.g.
-costs outlined in the Telemetry guidance report from the MonitoringWG: "Expenses to consider in radio-collared monitoring will vary, based on: helicopter (or other aircraft) use, purchase of tracking collars, data fees associated with the collars (the cost of which varies by operating system, recording frequency and data transmission frequency), and live-capture equipment (3 or more sets of capture nets and associated buckets, a netting gun, and ankle hobbles)"]
--->perhaps the keen MonitoringWG contacts would be willing to provide even really rough cost estimates, and/or even jsut a relative sense of how different it would be to e.g.
-monitor a few pops for many yrs vs lots of pops for a short time...?
+We should likely also have a pgph about costs of monitoring - it'll be
+pretty short, cause basically we have no idea what it costs since nobody
+will tell us :) Through the NBCKC Monitoring Working Group work, we
+tried over and over to get some basic cost estimates from P/T reps and
+practitioners, and hit a wall - nobody wanted to tell us a thing. Don't
+know if we'd now find anyone is more open to sharing, but don't think we
+should get our hopes up. But at the very least we'll want to acknowledge
+that we haven't accounted for relative costs of monitoring... [e.g.
+costs outlined in the Telemetry guidance report from the MonitoringWG:
+"Expenses to consider in radio-collared monitoring will vary, based on:
+helicopter (or other aircraft) use, purchase of tracking collars, data
+fees associated with the collars (the cost of which varies by operating
+system, recording frequency and data transmission frequency), and
+live-capture equipment (3 or more sets of capture nets and associated
+buckets, a netting gun, and ankle hobbles)"] --\>perhaps the keen
+MonitoringWG contacts would be willing to provide even really rough cost
+estimates, and/or even jsut a relative sense of how different it would
+be to e.g. monitor a few pops for many yrs vs lots of pops for a short
+time...?
 
-Re paragraph on costs of monitoring.
-The reality is that costs vary with remoteness and the difficulty of finding animals to collar (which also depends on how many collars are deployed).
-Since costs are so context-specific, there isn't a good way to incorporate them into this general analysis.
-It would be interesting to work through a case study (i.e.
-Newfoundland) informed by context-specific costs, and consider how incorporation of those costs may change the relative value of alternative monitoring scenarios.
+Re paragraph on costs of monitoring. The reality is that costs vary with
+remoteness and the difficulty of finding animals to collar (which also
+depends on how many collars are deployed). Since costs are so
+context-specific, there isn't a good way to incorporate them into this
+general analysis. It would be interesting to work through a case study
+(i.e. Newfoundland) informed by context-specific costs, and consider how
+incorporation of those costs may change the relative value of
+alternative monitoring scenarios.
 
-Biases in composition survey estimates also vary with vegetation cover (i.e.
-the difficulty of seeing animals) and the care taken to correctly identify the sex of unknown animals.
-Misidentification rates could be estimated from discrepancies among results from multiple independent observers analyzing photos of animal groups.
+Biases in composition survey estimates also vary with vegetation cover
+(i.e. the difficulty of seeing animals) and the care taken to correctly
+identify the sex of unknown animals. Misidentification rates could be
+estimated from discrepancies among results from multiple independent
+observers analyzing photos of animal groups.
 
-The tool we have developed is a proof of concept, designed to demonstrate possibilities and help users develop a general understanding of the relative value of alternative monitoring strategies for boreal caribou.
-In current form the tool will have limited value for managers who have some local data and wish to know the value of collecting more, because it is designed to either integrate real local data or simulated local data, but not both.
-Development of a tool designed to help decision makers assess the relative value of strategies for additional monitoring given their existing local data and national-demographic disturbance relationships would be a useful followup to this work.
+The tool we have developed is a proof of concept, designed to
+demonstrate possibilities and help users develop a general understanding
+of the relative value of alternative monitoring strategies for boreal
+caribou. In current form the tool will have limited value for managers
+who have some local data and wish to know the value of collecting more,
+because it is designed to either integrate real local data or simulated
+local data, but not both. Development of a tool designed to help
+decision makers assess the relative value of strategies for additional
+monitoring given their existing local data and national-demographic
+disturbance relationships would be a useful followup to this work.
 
-A tool that integrates real local data with national demographic disturbance relationships can also be used to provide consistent and transparent assessments of the strength of evidence for local deviations from the expected mean of national demographic-disturbance relationships.
-Help distinguish real variation among populations from apparent variation that can be attributed to lack of local data.
+A tool that integrates real local data with national demographic
+disturbance relationships can also be used to provide consistent and
+transparent assessments of the strength of evidence for local deviations
+from the expected mean of national demographic-disturbance
+relationships. Help distinguish real variation among populations from
+apparent variation that can be attributed to lack of local data.
 
@@
 
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
+This report was generated on `r Sys.time()` using the following
+computational environment and dependencies:
 

### 2023-09-05 analysis/paper/paper.Rmd
edit to UI paragraph

@@
 
-Integration into caribouMetrics, shiny UI, integration into SpaDES &
-SyncroSim.
-
 To make this work more user friendly and accessible to decision makers
 we built a user interface that allows for the exploration of different
-monitoring and anthropogenic disturbance scenarios. We have also made
-the R code and functions needed to carry out this analysis available as
-an R package caribouMetrics. We built the user interface as an R Shiny
-app [@chang2023], building off and expanding an app published by Eacker
-et al. [-@eacker_web-based_2019]. The app allows users to provide
-existing local data or simulate data based on national models as shown
-above, define an anthropogenic disturbance scenario and make
-projections. The results shown are similar to Figure
+monitoring and anthropogenic disturbance scenarios. We built the user
+interface as an R Shiny app [@chang2023], building off and expanding an
+app published by Eacker et al. [-@eacker_web-based_2019]. The app allows
+users to provide existing local data or simulate data based on national
+models as shown above, define an anthropogenic disturbance scenario and
+make projections. The results shown are similar to Figure
 \@ref(fig:examplePredictions) and allow for comparisons to the national
@@
 quickly while advanced user can customise the projections to suite their
-needs. The caribouMetrics package is available on GitHub at
-<https://github.com/LandSciTech/caribouMetrics> and a package website
-with detailed documentation is also available
+needs. We have also made the R code needed to run the app and carry out
+the above analysis available as an R package, caribouMetrics. Making our
+code available as an R package ensures that it is flexible and can be
+re-used or expanded upon in future projects or combined with other tools
+and packages (eg SpaDES or SyncroSym). The caribouMetrics package is
+available on GitHub at <https://github.com/LandSciTech/caribouMetrics>
+and a package website with detailed documentation is also available
 (<https://landscitech.github.io/caribouMetrics/>).

### 2023-09-20 analysis/paper/supplement.Rmd
updates for name change in getPriors function

@@
   #library(ggplot2);fig_widths=data.frame(two=7)
-  KS$SigmaMean = KS$sInterannualVar
-  KS$SigmaSD = KS$sInterannualVarSE
   KS=subset(KS,scn=="All")
@@
 
-  KSRec$SigmaMean = KSRec$rInterannualVar
-  KSRec$SigmaSD = KSRec$rInterannualVarSE
   KSRec=subset(KSRec,scn=="All")
@@
 
-knitr::include_graphics(here::here("figs/SurvInterceptAllsInterannualVarSE0.03.png"))
+knitr::include_graphics(here::here("figs/SurvInterceptAllsSigmaSD0.03.png"))
 ```
@@
 
-knitr::include_graphics(here::here("figs/SurvsurvInterceptLowsInterannualVarSE0.03.png"))
+knitr::include_graphics(here::here("figs/SurvsurvInterceptLowsSigmaSD0.03.png"))
 ```
@@
 Bands show 95% posterior predictive intervals ('local'), and the 2.5th and 97.5th percentiles for 1000 sample populations from the national model.",  cache=TRUE}
-knitr::include_graphics(here::here("figs/RecInterceptAllrInterannualVarSE0.22.png"))
+knitr::include_graphics(here::here("figs/RecInterceptAllrSigmaSD0.22.png"))
 ```
@@
 Bands show 95% posterior predictive intervals, black dots show true recruitment rates from the simulated population, and open triangles show observed recruitment rates.", cache=TRUE}
-knitr::include_graphics(here::here("figs/RecInterceptHighrInterannualVarSE0.22.png"))
+knitr::include_graphics(here::here("figs/RecInterceptHighrSigmaSD0.22.png"))
 ```

### 2023-09-20 R/CaribouDemographicFunctions.R

@@ -43,14 +43,14 @@ runScenario<-function(scns,quants=NULL,Anthro=NULL,survAnalysisMethod="KaplanMei
   return(scResults)
 }
 
-makeInterceptPlots <- function(scResults, addBit = "", facetVars = c("sIntSEMod","sInterannualVar"),
+makeInterceptPlots <- function(scResults, addBit = "", facetVars = c("sIntSEMod","sSigmaMean"),
                                loopVars = NULL,
                                whichPlots = c("Adult female survival",
                                               "Population growth rate",
                                               "Recruitment",
                                               "Female population size"),
                                survLow = 0.6, type = "png", useNational = T) {
-  # facetVars=c("sIntSEMod","sInterannualVar");loopVars="sInterannualVarSE";scResults=scResults
+  # facetVars=c("sIntSEMod","sSigmaMean");loopVars="sSigmaSD";scResults=scResults
   #useNational=T;type="png"; survLow=0.6
   if (!is.null(loopVars)) {
     loopSet <- unique(subset(scResults$rr.summary.all, select = loopVars))

### 2023-09-20 analysis/scripts/calibratePriors.R

@@ -6,45 +6,45 @@ scnsLow = data.frame(obsYears=15,collarCount=60,cowMult=9,projYears=1,rQuantile=
 scnsHigh = scnsLow;scnsHigh$rQuantile =0.975;scnsHigh$sQuantile=0.975
 ########################
 #Survival
-#vary sIntSEMod, sInterannualVar, and sInterannualVarSE in factorial array. Outcome of interest is KS distance
-sIntSEMod=seq(1,10,by=2);sInterannualVar=0.08696*seq(0.2,1,by=0.2);sInterannualVarSE=seq(0.01,0.09,by=0.02)
+#vary sIntSEMod, sSigmaMean, and sSigmaSD in factorial array. Outcome of interest is KS distance
+sIntSEMod=seq(1,10,by=2);sSigmaMean=0.08696*seq(0.2,1,by=0.2);sSigmaSD=seq(0.01,0.09,by=0.02)
 
 #Looking for low KS distance from full range of input sims when given almost no info
 scnsSInt=expand.grid(obsAnthroSlope = 0, projAnthroSlope = 0,sIntSEMod=sIntSEMod,
-                     sInterannualVar=sInterannualVar,sInterannualVarSE=sInterannualVarSE)
-scnsSInt$SigmaMean = scnsSInt$sInterannualVar
+                     sSigmaMean=sSigmaMean,sSigmaSD=sSigmaSD)
+scnsSInt$SigmaMean = scnsSInt$sSigmaMean
 scns=merge(scnsSInt,scnsApplyAll);scns=merge(scns,scnsNoData)
 scResults = runScenario(scns,survAnalysisMethod = "Exponential")
 KSAll = subset(scResults$ksDists,(Year==2023)&(Parameter=="Adult female survival"))
 addBit = paste0("InterceptAll")
 scResults$obs.all = NULL
-makeInterceptPlots(scResults,addBit=addBit,facetVars=c("SigmaMean","sIntSEMod"),loopVars="sInterannualVarSE",whichPlots=c("Adult female survival"))
+makeInterceptPlots(scResults,addBit=addBit,facetVars=c("SigmaMean","sIntSEMod"),loopVars="sSigmaSD",whichPlots=c("Adult female survival"))
 
 scns=merge(scnsSInt,scnsApplyAll);scns=merge(scns,scnsLow)
[…]

### 2023-09-20 analysis/scripts/bayesianModelExamples.R
changes to bayesian model examples figure to align legend and scales

@@ -38,32 +38,58 @@ labFontSize = 10; breakInterval=5
 
 priorResult = caribouMetrics:::runScnSet(scns[1,],eParsIn,simBig,getKSDists=F,printProgress=F)
 priorResult$obs.all=NULL
+yr_scale1 <- scale_x_continuous(breaks = 2023:2072 %>% .[0:9*5+1],
+                                labels = 2009:2058 %>% .[0:9*5+1] )
+yr_scale2 <- scale_x_continuous(breaks = 2009:2058 %>% .[0:9*5+1],
+                                labels = 2009:2058 %>% .[0:9*5+1] )
 recPrior =  plotRes(priorResult, "Recruitment", lowBound=0, highBound = 0.75,
-                    legendPosition="left",breakInterval=breakInterval,labFontSize=labFontSize)
+                    legendPosition="none",breakInterval=breakInterval,
+                    labFontSize=labFontSize) +
+  yr_scale1 +
+  labs(tag = "a")
 plot(recPrior)
 survPrior =  plotRes(priorResult, "Adult female survival", lowBound=0.6,
-                     legendPosition="none",breakInterval=breakInterval,labFontSize=labFontSize)
+                     legendPosition="none",breakInterval=breakInterval,
+                     labFontSize=labFontSize)+
+  yr_scale1
 plot(survPrior)
 lambdaPrior =  plotRes(priorResult, "Population growth rate", lowBound=0.6,
-                       legendPosition="none",breakInterval=breakInterval,labFontSize=labFontSize)
+                       legendPosition="none",breakInterval=breakInterval,
+                       labFontSize=labFontSize)+
+  yr_scale1 +
+  ylim(c(0, 1.2))
 plot(lambdaPrior)
 
 posteriorResult = caribouMetrics:::runScnSet(scns[2,],eParsIn,simBig,getKSDists=F,printProgress=F)
 recPosterior =  plotRes(posteriorResult, "Recruitment", lowBound=0,highBound = 0.75,
-                        legendPosition="left",breakInterval=breakInterval,labFontSize=labFontSize)
+                        legendPosition="none",breakInterval=breakInterval,
+                        labFontSize=labFontSize)+
+  yr_scale2 +
+  labs(tag = "b")
 plot(recPosterior)
 survPosterior =  plotRes(posteriorResult, "Adult female survival", lowBound=0.6,
[…]

### 2023-09-20 analysis/scripts/bayesianModelExamples.R
fix background color

@@ -87,7 +87,7 @@ leg <- ggpubr::get_legend(leg)
 plts <- ggpubr::ggarrange(recPrior, survPrior, lambdaPrior,
                           recPosterior, survPosterior, lambdaPosterior, labels = "",
                           ncol = 3, nrow = 2, vjust = 1)
-ggpubr::ggarrange(plts, leg, ncol = 2, widths = c(6,1))
+ggpubr::ggarrange(plts, leg, ncol = 2, widths = c(6,1))+bgcolor("white")
 
 ggsave(paste0(baseDir,"/analysis/paper/figs/bayesianModelExamples.png"),
        width = 12*0.8, height = 3.6*2, units = "in",
